### 2021-04-14 analysis/OTT.Rmd
wflow_publish("analysis/OTT.Rmd")

---
title: "R Notebook"
output: html_notebook
---

Here I make a subtree from the Open Tree Of Life API for the species we have.
```{r setup}
library(rotl)
library(ggtree)
library(cowplot)
library(ggplot2)
library(tidyverse)
```

```{r}
species <- c('Zostera marina', 'Zostera muelleri', 'Arabidopsis thaliana', 'Thellungiella parvula', 'Populus trichocarpa', 'Vitis vinifera', 'Amborella trichopoda', 'Oryza sativa', 'Zea mays', 'Brachypodium distachyon', 'Spirodela polyrhiza', 'Selaginella moellendorffi', 'Physcomitrella patens', 'Chlamydomonoas reinhardtii', 'Ostreococcus lucimarinus', 'Lemna gibba', 'Posidonia australis' ,'Amphibolis antarctica')
```

Let's call the API to get the OTT IDs for these species, and  to double-check we got the right names :)

We should get a tree for `r length(species)` species.

```{r}
taxon_search <- tnrs_match_names(names = species, context_name = "All life")
```

```{r}
knitr::kable(taxon_search)
```

Let's write the species names out again for timetree.org
```{r}
cat(capture.output(cat(taxon_search$unique_name, sep='\n'), file="data/species.csv"))
```

That looks good to me!

```{r}
ott_in_tree <- ott_id(taxon_search)[is_in_tree(ott_id(taxon_search))]
```

```{r}
tr <- tol_induced_subtree(ott_ids = ott_in_tree)
```

```{r}
plot(tr)
```

Let's get rid of those OTT IDs from the tree, I don't like that. There's a helper function for that

```{r}
tr$tip.label <- strip_ott_ids(tr$tip.label, remove_underscores = TRUE)
```

```{r}
ggtree(tr) + 
  geom_tree() + 
  geom_tiplab(size=3,
              fontface='italic') +
  theme_tree() +
  xlim(0, 15)
```

Cool! I took the above species list and uploaded to timetree (see Species_timetree.svg)

```{r fig.width=14}
plot(ggdraw()+cowplot::draw_image('data/Species_timetree.svg'))
```

Timetree also lets you export the tree timed with MYA

```{r}
tree <- read.tree('data/timetree_species.nwk')
```

```{r}
tree$tip.label <- str_replace_all(tree$tip.label, '_', ' ')
p <- ggplot(tree) + 
  geom_tree() + 
  theme_tree2() +
  scale_x_continuous(labels = abs) + 
  geom_tiplab(size=3,  fontface='italic')

revts(p) + xlim(-1200, 300)
```

### 2021-04-14 analysis/index.Rmd
wflow_publish(c("analysis/plot.Rmd", "analysis/_site.yml", "analysis/index.Rmd"),     republish = TRUE)

---
title: "Home"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: console
---

This website collects comparisons of Amphibolis and Posidonia comparative analyses.

[Repeat comparison plots](plot.html)

[Phylogenies](OTT.html)

### 2021-04-14 analysis/plot.Rmd
wflow_publish(c("analysis/plot.Rmd", "analysis/_site.yml", "analysis/index.Rmd"),     republish = TRUE)

---
title: "Untitled"
author: "Philipp Bayer"
date: "4 March 2021"
output: html_document
---

```{r setup}
library(tidyverse)
library(cowplot)
theme_set(theme_cowplot())
library(RColorBrewer)
library(patchwork)
```

```{r}
# reps <- data.frame(
#   stringsAsFactors = FALSE,
#              Class = c("LINEs","LTRs","DNA",
#                        "Unclassified","Non-repetitive","LINEs","LTRs","DNA","Unclassified",
#                        "Non-repetitive","LINEs","LTRs","DNA","Unclassified",
#                        "Non-repetitive","LINEs","LTRs","DNA","Unclassified","Non-repetitive"),
#           Assembly = c("A. antarctica","A. antarctica",
#                        "A. antarctica","A. antarctica","A. antarctica","P. australis",
#                        "P. australis","P. australis","P. australis","P. australis","Z. marina",
#                        "Z. marina","Z. marina","Z. marina","Z. marina","Z. muelleri",
#                        "Z. muelleri","Z. muelleri","Z. muelleri","Z. muelleri"),
#                 BP = c(805898L,45776664L,37822890L,
#                        1546483L,158725395L,38319733L,611240734L,100034681L,
#                        37142931L,428369702L,3103364L,123623579L,27390818L,
#                        6060480L,96439714L,14331182L,164253396L,142828758L,
#                        21922574L,274954485L),
#            Percent = c(0.33,18.71,15.46,0.63,
#                        64.87131,3.15,50.3,8.23,3.06,35.25364,1.19,47.46,10.52,
#                        2.33,37.02,2.32,26.57,23.1,3.55,44.46)
# )

reps <- readxl::read_xlsx('data/Repeat_coding.xlsx')

reps <- reps %>% dplyr::filter(Class != 'Total')
reps$Class <- factor(reps$Class, levels = c('LINEs', 'LTRs', 'DNA', 'Unclassified', 'Total CDS', 'Non-repetitive'))
reps
```

```{r}
p1 <- reps %>% ggplot(aes(x=Assembly, fill=Class, y = BP/1000000)) + 
geom_bar(position='stack', stat='identity') + ylab('Size (Mbp)') +
  scale_fill_brewer(palette='Dark2') +
  theme(axis.text.x =  element_text(face="italic"))

p1
```

```{r}
p2 <- reps %>% ggplot(aes(x=Assembly, fill=Class, y = Percent)) + geom_bar(position='stack', stat='identity')+
  scale_fill_brewer(palette='Dark2') +
  theme(axis.text.x =  element_text(face="italic")) +
  ylab('Percent\n of assembly')
p2
```
```{r fig.width = 10, dpi=300}
p1/p2 + plot_annotation(tag_levels = 'A')
```

```{r}
reps %>% filter(Class=='Non-repetitive') %>% ggplot(aes(x=Assembly, fill=Class, y = BP)) + geom_bar(position='stack', stat='identity')

```

### 2021-04-15 analysis/Orthofinder.Rmd
wflow_publish(files = list.files("analysis/", pattern = "*Rmd",     full.names = T))

---
title: "Clustering/Orthofinder analysis"
author: "Philipp Bayer"
date: "15 April 2021"
output: html_document
---

```{r setup}
library(tidyverse)
library(cowplot)
theme_set(theme_cowplot())
library(RColorBrewer)
library(patchwork)
library(UpSetR)
```

```{r}
groups <- read_tsv('data/Orthogroups.tsv')

names(groups) <- c('Orthogroup', 'A. antarctica', 'A. trichopada', 'B. distachyon', 'C. reinhardtii', 'L. gibba', 'O. sativa', 'P. australis', 'P. patens', 'P. trichocarpa', 'S. moellendorffii', 'S. polyrhiza', 'A. thaliana', 'T. parvula', 'V. vinifera', 'Z. mays', 'Z. muelleri', 'Z. marina', 'O. lucimarinus')
```

```{r}
# for upsetr, we need to know only which OG-groups are shared between species, the actual genes don't matter
per_spec <- groups %>% pivot_longer(-Orthogroup) %>% 
  filter(!is.na(value)) %>% # species not in an orthogroup are still listed, they just have NA genes for this group
  select(-value) # don't need all gene names, speed things up
```

```{r}
# now I want the data in this format:
# listInput <- list(one = c(1, 2, 3, 5, 7, 8, 11, 12, 13), two = c(1, 2, 4, 5, 
#   10), three = c(1, 5, 6, 7, 8, 9, 10, 12, 13))
x <- per_spec %>% 
  select(name, Orthogroup) %>% # turn the table around
  deframe() # convert to named vector
```

```{r}
mylist <- lapply(split(x, names(x)), unname) # yuck - ugly code to convert the named vector to a list
```

```{r}
x <- upset(fromList(mylist), order.by='freq', nsets = length(groups) - 1)
```

```{r fig.width=10}
x
```

Let's get the species-only cluster numbers

```{r}
species_specific_orthos <- per_spec %>% 
  group_by(Orthogroup) %>% 
  summarise(counts = length(name)) %>%
  filter(counts == 1)
```

```{r}
per_spec %>% 
  filter(Orthogroup %in% species_specific_orthos$Orthogroup) %>% 
  group_by(name) %>% 
  count() %>% 
  arrange(n) %>% 
  knitr::kable()
```

```{r}
species_names <- per_spec %>% 
  group_by(Orthogroup) %>% 
  group_by(Orthogroup) %>% 
  summarise(catty = paste0(sort(name), collapse = ', '))
```

How many orthogroups are shared between the four seagrasses?

```{r}
species_names %>% dplyr::filter(catty == 'A. antarctica, P. australis, Z. marina, Z. muelleri') %>% nrow()
```

```{r}
species_names %>% dplyr::filter(catty == 'A. antarctica, P. australis') %>% nrow()

```

```{r}
species_names %>% dplyr::filter(catty == 'A. antarctica, P. australis, Z. marina') %>% nrow()

```
```{r}
species_names %>% dplyr::filter(catty == 'A. antarctica, P. australis, Z. muelleri') %>% nrow()
```

```{r}
species_names %>% dplyr::filter(catty == 'P. australis, Z. marina, Z. muelleri') %>% nrow()
```

I should automate this.... and I can!

```{r}
newlist <- mylist[c('A. antarctica', 'Z. marina', 'P. australis', 'Z. muelleri')]

x <- upset(fromList(newlist), order.by='freq', nsets = 4)

```
```{r}
x
```

This can't be right, how are 6532 clusters suddenly shared.... that's less than the overall. Need to subset the original clusters, NOT the species.

### 2021-10-07 analysis/GOenrichment.Rmd
wflow_publish(files = c("analysis/*"))

---
title: "GO enrichment"
author: "Philipp Bayer"
date: "01/10/2021"
output: html_document
---

```{r setup, include=FALSE}
library(topGO)
library(tidyverse)
library(rvest)
library(wesanderson)
library(httr)
library(stringi)
library(ggrepel)
library(eulerr)
library(UpSetR)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

# Lost in seagrasses

The GO enrichment does not work well on my laptop so I'm setting this to eval=FALSE and run it on a remote server.
The script which writes the input files is a Python script in the scripts/ folder: findClustersUniqueToAquatic.py. This script parses Orthofinder output to pull out genome-specific groups and genes.

```{r eval=FALSE}
# give properly formatted background in format: GO:0005838	GSBRNA2T00088508001;GSBRNA2T00088313001;GSBRNA2T00035842001 
#annAT <- readMappings('BACKGROUND.txt', sep="\t", IDsep=";")
#save(annAT, file='annAtObject.RData')

load('annAtObject.RData')
allgenes <- unique(unlist(annAT))

compare <- function(genelistfile, outname, allgenes, annAT) {
  # give file with your genes of interest, one gene_id per line

  mygenes <-scan(genelistfile ,what="")
  geneList <- factor(as.integer(allgenes %in% mygenes))
  names(geneList) <- allgenes

  GOdata <-new ("topGOdata", ontology = 'BP', allGenes = geneList, nodeSize = 5, annot=annFUN.GO2genes, GO2genes=annAT)
  # using ClassicCount:
  #test.stat <-new ("classicCount", testStatistic = GOFisherTest, name = "Fisher Test")
  #resultsFisherC <-getSigGroups (GOdata, test.stat)
  # using weight01:
  weight01.fisher <- runTest(GOdata, statistic = "fisher")
  # using ClassicCount:
  # allRes <- GenTable(GOdata, classicFisher= resultsFisherC, topNodes = 30)
  # using weight01:
  allRes <- GenTable(GOdata, classicFisher=weight01.fisher,topNodes=30)#,topNodes=100)
  names(allRes)[length(allRes)] <- "p.value"
  p_values <- score(weight01.fisher)
  adjusted_p <- p.adjust(p_values)
  adjusted_p[adjusted_p < 0.05] %>% enframe() %>% write_csv('data/' + outname)

}

```

# lost in seagrasses, each one uniquely vs all terrestrials
```{r eval=FALSE}
compare('lost_in_amphi_vs_all.txt', 'missing_amphi_vs_all_GO.txt', allgenes, annAT)
compare('lost_in_posi_vs_all.txt', 'missing_posi_vs_all_GO.txt', allgenes, annAT)
compare('lost_in_zmar_vs_all.txt', 'missing_zmar_vs_all_GO.txt', allgenes, annAT)
compare('lost_in_zmuel_vs_all.txt', 'missing_zmuel_vs_all_GO.txt', allgenes, annAT)
```

# lost in seagrasses vs all terrestrials
Here we compare GO terms for seagrasses and aquatics (seagrasses+duckweeds) vs all terrestrials

```{r eval=FALSE}

compare('lost_in_seagrasses.txt', 'missing_seagrasses_GO.txt', allgenes, annAT)
compare('lost_in_aquatics.txt', 'missing_aquatics_GO.txt', allgenes, annAT)
compare('only_in_seagrasses.txt', 'only_seagrasses_GO.txt', allgenes, annAT)

```

# Seagrass only comparisons

Now we compare seagrasses within each other.

For the seagrass-only comparisons, I'm using a Seagrass-only background as that makes more biological sense to me.s

```{r eval=FALSE}
# give properly formatted background in format: GO:0005838	GSBRNA2T00088508001;GSBRNA2T00088313001;GSBRNA2T00035842001 
#sannAT <- readMappings('SEAGRASSBACKGROUND.txt', sep="\t", IDsep=";")
#save(sannAT, file='sannAtObject.RData')

load('annAtObject.RData')
sallgenes <- unique(unlist(sannAT))

compare('lost_in_amphi.txt', 'lost_in_amphi_vs_seagrasses_GO.txt', sallgenes, sannAT)
compare('lost_in_posi.txt', 'lost_in_posi_vs_seagrasses_GO.txt', sallgenes, sannAT)
compare('lost_in_zmar.txt', 'lost_in_zmar_vs_seagrasses_GO.txt', sallgenes, sannAT)
compare('lost_in_zmuel.txt', 'lost_in_zmuel_vs_seagrasses_GO.txt', sallgenes, sannAT)
```

# Present in seagrasses

The opposite - which GO-terms are present only in one of the four species?

```{r eval=FALSE}
compare('only_in_amphi.txt', 'only_in_amphi_vs_seagrasses_GO.txt', sallgenes, sannAT)
compare('only_in_posi.txt', 'only_in_posi_vs_seagrasses_GO.txt', sallgenes, sannAT)
compare('only_in_zmar.txt', 'only_in_zmar_vs_seagrasses_GO.txt', sallgenes, sannAT)
compare('only_in_zmuel.txt', 'only_in_zmuel_vs_seagrasses_GO.txt', sallgenes, sannAT)
```

Alright now we have all these different GO terms in all these files - we can send them to revigo for visualiation and some deduplication!

# Revigo

This code is based on http://revigo.irb.hr/CodeExamples/revigo.R.txt

```{r}
results_list <- list()

for (f in list.files('./data/', pattern='GO.txt')){
  filename <- paste('./data/', f, sep='')
  go_and_pvalues <- readChar(filename, file.info(filename)$size)
  go_and_pvalues <- gsub(',', ' ', go_and_pvalues)

  httr::POST(
    url = "http://revigo.irb.hr/StartJob.aspx",
    body = list(
      cutoff = "0.7",
      valueType = "pvalue",
     # speciesTaxon = "4577", # zea mays
     #speciesTaxon = '39947', # japonica
     speciesTaxon = '3702', # arabidopsis
      measure = "SIMREL",
      goList = go_and_pvalues
    ),
    # application/x-www-form-urlencoded
    encode = "form"
  ) -> res

  dat <- httr::content(res, encoding = "UTF-8")

  jobid <- jsonlite::fromJSON(dat,bigint_as_char=TRUE)$jobid

  # Check job status
  running <- "1"
  while (running != "0" ) {
      httr::POST(
        url = "http://revigo.irb.hr/QueryJobStatus.aspx",
        query = list( jobid = jobid )
      ) -> res2
      dat2 <- httr::content(res2, encoding = "UTF-8")
      running <- jsonlite::fromJSON(dat2)$running
      Sys.sleep(1)
  }

  # Fetch results
  httr::POST(
    url = "http://revigo.irb.hr/ExportJob.aspx",
    query = list(
      jobid = jobid, 
      namespace = "1",
      type = "CSVTable"
    )
  ) -> res3

  dat3 <- httr::content(res3, encoding = "UTF-8")

  dat3 <- stri_replace_all_fixed(dat3, "\r", "")
  # Now we have a csv table in a string!

  # read_csv does not like the ', ', it wants ','
  dat <- read_csv(gsub(', ', ',', dat3), show_col_types = FALSE)

  # do we even have results?
  if(nrow(dat) == 0){next}
  results_list[[f]] <- dat

}
```

OK we have a list with all results in a big list. Now we can plot!

```{r warning=FALSE, echo=FALSE, message=FALSE}
# lots of warnings and messages here so I'm hiding these
plot_list <- list()
for (f in names(results_list)) {
  dat <- results_list[[f]]
  if(nrow(dat) == 0) {next}

  names(dat) <- c("term_ID","description","frequency", 'plot_X', 'plot_Y', 'log_size', 'value', 'uniqueness', 'dispensability', 'eliminated', 'representative')
  one.data <- dat

  one.data <- one.data [(one.data$plot_X != "null" & one.data$plot_Y != "null") & (! is.na(one.data$frequency)) & (one.data$value != 'null'), ];
  one.data <- type_convert(one.data)
  one.data$pvalue <- p.adjust(exp(one.data$value))
  #one.data <- one.data %>% filter(pvalue < 0.05) # NO FILTERING OF P-VALUES - for now?

  p1 <-
    ggplot(data = one.data, aes(plot_X, plot_Y, label = description)) +
    geom_point(aes(colour = log(pvalue), size = log_size), alpha = I(0.6)) +
    scale_size_area() +
    scale_colour_gradientn(colours = rev(wes_palette("Zissou1", 100, type = "continuous")), 
                           limits = c(min(log(one.data$pvalue)), 0)) +
    geom_point(
      aes(plot_X, plot_Y, size = log_size),
      shape = 21,
      fill = "transparent",
      colour = I (alpha ("black", 0.6))
    ) +
    scale_size_area() +
    geom_label_repel(max.overlaps = 5, box.padding = 0.5,
                     aes(point.size = log_size)) +
    theme_minimal() +
    labs(y = "Semantic space x",
         x = "Semantic space y",
         color = 'Log(p-value)',
         size = 'Frequency') +
    theme(
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank()
    ) +
    ggtitle(f)

  plot_list[[f]] <- p1

}
```

# Venn diagram GO overlaps

Let's also check which GO terms overlap between the 4 seagrasses!

Let's have a look at all of these plots. Manually zooming in leads to ggrepel reloading labels, so on the small scale a lot of these plots don't have labels.
```{r}
plot_list
```

How many shared lost GO-terms are there? Hopefully, all four species will have lost the most GO-terms.
```{r}
a <- list(`P. australis` = results_list$missing_posi_vs_all_GO.txt$Name,
          `A. antarctica` = results_list$missing_amphi_vs_all_GO.txt$Name,
          `Z. marina` = results_list$missing_zmar_vs_all_GO.txt$Name,
          `Z. muelleri` = results_list$missing_zmuel_vs_all_GO.txt$Name,
          `A. thaliana` = results_list$missing_arabidopsis_vs_all_GO.txt$Name)

a_go_ids <- list(`P. australis` = results_list$missing_posi_vs_all_GO.txt$TermID,
          `A. antarctica` = results_list$missing_amphi_vs_all_GO.txt$TermID,
          `Z. marina` = results_list$missing_zmar_vs_all_GO.txt$TermID,
          `Z. muelleri` = results_list$missing_zmuel_vs_all_GO.txt$TermID,
          `A. thaliana` = results_list$missing_arabidopsis_vs_all_GO.txt$TermID)
```

```{r}
plot(euler(a), 
     quantities = TRUE,
     fill = rev(wes_palette("Zissou1", 15, type = 'continuous')),
    alpha = 0.5,
     labels = list(font = 4))
```

```{r}
upset(fromList(a), order.by='freq', )
```

```{r}
a_no_ara <- list(`P. australis` = results_list$missing_posi_vs_all_GO.txt$Name,
          `A. antarctica` = results_list$missing_amphi_vs_all_GO.txt$Name,
          `Z. marina` = results_list$missing_zmar_vs_all_GO.txt$Name,
          `Z. muelleri` = results_list$missing_zmuel_vs_all_GO.txt$Name)
```

What are the shared GO-terms in seagrasses, WITHOUT the Ara loss??
```{r}
setdiff(Reduce(intersect, a_no_ara), Reduce(intersect, a)) %>% enframe() %>% writexl::write_xlsx('./data/shared_lost_genes.xlsx')
```

What if we remove Posidonia?

```{r}
b <- list(`A. antarctica` = results_list$missing_amphi_vs_all_GO.txt$Name,
          `Z. marina` = results_list$missing_zmar_vs_all_GO.txt$Name,
          `Z. muelleri` = results_list$missing_zmuel_vs_all_GO.txt$Name)
intersections <- Reduce(intersect, b)
intersections[grepl('ethylene', intersections)]
```
OK we need a big list of all GO-terms here - which GO-term is lost in which species. That will be a supplementary table.

```{r}
all_species <- c("P. australis","A. antarctica","Z. marina","Z. muelleri", 'A. thaliana')  
all_go_terms <- Reduce(union, a)
all_go_ids <- Reduce(union, a_go_ids)
results_d <- data.frame('GOID' = character(),
                        'GO' = character(),
                        'P. australis' = character(),
                        'A. antarctica' = character(),
                        'Z. marina' = character(),
                        'Z. muelleri' = character(),
                        'A. thaliana' = character())

for (index in seq_along(all_go_terms)){
  go <- all_go_terms[index]
  go_id <- all_go_ids[index]
  specs <- c()
  for (species in names(a)) {
    if ( length(a[[species]][grep(paste('^', go, '$', sep=''), a[[species]])]) > 0 ) {
      specs <- c(specs, species)
    }
  }

  results_d[index,] <- c(go_id, go, gsub('FALSE', 'Present', gsub('TRUE', 'Lost', all_species %in% specs)))
}
```

```{r}
writexl::write_xlsx(results_d, 'data/Lost_GO_terms_in_five_species.xlsx')
```

# after filtering for plant-specific GO-terms

We will use the GO-terms that are plant-specific as identified by the GOMAP paper.
See https://github.com/wkpalan/GOMAP-maize-analysis/blob/main/6.plantSpecific/1.getSppSpecific.R
or https://plantmethods.biomedcentral.com/articles/10.1186/s13007-021-00754-1

```{r}
go_plant <- read_tsv('https://raw.githubusercontent.com/wkpalan/GOMAP-maize-analysis/main/data/go/speciesSpecificGOTerms.txt')
```

```{r}
# taxon 33090 is Viridiplantae
plantSpecificGO <- go_plant %>% dplyr::filter(`NCBITaxon:33090`==1) %>% pull(GOterm)
plantSpecificGO <- c(plantSpecificGO,c("GO:0005575","GO:0008150","GO:0003674"))
```

```{r}
results_d %>% filter(GOID %in% plantSpecificGO) %>% writexl::write_xlsx('data/Lost_GO_terms_in_five_species.PlantSpecific.xlsx')
```

Now let's redo the Venn diagram with those filtered GO-terms

```{r}
filters <- lapply(a_go_ids, function(ch) ch %in% plantSpecificGO)

newa <- list()
for (species in names(filters)) {
  before <- a[[species]]
  after <- before[filters[[species]]]
  newa[[species]] <- after
}

```

```{r}
plot(euler(newa), 
     quantities = TRUE,
     fill = rev(wes_palette("Zissou1", 15, type = 'continuous')),
    alpha = 0.5,
     labels = list(font = 4))
```

Not much difference?

### 2021-10-07 analysis/OTT.Rmd
Add missing files
wflow_publish(files = list.files("analysis/", pattern = "*Rmd",     full.names = T))
wflow_publish(files = c("analysis/*"))

@@
 ---
-title: "R Notebook"
+title: "Phylogenetics"
 output: html_notebook
@@
 ```{r}
-species <- c('Zostera marina', 'Zostera muelleri', 'Arabidopsis thaliana', 'Thellungiella parvula', 'Populus trichocarpa', 'Vitis vinifera', 'Amborella trichopoda', 'Oryza sativa', 'Zea mays', 'Brachypodium distachyon', 'Spirodela polyrhiza', 'Selaginella moellendorffi', 'Physcomitrella patens', 'Chlamydomonoas reinhardtii', 'Ostreococcus lucimarinus', 'Lemna gibba', 'Posidonia australis' ,'Amphibolis antarctica')
+species <- c('Zostera marina', 'Zostera muelleri', 'Arabidopsis thaliana', 'Thellungiella parvula', 'Populus trichocarpa', 'Vitis vinifera', 'Amborella trichopoda', 'Oryza sativa', 'Zea mays', 'Brachypodium distachyon', 'Spirodela polyrhiza', 'Selaginella moellendorffi', 'Physcomitrella patens', 'Chlamydomonoas reinhardtii', 'Ostreococcus lucimarinus', 'Lemna gibba', 'Posidonia australis' ,'Amphibolis antarctica', 'Wolffia australiana')
 ```
@@
 ```{r fig.width=14}
-plot(ggdraw()+cowplot::draw_image('data/Species_timetree.svg'))
+plot(ggdraw()+cowplot::draw_image('./data/species_timetree.svg'))
 ```
@@
 ```{r}
-tree <- read.tree('data/timetree_species.nwk')
+tree <- read.tree('./data/timetree_species.nwk')
 ```

### 2021-10-07 analysis/Orthofinder.Rmd
Missing commits
wflow_publish(files = c("analysis/*"))

@@
 library(UpSetR)
+
 ```
@@
 ```{r}
-groups <- read_tsv('data/Orthogroups.tsv')
+groups <- read_tsv('./data/Orthogroups.tsv')
 
@@
 marina', 'O.
-lucimarinus')
+lucimarinus', 'W.
+australis')
 ```
@@
 
-```{r}
-species_names <- per_spec %>% 
-  group_by(Orthogroup) %>% 
-  group_by(Orthogroup) %>% 
-  summarise(catty = paste0(sort(name), collapse = ', '))
-```
-
 How many orthogroups are shared between the four seagrasses?
-
-```{r}
-species_names %>% dplyr::filter(catty == 'A. antarctica, P. australis, Z. marina, Z. muelleri') %>% nrow()
-```
-
-```{r}
-species_names %>% dplyr::filter(catty == 'A. antarctica, P. australis') %>% nrow()
-
-```
-
-```{r}
-species_names %>% dplyr::filter(catty == 'A. antarctica, P. australis, Z. marina') %>% nrow()
-
-```
-```{r}
-species_names %>% dplyr::filter(catty == 'A. antarctica, P. australis, Z. muelleri') %>% nrow()
-```
-
-```{r}
-species_names %>% dplyr::filter(catty == 'P. australis, Z. marina, Z. muelleri') %>% nrow()
-```
-
-I should automate this.... and I can!
 
@@
 ```
-
-This can't be right, how are 6532 clusters suddenly shared.... that's less than the overall. Need to subset the original clusters, NOT the species.

### 2021-10-07 analysis/plot.Rmd
Add missing files
wflow_publish(files = list.files("analysis/", pattern = "*Rmd",     full.names = T))
wflow_publish(files = c("analysis/*"))

@@
 ---
-title: "Untitled"
+title: "Genomic comparison"
 author: "Philipp Bayer"
@@
 library(RColorBrewer)
+library(wesanderson)
 library(patchwork)
@@
 
-reps <- readxl::read_xlsx('data/Repeat_coding.xlsx')
+reps <- readxl::read_xlsx('./data/Repeat_coding.xlsx')
 
@@
 ```{r}
+pal <- wes_palette("Zissou1", 6, type = "continuous")
 p1 <- reps %>% ggplot(aes(x=Assembly, fill=Class, y = BP/1000000)) + 
 geom_bar(position='stack', stat='identity') + ylab('Size (Mbp)') +
-  scale_fill_brewer(palette='Dark2') +
+  #scale_fill_brewer(palette='Dark2') +
+  scale_fill_manual(values=pal) +  
   theme(axis.text.x =  element_text(face="italic"))
@@
 p2 <- reps %>% ggplot(aes(x=Assembly, fill=Class, y = Percent)) + geom_bar(position='stack', stat='identity')+
-  scale_fill_brewer(palette='Dark2') +
+  #scale_fill_brewer(palette='Dark2') +
+    scale_fill_manual(values=pal) +  
   theme(axis.text.x =  element_text(face="italic")) +
@@
 ```{r fig.width = 10, dpi=300}
-p1/p2 + plot_annotation(tag_levels = 'A')
+patch <- p1/p2 + plot_annotation(tag_levels = 'A')
+patch[[1]] = patch[[1]] + theme(axis.text.x = element_blank(),
+                                        axis.ticks.x = element_blank(),
+                                        axis.title.x = element_blank() )
+
+patch
 ```

### 2021-10-07 analysis/plotGenes.Rmd
wflow_publish(files = c("analysis/*"))

---
title: "Making gene number plots"
author: "Philipp Bayer"
date: "14/09/2021"
output: html_document
---
```{r}
library(tidyverse)
```

This page plots numbers of genes based on GO-terms, no clustering involved.
```{r}
df <- data.frame(
  stringsAsFactors = FALSE,
           Species = c("A. antarctica","O. sativa",
                       "P. australis","Z. muelleri","Z. marina"),
          Salinity = c(3L, 35L, 3L, 9L, 2L),
               PPR = c(254L, 590L, 291L, 913L, 528L),
        Heat.shock = c(42L, 102L, 56L, 139L, 46L),
    Pectinesterase = c(21L, 51L, 30L, 76L, 67L)
)
```

```{r}
df %>% pivot_longer(-Species) %>% ggplot(aes(x=factor(Species, levels=rev(c('O. sativa', 'A. antarctica', 'P. australis', 'Z. marina', 'Z. muelleri'))), y=value)) + 
  geom_bar(stat='identity') + 
  facet_wrap(~name, scales='free_x' ) +
  ylab('Count') + xlab('Species') + 
  cowplot::theme_minimal_hgrid() + coord_flip() +
  theme(axis.text.y = element_text(face = "italic"))
```

### 2021-10-07 analysis/plotRgenes.Rmd
wflow_publish(files = c("analysis/*"))

---
title: "R genes"
output: html_notebook
---
```{r setup}
library(tidyverse)
library(wesanderson)
```

Visualising R-gene differences here

```{r}
pal <- wes_palette("Zissou1", 3, type = "continuous")

df <- readxl::read_xlsx('./data/R_genes.xlsx')
df %>% ggplot(aes(fill=Class, x=Genome, y=Count)) + geom_bar(stat='identity') +
    scale_fill_manual(values=pal) +  
 coord_flip() +  cowplot::theme_minimal_vgrid() +
    theme(axis.text.y = element_text(face = "italic"))

```

```{r}
pal <- wes_palette("Zissou1", 4, type = "continuous")

df2 <- df %>% mutate(Class2= case_when(str_detect(Subclass, pattern = 'TN') ~ 'NLR (TNL)',
                                str_detect(Subclass, pattern = '^R') ~ Subclass,
                                TRUE ~ 'NLR (CNL)'))
df2 %>% ggplot(aes(fill=Class2, x=Genome, y=Count)) + geom_bar(stat='identity') +
    scale_fill_manual(values=pal)  + coord_flip() +  cowplot::theme_minimal_vgrid() +
    theme(axis.text.y = element_text(face = "italic")) 
```

```{r}
df2 %>% dplyr::filter(Genome %in% c('P. australis', 'Z. marina', 'Z. muelleri', 'O. sativa', 'A. antarctica', 'A. thaliana')) %>% 
  ggplot(aes(fill=Class2, x=factor(Genome, levels=c('Z. marina', 'Z. muelleri', 'P. australis', 'A. antarctica', 'O. sativa', 'A. thaliana' )), y=Count)) + 
  geom_bar(stat='identity') +
    scale_fill_manual(values=pal)  + coord_flip() +  cowplot::theme_minimal_vgrid() +
    theme(axis.text.y = element_text(face = "italic")) +
  xlab('Genome') +
  labs(fill='Class')
```

### 2021-10-07 docs/associateGenes.Rmd
Missing file

---
title: "Comparing gene loss with overall gene numbers"
author: "Philipp Bayer"
date: "14/09/2021"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

Is there an association between the lost number of *candidate* genes and the overall gene number? In other words, is the loss of our ~50 candidate genes driven by overall gene loss, or is it driven by adaptation to aquatic life?

If there's no association between overall gene numbers and the status of our ~50 genes then there might be selection pressure to lose these genes. If there's an association then these candidate genes may not be adaptation to aquatic life, but instead may be lost due to overall genome shrinkage pressure.

This is not currently part of the paper.

Data comes from Gene_content_Amphi_posi.xlsx

```{r}
data.frame(
  stringsAsFactors = FALSE,
                Species = c("A. antarctica","P. australis",
                       "Z. muelleri","Z. marina","L. gibba","S. polyrhiza",
                       "O. lucimarinus","C. reinhardtii","T. salsuginea",
                       "A. trichopada","B. distachyon","O. sativa","P. patens",
                       "P. trichocarpa","S. moellendorfii","A. thaliana",
                       "V. vinifera","Z. mays"),
                `Lost genes` = c(46L,30L,21L,27L,10L,23L,
                       0L,50L,10L,7L,0L,1L,24L,1L,24L,0L,5L,0L),
                Genes = c(16176L,21126L,35875L,21483L,
                       21830L,20661L,7742L,17741L,27132L,26846L,32439L,
                       42189L,32926L,34699L,22285L,35836L,31845L,39498L)
) -> df
```

```{r}
df
```
from getRandomGeneCounts.py

```{r}
random_lost <- data.frame(
  stringsAsFactors = FALSE,
                               Species = c("Z. muelleri","A. trichopada","O. sativa",
                                      "T. salsuginea","P. trichocarpa",
                                      "P. patens","Z. marina","S. polyrhiza",
                                      "C. reinhardtii","S. moellendorffii",
                                      "O. lucimarinus","Z. mays","P. australis",
                                      "L. gibba","V. vinifera","A. antarctica",
                                      "A. thaliana","B. distachyon"),
                               RandomLost = c(34L,34L,28L,36L,37L,40L,38L,40L,39L,
                                      39L,41L,33L,38L,39L,36L,35L,36L,
                                      33L)
               )

both <- left_join(df, random_lost)
bothlong <- both %>% pivot_longer( c('Lost.genes', 'RandomLost'), names_to = 'Type', values_to='Count')
```

```{r}
bothlong %>% ggplot(aes(y = Genes, x= Count, color=Type)) + 
  geom_point() + 
  cowplot::theme_minimal_grid() +
  xlab('Lost genes') 
```

```{r}
fit1 <- lm(Genes ~ Lost.genes, data=df)
summary(fit1)
```

```{r}
fit2 <- lm(Genes ~ RandomLost, data=both)
summary(fit2)
```

```{r}
ggplotRegression <- function (fit) {

require(ggplot2)

ggplot(fit$model, aes_string(x = names(fit$model)[2], y = names(fit$model)[1])) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red") +
  labs(title = paste("Adj R2 = ",signif(summary(fit)$adj.r.squared, 5),
                     " Slope =",signif(fit$coef[[2]], 5),
                     " P =",signif(summary(fit)$coef[2,4], 5))) +
  cowplot::theme_minimal_grid() +
  xlab('Lost genes')
}

ggplotRegression(fit1)
```

```{r}
fit2 <- lm(Genes ~ Lost.genes, data=df %>% filter(Genes > 10000))
summary(fit2)
```
```{r}
ggplotRegression(fit2)
```

```{r}
library(patchwork)

ggplotRegression(fit1) + ggplotRegression(fit2)
```

### 2021-10-07 analysis/GOenrichment.Rmd
Add missing data
Updated GOenrichment

@@
 # give properly formatted background in format: GO:0005838	GSBRNA2T00088508001;GSBRNA2T00088313001;GSBRNA2T00035842001 
-#annAT <- readMappings('BACKGROUND.txt', sep="\t", IDsep=";")
+#annAT <- readMappings('BACKGROUND.txt.gz', sep="\t", IDsep=";")
 #save(annAT, file='annAtObject.RData')

### 2021-10-18 analysis/GOenrichment.Rmd
wflow_publish(c("analysis/GOenrichment.Rmd", "analysis/plotGenes.Rmd",     "analysis/plotRgenes.Rmd"))

@@
 library(UpSetR)
+library(rrvgo)
 knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
@@
   one.data <- dat
-
+  if (one.data[1,] == '<html>') {next} # some datasets result in error
   one.data <- one.data [(one.data$plot_X != "null" & one.data$plot_Y != "null") & (! is.na(one.data$frequency)) & (one.data$value != 'null'), ];
   one.data <- type_convert(one.data)
-  one.data$pvalue <- p.adjust(exp(one.data$value))
-  #one.data <- one.data %>% filter(pvalue < 0.05) # NO FILTERING OF P-VALUES - for now?
 
@@
     ggplot(data = one.data, aes(plot_X, plot_Y, label = description)) +
-    geom_point(aes(colour = log(pvalue), size = log_size), alpha = I(0.6)) +
+    geom_point(aes(colour = value, size = log_size), alpha = I(0.6)) +
     scale_size_area() +
     scale_colour_gradientn(colours = rev(wes_palette("Zissou1", 100, type = "continuous")), 
-                           limits = c(min(log(one.data$pvalue)), 0)) +
+                           limits = c(min(one.data$value), 0)) +
     geom_point(

### 2021-10-27 analysis/Orthofinder.Rmd
Add missing files

@@
 library(UpSetR)
+knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
 
@@
 ```
+
+# Connect Orthofinder results with functional table of Arabidopsis genes
+
+File comes from https://www.arabidopsis.org/download_files/Genes/TAIR10_genome_release/TAIR10_functional_descriptions
+
+```{r eval=FALSE}
+download.file('https://www.arabidopsis.org/download_files/Genes/TAIR10_genome_release/TAIR10_functional_descriptions', 'data/TAIR10_functional_descriptions')
+```
+```{r}
+functions <- read_tsv('data/TAIR10_functional_descriptions')
+```
+I also downloaded the gene symbols:
+
+```{r eval=FALSE}
+esearch -db gene -query "Arabidopsis thaliana [ORGN]" | esummary | xtract -pattern DocumentSummary -element Name,OtherAliases |  awk -F "\t|," '{OFS="\t"}{print $2,$1}' > arabidopsis_gene_symbols.txt
+```
+
+```{r}
+symbols <- read_tsv('./data/arabidopsis_gene_symbols.txt', col_names = c('Gene','Symbol'))
+```
+
+OK now I need to flip the Orthofinder table to then join the functional table
+
+```{r}
+ara_groups <- groups %>% select(Orthogroup, `A. thaliana`) %>% filter(!is.na(`A. thaliana`)) %>% separate_rows(Orthogroup, `A. thaliana`, convert=T)
+ara_groups <- ara_groups %>% separate(`A. thaliana`, c('Gene','Number'), remove=FALSE)
+ara_groups <- left_join(ara_groups, symbols)
+ara_groups$Number <- NULL
+head(ara_groups)
+```
+
+```{r}
+ara_joined <- left_join(ara_groups, functions, by=c(`A. thaliana`='Model_name'))
+```
+
+You know what? we shouldn't directly compare a dicot (A. thaliana) with monocots (seagrasses), should also add another regular monocot (rice)
+
+```{r}
+posi_groups <- groups %>% select(Orthogroup, `P. australis`) %>% filter(!is.na(`P. australis`)) %>% separate_rows(Orthogroup, `P. australis`, convert=T, sep = ', ')
+amphi_groups <- groups %>% select(Orthogroup, `A. antarctica`) %>% filter(!is.na(`A. antarctica`)) %>% separate_rows(Orthogroup, `A. antarctica`, convert=T, sep = ', ')
+zmar_groups <- groups %>%  select(Orthogroup, `Z. marina`) %>% filter(!is.na(`Z. marina`)) %>% separate_rows(Orthogroup, `Z. marina`, convert=T, sep = ', ')
+zmuel_groups <- groups %>%  select(Orthogroup, `Z. muelleri`) %>% filter(!is.na(`Z. muelleri`)) %>% separate_rows(Orthogroup, `Z. muelleri`, convert=T, sep = ', ')
+w_australis_groups <- groups %>%  select(Orthogroup, `W. australis`) %>% filter(!is.na(`W. australis`)) %>% separate_rows(Orthogroup, `W. australis`, convert=T, sep = ', ')
+s_polyrhiza_groups <- groups %>%  select(Orthogroup, `S. polyrhiza`) %>% filter(!is.na(`S. polyrhiza`)) %>% separate_rows(Orthogroup, `S. polyrhiza`, convert=T, sep = ', ')
+l_gibba_groups <- groups %>%  select(Orthogroup, `L. gibba`) %>% filter(!is.na(`L. gibba`)) %>% separate_rows(Orthogroup, `L. gibba`, convert=T, sep = ', ')
+rice_groups<- groups %>%  select(Orthogroup, `O. sativa`) %>% filter(!is.na(`O. sativa`)) %>% separate_rows(Orthogroup, `O. sativa`, convert=T, sep = ', ')
+```
+
+```{r}
+big_joined <- ara_joined %>% mutate(
+  present_in_zmuel = case_when(Orthogroup %in% unique(zmuel_groups$Orthogroup) ~ T,
+                               TRUE ~ F),
+  present_in_zmar = case_when(Orthogroup %in% unique(zmar_groups$Orthogroup) ~ T,
+                              TRUE ~ F),
+  present_in_amphi = case_when(Orthogroup %in% unique(amphi_groups$Orthogroup) ~ T,
+                               TRUE ~ F),
+  present_in_posi = case_when(Orthogroup %in% unique(posi_groups$Orthogroup) ~ T,
+                              TRUE ~ F),
+  present_in_waustralis = case_when(Orthogroup %in% unique(w_australis_groups$Orthogroup) ~ T,
+                              TRUE ~ F),
+  present_in_spolyrhiza = case_when(Orthogroup %in% unique(s_polyrhiza_groups$Orthogroup) ~ T,
+                              TRUE ~ F),
+  present_in_lgibba = case_when(Orthogroup %in% unique(l_gibba_groups$Orthogroup) ~ T,
+                              TRUE ~ F),
+  present_in_rice = case_when(Orthogroup %in% unique(rice_groups$Orthogroup) ~ T,
+                              TRUE ~ F)
+  )
+```
+
+Done :) Now we have a big table with all *A. thaliana* gene functions and whether these genes are present in clusters of the four seagrasses and three duckweeds.
+
+Let's also add counts - how often is this gene lost?
+```{r}
+big_joined <- big_joined %>% mutate(Lost_in_Seagrasses = 4 - (present_in_posi + present_in_zmuel + present_in_zmar + present_in_amphi),
+                      Lost_in_duckweeds=3 - (present_in_lgibba + present_in_spolyrhiza + present_in_waustralis),
+                      Lost_in_both = (Lost_in_Seagrasses + Lost_in_duckweeds))
+```
+
+```{r}
+big_joined %>% writexl::write_xlsx('data/arabidopsis_gene_level_comparison.xlsx')
+big_joined %>% filter_at(vars(starts_with('present')), any_vars(. == FALSE)) %>% writexl::write_xlsx('data/arabidopsis_gene_level_comparison_only_losts.xlsx')
+```
+
+That was about loss - can we also check for gene family extension using the *A. thaliana* genes?
+
+# Gene family extension
+
+First, let's count how many members each of these orthogroups has per species.
+```{r}
+ara_count <- ara_joined %>% count(Orthogroup)
+posi_count <- posi_groups %>% count(Orthogroup)
+amphi_count <- amphi_groups %>% count(Orthogroup)
+zmar_count <- zmar_groups %>% count(Orthogroup)
+zmuel_count <- zmuel_groups %>% count(Orthogroup)
+lgibba_count <- l_gibba_groups %>% count(Orthogroup)
+s_polyrhi_count <- s_polyrhiza_groups %>% count(Orthogroup)
+w_aus_count <- w_australis_groups %>% count(Orthogroup)
+rice_count <- rice_groups %>% count(Orthogroup)
+```
+
+```{r}
+counts <- plyr::join_all(list(ara_count, posi_count, amphi_count, zmar_count, zmuel_count, lgibba_count, s_polyrhi_count, w_aus_count, rice_count), by='Orthogroup', type='left') 
+names(counts) <- c('Orthogroup', 'A. thaliana', 'P. australis', 'A. antarctica', 'Z. marina', 'Z. muelleri', 'L. gibba', 'S. polyrhiza', 'W. australis', 'O. sativa')
+counts
+```
+
+```{r}
+joined_counts <- left_join(big_joined, counts, by='Orthogroup') %>% select(-starts_with('present')) %>% 
+  select(-starts_with('Lost_i'))
+joined_counts %>% writexl::write_xlsx('data/arabidopsis_gene_level_counts.xlsx')
+```

### 2021-12-20 analysis/GOenrichment.Rmd
add tables

@@
   one.data <- type_convert(one.data)
-
+  top_ten_values <- one.data %>% arrange(value) %>% head(10) %>% pull(term_ID) 
+  one.data <- one.data %>% mutate(description2 = case_when(
+    term_ID %in% top_ten_values ~ description, 
+                                                           TRUE ~ ''))
   p1 <-
-    ggplot(data = one.data, aes(plot_X, plot_Y, label = description)) +
+    ggplot(data = one.data, aes(plot_X, plot_Y, label = description2)) +
     geom_point(aes(colour = value, size = log_size), alpha = I(0.6)) +
@@
     scale_size_area() +
-    geom_label_repel(max.overlaps = 5, box.padding = 0.5,
+    geom_label_repel(max.overlaps = Inf, box.padding = 0.5,
                      aes(point.size = log_size)) +
@@
     ) +
-    ggtitle(f)
+    NULL
+    #ggtitle(f)
 
@@
 }
+```
+
+# Making Supplementary Tables for the REVIGO results
+
+Let's also pull these terms out as tables
+
+```{r}
+for( i in names(results_list)) {
+  print(i)
+  knitr::kable(results_list[[i]] %>% filter(Eliminated == FALSE), label=i)
+}
+
 ```
@@
 ```{r}
-plot_list
+for(i in names(plot_list)) {
+  file_name = paste('output/',  i, '.png', sep='')
+
+  cowplot::save_plot(file_name, plot_list[[i]], base_height=8)
+
+}
 ```

### 2021-12-20 analysis/GOenrichment.Rmd
add tables
fix tables
fix tables again
fix tables again, finally :)

@@
 
-Let's also pull these terms out as tables
+Let's also pull these terms out as tables. `Value` is the unadjusted p-value.
 
-```{r}
+```{r results='asis'}
 for( i in names(results_list)) {
   print(i)
-  knitr::kable(results_list[[i]] %>% filter(Eliminated == FALSE), label=i)
+  print(knitr::kable(results_list[[i]] %>% filter(Eliminated == FALSE) %>% select(TermID, Name, Value), label=i))
+  cat("\n")
 }

### 2021-12-20 analysis/GOenrichment.Rmd
fix tables again, finally :) haha

--- 2021-12-20 analysis/GOenrichment.Rmd
@@
 
-Let's also pull these terms out as tables
+Let's also pull these terms out as tables. `Value` is the unadjusted p-value (-3.7 = 10 * -3.7)
 
-```{r}
+```{r results='asis'}
 for( i in names(results_list)) {
   print(i)
-  knitr::kable(results_list[[i]] %>% filter(Eliminated == FALSE), label=i)
+  print(knitr::kable(results_list[[i]] %>% filter(Eliminated == FALSE) %>% select(TermID, Name, Value) %>% arrange(Value), caption=i))
+  cat("\n")
 }

### 2021-12-20 data/Chloroplast_assemblies/compareArabidopsisgenes.py
Add missing files

@@ -1,0 +1,38 @@
+ara_genes = {}
+for line in open('NC_000932.1.gff3'):
+    ll = line.split()
+    if len(ll) < 3: continue
+    thistype = ll[2]
+    if thistype != 'gene':
+        continue
+    # ['NC_000932.1', 'Chloe', 'gene', '14024', '14770', '.', '-', '.', 'ID=atpI;Name=atpI']
+    start = ll[3]
+    end = ll[4]
+    length = abs(int(end) - int(start))
+    name = ll[-1].split(';')[0].replace('ID=','')
+    ara_genes[name] = length
+
+
+from glob import glob
+from collections import defaultdict
+for l in glob('*gff3'):
+    if 'NC' in l: continue
+    thisgenes = defaultdict(int)
+    print('----')
+    print(l)
+    for line in open(l):
+        ll = line.split()
+        if len(ll) < 3: continue
+        thistype = ll[2]
+        if thistype != 'gene':
+            continue
+        # ['NC_000932.1', 'Chloe', 'gene', '14024', '14770', '.', '-', '.', 'ID=atpI;Name=atpI']
+        start = ll[3]
+        end = ll[4]
+        length = abs(int(end) - int(start))
+        name = ll[-1].split(';')[0].replace('ID=','')
+        thisgenes[name] = length
+
+    for ara_gene in sorted(ara_genes.keys()):
+        perc_diff = abs(ara_genes[ara_gene] - thisgenes[ara_gene]) / ara_genes[ara_gene] * 100
+        print(ara_gene, ara_genes[ara_gene], thisgenes[ara_gene], '%.1f'%perc_diff)

### 2022-01-16 analysis/Orthofinder.Rmd
Add missing files
better analysis

@@
 ---
+
+Here I use Orthofinder results to pull out genes unique to seagrasses, unique to duckweeds, and make some nice summary tables
 
@@
 ```
+
+# FINAL version redoing - terrestrial vs aquatics vs seagrasses
+
+Plotting all species separately is messy as I have so many species. I'm joining the species into three groups: aquatics, terrestrials, and seagrasses!
+
+```{r}
+per_group_spec <-
+  per_spec %>% mutate(group = case_when(
+    name %in% c(
+      'A. trichopada',
+      'B. distachyon',
+      'O. sativa',
+      'P. patens',
+      'P. trichocarpa',
+      'S. moellendorffii',
+      'A. thaliana',
+      'T. parvula',
+      'V. vinifera',
+      'Z. mays'
+    ) ~ 'Terrestrial',
+    name %in% c(
+      'C. reinhardtii',
+      'L. gibba',
+      'S. polyrhiza',
+      'O. lucimarinus',
+      'W. australis'
+    ) ~ 'Aquatic',
+    name %in% c('P. australis', 'Z. muelleri', 'Z. marina', 'A. antarctica') ~ 'Seagrasses'
+  )) 
+```
+
+```{r}
+# now I want the data in this format:
+# listInput <- list(one = c(1, 2, 3, 5, 7, 8, 11, 12, 13), two = c(1, 2, 4, 5, 
+#   10), three = c(1, 5, 6, 7, 8, 9, 10, 12, 13))
+groupx <- per_group_spec %>% 
+  select(group, Orthogroup) %>% # turn the table around
+  deframe() # convert to named vector
+```
+
+```{r}
+mylistgroup <- lapply(split(groupx, names(groupx)), unname) # yuck - ugly code to convert the named vector to a list
+```
+
+```{r}
+xgroup <- upset(fromList(mylistgroup), order.by='freq', nsets = length(groups) - 1)
+```
+
+```{r fig.width=10}
+xgroup
+```
+
+MUCH better.
+The biggest chunk of orthogroups is not in aquatics or seagrasses with 13562 orthogroups.
+3437 orthogroups are unique to aquatics, but not in seagrasses - 1183 orthogroups are unique in seagrasses, but not in aquatics.
+404 + 3437 + 2217 orthogroups are not in terrestrials.

### 2022-02-21 analysis/GeneLossComparisonReads.Rmd
New analysis yay

---
title: "Visualising gene differences based on read alignments"
author: "Philipp Bayer"
date: "21 February 2022"
output: html_document
---

Here I compare read alignment differences with known Arabidopsis genes.

```{r setup}
library(tidyverse)
library(cowplot)
theme_set(theme_cowplot())
library(RColorBrewer)
library(wesanderson)
library(gghighlight)
library(patchwork)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

```

The data is based on Supplementary Table S6 and S7.

# Stomata genes

```{r}
stomata <- tibble::tribble(
             ~Gene.symbol, ~A..antarctica, ~P..australis, ~H..ovalis, ~Z..muelleri, ~Z..marina,
                 "SBT1.2",              0,             0,                  5.63,          8.2,       4.34,
                  "EPFL9",              0,             0,                     0,            0,          0,
                   "EPF1",              0,         25.08,                     0,            0,          0,
                    "TMM",           3.35,             0,                     0,            0,          0,
                   "EPF2",              0,             0,                     0,            0,          0,
                  "MYB88",              0,             0,                  6.74,         2.34,          0,
                   "MUTE",              0,             0,                     0,            0,          0,
                   "SPCH",           7.95,          3.29,                  5.84,         4.47,          0,
                   "FAMA",           6.99,         20.64,                  6.83,         4.98,       4.98,
                  "SCRM2",           21.8,         25.57,                 19.66,        26.16,      22.25,
                    "FLP",            3.2,          8.77,                  4.42,            0,          0
             )

```

```{r}
stomata_long <- stomata %>% pivot_longer(-Gene.symbol, values_to = 'Coverage', names_to='Species') %>% 
  mutate(Species = gsub('\\.\\.', '. ', Species))
```

```{r}
stom_p <- stomata_long %>% ggplot(aes(x=forcats::fct_rev(Species), y=Coverage, fill=Species, color=Species)) + 
  geom_boxplot(
    aes(fill = after_scale(colorspace::lighten(fill, .9))),
    size = 1.5, outlier.shape = NA
  ) +
  geom_jitter(width = .1, size = 4, alpha = .5) +
  ylim(c(-1,100)) +

  theme(legend.position = "None",
  axis.text.y = element_text(face = "italic")) +
  scale_color_brewer(palette='Dark2') +
  ylab('Coverage (%)') +
  coord_flip() + xlab('Species')

stom_p
```

# Ethylene

```{r}
ethylene <- tibble::tribble(
              ~Gene.symbol, ~A..antarctica, ~P..australis, ~H..ovalis, ~Z..muelleri, ~Z..marina,
                    "ACO1",              0,         53.16,          0,            0,          0,
                    "ACO2",              0,         11.53,          0,            0,          0,
                    "ACO4",              0,         15.74,          0,            0,          0,
                    "ACO5",              0,             0,          0,         6.39,        5.3,
                    "ACS1",           9.34,          5.11,          0,            0,          0,
                    "ACS2",           7.24,          5.97,          0,         4.43,          0,
                    "ACS4",          18.74,         25.19,          0,         8.49,       8.21,
                    "ACS5",          19.25,         50.39,       4.46,            0,          0,
                    "ACS6",              0,             0,          0,            0,          0,
                    "ACS7",          10.27,         66.29,          0,         6.47,          0,
                    "ACS8",          38.58,         42.62,       6.24,            0,          0,
                    "ACS9",          20.52,         36.94,          0,            0,          0,
                   "ACS11",              0,         22.05,       7.45,            0,          0,
                    "ERS1",          35.23,         40.88,          0,          2.5,          0,
               "ETR1/EIN1",          49.57,         51.87,          0,         3.83,          0,
                    "ETR2",              0,         21.96,          0,            0,          0,
                    "EIN4",           1.74,         31.33,          0,            0,       1.78,
                    "CTR1",          26.68,         31.55,       5.96,            0,          0,
                    "EIN2",              0,         14.54,      11.12,            0,          0,
                    "EBF1",              0,             0,        4.5,            0,          0,
                    "EBF2",              0,         12.45,       4.22,            0,          0,
                    "EIN3",          27.03,         43.08,      36.83,        31.27,      31.11,
                    "MYC2",           4.43,         27.83,      10.31,        23.61,       23.5,
                    "MYC3",             19,         17.14,      10.06,        12.14,      13.77,
                    "MYC4",              0,          3.84,       8.19,        20.96,      11.75
              )

```

```{r}
ethylene_long <- ethylene %>% pivot_longer(-Gene.symbol, values_to = 'Coverage', names_to='Species') %>% 
  mutate(Species = gsub('\\.\\.', '. ', Species))
```

```{r}
eth_p <- ethylene_long %>% ggplot(aes(x=forcats::fct_rev(Species), y=Coverage, fill=Species, color=Species)) + 
  geom_boxplot(
    aes(fill = after_scale(colorspace::lighten(fill, .9))),
    size = 1.5, outlier.shape = NA
  ) +
  geom_jitter(width = .1, size = 4, alpha = .5) +
  ylim(c(-1,100)) +

  theme(legend.position = "None",
  axis.text.y = element_text(face = "italic")) +
  scale_color_brewer(palette='Dark2') +
  ylab('Coverage (%)') +
  coord_flip() +
  xlab('Species')
eth_p
```
# NDH genes

```{r}
ndh <- tibble::tribble(
         ~Gene.symbol, ~A..antarctica, ~P..australis, ~H..ovalis, ~Z..muelleri, ~Z..marina,
               "NDHL",              0,             0,          0,        16.67,      16.67,
               "NDHM",              0,             0,          0,        36.24,       34.4,
               "NDHN",          48.73,             0,          0,        66.35,      59.68,
               "NDHO",              0,             0,          0,        33.54,      22.22,
               "NDHS",          12.35,         28.29,          0,         34.4,      30.15,
               "NDHT",              0,          8.67,          0,        19.47,      16.67,
               "NDHU",              0,             0,          0,         9.89,          0,
              "PNSB1",           4.26,         42.42,          0,         32.4,      31.24,
              "PNSB2",              0,             0,          0,            0,          0,
              "PNSB3",              0,          8.46,          0,        16.26,       9.76,
              "PNSB4",              0,         19.13,          0,        18.75,      18.75,
              "PNSL1",              0,             0,          0,            0,          0,
              "PNSL2",          16.58,             0,          0,        45.03,      15.01,
              "PNSL3",              0,             0,          0,            0,          0,
              "PNSL4",           7.19,         23.55,          0,        31.04,      45.11,
              "PNSL5",          52.82,         62.82,      57.31,        48.72,      46.28,
               "CRR6",              0,             0,          0,        43.86,      33.06,
               "CRR7",              0,             0,          0,            0,          0,
              "Lhca5",              0,         35.15,          0,        18.29,      27.76,
              "Lhca6",              0,         47.85,          0,        58.43,      60.27,
              "CRR27",           21.9,           9.2,       9.26,        36.38,      41.78,
              "CRR41",              0,             0,          0,        44.03,       7.86,
               "PGR5",          55.97,         56.72,      53.98,        55.22,      55.22,
             "PGRL1A",          17.54,         23.79,      29.54,        33.54,      31.08,
               "CRR2",          41.49,          61.4,       2.94,        23.51,      13.93,
               "CRR3",              0,             0,          0,            0,          0,
              "CRR42",              0,             0,          0,        30.89,      28.44,
               "PQL3",              0,             0,          0,            0,          0,
               "NDF5",              0,             0,          0,            0,          0
         )

```

```{r}
ndh_long <- ndh %>% pivot_longer(-Gene.symbol, values_to = 'Coverage', names_to='Species') %>% 
  mutate(Species = gsub('\\.\\.', '. ', Species))
```

```{r}
ndh_p <- ndh_long %>% ggplot(aes(x=forcats::fct_rev(Species), y=Coverage, fill=Species, color=Species)) + 
  geom_boxplot(
    aes( fill = after_scale(colorspace::lighten(fill, .9))),
    size = 1.5, outlier.shape = NA
  ) +
  geom_jitter(width = .1, size = 4, alpha = .5) +
  ylim(c(-1,100)) +

  theme(legend.position = "None",
  axis.text.y = element_text(face = "italic")) +
  scale_color_brewer(palette='Dark2') +
  ylab('Coverage %') +
  coord_flip() +
  xlab('Species')
ndh_p
```

The following data is slightly different because it compares lengths of annotated genes in the chloroplast assembly.

```{r}
ndh_chloro <- tibble::tribble(
  ~Gene.symbol, ~A..antarctica, ~H..ovalis, ~P..australis, ~Z..marina, ~Z..muelleri,
        "ndhA",              0,          0,             0,      101.8,        101.4,
        "ndhB",              0,          0,             0,      100.4,        100.9,
      "ndhB-2",              0,          0,             0,      100.4,        100.9,
        "ndhC",           92.5,          0,           100,        100,          100,
        "ndhD",           29.4,          0,             0,        100,          100,
        "ndhE",              0,          0,          45.4,        100,          100,
        "ndhF",              0,          0,             0,       98.4,         98.3,
        "ndhG",           45.9,          0,             0,        100,          100,
        "ndhH",             34,       63.1,          63.8,        100,          100,
        "ndhI",              0,          0,             0,      110.5,        110.5,
        "ndhJ",           70.8,          0,             0,        100,          100,
        "ndhK",             55,          0,         109.3,      109.8,        109.8
  )

```

```{r}
ndh_chl_long <- ndh_chloro %>% pivot_longer(-Gene.symbol, values_to = 'Coverage', names_to='Species') %>% 
  mutate(Species = gsub('\\.\\.', '. ', Species))
```

Stylistic choice - if the new gene is longer than the Arabidopsis one, set to 100%.

```{r}
ndh_chl_p <- ndh_chl_long %>%
  mutate(Coverage = case_when(Coverage > 100 ~ 100,
                              TRUE ~ Coverage)) %>% 
  ggplot(aes(x=forcats::fct_rev(Species), y=Coverage, fill=Species, color=Species)) + 
  geom_boxplot(
    aes( fill = after_scale(colorspace::lighten(fill, .9))),
    size = 1.5, outlier.shape = NA
  ) +
  geom_jitter(width = .1, size = 4, alpha = .5) +
  ylim(c(-1,100.1)) +

  theme(legend.position = "None",
  axis.text.y = element_text(face = "italic")) +
  scale_color_brewer(palette='Dark2') +
  ylab('Gene size difference (%)') +
  xlab('Species') +
  coord_flip()
ndh_chl_p
```

# all together

```{r fig.width=10}
patchwork <- (eth_p + stom_p) / (ndh_p + ndh_chl_p)
# patchwork[[2]] = patchwork[[2]] + theme(axis.text.y = element_blank(),
#                                         axis.ticks.y = element_blank(),
#                                         axis.title.y = element_blank() )
# patchwork[[1]] = patchwork[[1]] + theme(axis.text.y = element_blank(),
#                                         axis.ticks.y = element_blank(),
#                                         axis.title.y = element_blank() )
patchwork+ plot_annotation(tag_levels = "A")
```

### 2022-02-21 analysis/Orthofinder.Rmd
add some more changes

@@
 library(RColorBrewer)
+library(wesanderson)
 library(patchwork)
 library(UpSetR)
+library(ggVennDiagram)
 knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
@@
 names(counts) <- c('Orthogroup', 'A. thaliana', 'P. australis', 'A. antarctica', 'Z. marina', 'Z. muelleri', 'L. gibba', 'S. polyrhiza', 'W. australis', 'O. sativa')
-counts
+head(counts)
 ```
@@
       'Z. mays'
-    ) ~ 'Terrestrial',
+    ) ~ 'Terrestrials',
     name %in% c(
@@
       'W. australis'
-    ) ~ 'Aquatic',
+    ) ~ 'Aquatics',
     name %in% c('P. australis', 'Z. muelleri', 'Z. marina', 'A. antarctica') ~ 'Seagrasses'
@@
 mylistgroup <- lapply(split(groupx, names(groupx)), unname) # yuck - ugly code to convert the named vector to a list
+mylistgroup <- lapply(mylistgroup, unique)
 ```
@@
 404 + 3437 + 2217 orthogroups are not in terrestrials.
+
+OK now we need to do the GOenrichment for those three groups.
+
+Let's also make an Venn plot, we have only three groups:
+```{r}
+p <- ggVennDiagram(mylistgroup, label_alpha=0) + 
+  scale_fill_gradientn(colours=wes_palette("Zissou1", 100, type = "continuous")) + 
+  theme(legend.position = "none")
+```
+
+```{r fig.width=10}
+p
+```
+
+Nice!

### 2022-02-21 analysis/plotRgenes.Rmd
Add missing data
better plot

@@
 library(wesanderson)
+library(ggtree)
+library(ggtreeExtra)
+knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
+
 ```
 
-Visualising R-gene differences here
+# Visualising R-gene differences here
 
@@
 ```
+
+Let's link those to the phylogeny we got from timetree.org
+
+```{r}
+tree <- ape::read.tree('./data/timetree_species.nwk')
+tree$tip.label <- c('O.
+lucimarinus', 'C.
+reinhardtii', 'P.
+patens', 'S.
+moellendorffii', 'O.
+sativa', 'B.
+distachyon', 'Z.
+mays', 'P.
+australis', 'Z.
+marina', 'Z.
+muelleri', 'A.
+antarctica', 'S.
+polyrhiza', 'L.
+gibba', 'V.
+vinifera', 'A.
+thaliana', 'S.
+parvula', 'P.
+trichocarpa', 'A.
+trichopada')
+
+```
+```{r}
+p2 <- df2 %>% ggplot(aes(fill=Class2, x=Genome, y=Count)) + geom_bar(stat='identity') +
+    scale_fill_manual(values=pal)  + coord_flip() +  cowplot::theme_minimal_vgrid() +
+    theme(axis.text.y = element_text(face = "italic")) 
+```
+
+```{r}
+p1 <- ggtree(tree)
+p1
+```
+
+```{r}
+df2$label <- df2$Genome
+# get species not in tree
+subtree <- ape::drop.tip(tree, tree$tip.label[!tree$tip.label %in% df2$label])
+```
+
+```{r}
+p1 <- ggtree(subtree)
+p1 
+df3 <- as.data.frame(df2)
+df3$label <- df3$Genome
+df3$id <- df3$label
+
+```
+
+```{r}
+# code from https://thackl.github.io/ggtree-composite-plots
+
+tree_y <-  function(ggtree, data){
+  if(!inherits(ggtree, "ggtree"))
+    stop("not a ggtree object")
+  left_join(select(data, label), select(ggtree$data, label, y)) %>%
+    pull(y)
+}
+# overwrite the default expand for continuous scales
+scale_y_tree <- function(expand=expand_scale(0, 0.6), ...){
+    scale_y_continuous(expand=expand, ...)
+}
+
+# get the range of the ggtree y-axis data
+tree_ylim <- function(ggtree){
+  if(!inherits(ggtree, "ggtree"))
+    stop("not a ggtree object")
+  range(ggtree$data$y)
+}
+
+# plot data next to a ggtree aligned by shared labels
+ggtreeplot <- function(ggtree, data = NULL, mapping = aes(), flip=FALSE,
+     expand_limits=expand_scale(0,.6), ...){
+
+  if(!inherits(ggtree, "ggtree"))
+    stop("not a ggtree object")
+
+  # match the tree limits
+  limits <- tree_ylim(ggtree)
+  limits[1] <- limits[1] + (limits[1] * expand_limits[1]) - expand_limits[2]
+  limits[2] <- limits[2] + (limits[2] * expand_limits[3]) + expand_limits[4]
+
+  if(flip){
+    mapping <- modifyList(aes_(x=~x), mapping)
+    data <- mutate(data, x=tree_y(ggtree, data))
+    gg <- ggplot(data=data, mapping = mapping, ...) +
+      scale_x_continuous(limits=limits, expand=c(0,0))
+  }else{
+    mapping <- modifyList(aes_(y=~y), mapping)
+    data <- mutate(data, y=tree_y(ggtree, data))
+    gg <- ggplot(data=data, mapping = mapping, ...) +
+      scale_y_continuous(limits=limits, expand=c(0,0))
+  }
+  gg
+}
+
+# get rid of superfluous axis - this works after coord_flip, so it also works
+# for the rotated histogram
+no_y_axis <- function () 
+  theme(axis.line.y = element_blank(), 
+        axis.title.y = element_blank(),
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank())
+```
+
+```{r}
+p3 <-  ggtree(subtree) + geom_tiplab(align=TRUE, fontface='italic') +
+  scale_x_continuous(expand=expand_scale(0.8)) + scale_y_tree()
+
+myhist <- ggtreeplot(p3, df3, aes(y=Count, color=Class2, fill=Class2), flip=TRUE) +
+  geom_col(aes(fill=Class2,group=Class2,color=Class2)) + 
+  #theme(legend.position="none") +
+  coord_flip() + no_y_axis()  + 
+  theme(legend.position=c(0.6, 0.87)) + 
+  labs(fill='Class', color='Class')
+
+p3 + myhist
+
+```

### 2022-02-21 analysis/plotRgenes.Rmd
summaryTable

@@
 library(ggtreeExtra)
+library(kableExtra)
 knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
@@
 ```
+
+# Quick summary stats
+
+```{r}
+df %>% group_by(Genome, Class) %>% summarise(sum=sum(Count)) %>% kbl() %>%
+  kable_styling()
+```

### 2022-02-22 analysis/GeneLossComparisonReads.Rmd
add some more changes
Make order of plot nicer

@@
   scale_color_brewer(palette='Dark2') +
-  ylab('Gene size difference (%)') +
+  ylab('Gene size similarity (%)') +
   xlab('Species') +
@@
 ```{r fig.width=10}
-patchwork <- (eth_p + stom_p) / (ndh_p + ndh_chl_p)
+patchwork <- (eth_p + stom_p) / ( ndh_chl_p + ndh_p )
 # patchwork[[2]] = patchwork[[2]] + theme(axis.text.y = element_blank(),

### 2022-02-22 analysis/Orthofinder.Rmd
venn diagram
Add changes

@@
 library(UpSetR)
+library(kableExtra)
 library(ggVennDiagram)
@@
 ```{r}
-groups <- read_tsv('./data/Orthogroups.tsv')
+groups <- read_tsv('./data/Orthogroups.tsv.gz')
 
@@
   arrange(n) %>% 
-  knitr::kable()
+  kbl() %>%
+  kable_styling()
 ```
@@
 ```
+
 ```{r}
@@
 
-```{r fig.width=10}
+```{r fig.width=9}
 p
@@
 Nice!
+
+# Write out the per-group GO-terms
+
+Now we pull out all gene-IDs that are only in terrestrials, only in seagrasses, only in aquatics, and unique for each of these three groups, AND we pull out the IDs only present in the four seagrasses each compared with the others.
+
+```{r}
+gene_per_spec <- groups %>% pivot_longer(-Orthogroup) %>% 
+  filter(!is.na(value)) %>% 
+  mutate(group = case_when(
+    name %in% c(
+      'A. trichopada',
+      'B. distachyon',
+      'O. sativa',
+      'P. patens',
+      'P. trichocarpa',
+      'S. moellendorffii',
+      'A. thaliana',
+      'T. parvula',
+      'V. vinifera',
+      'Z. mays'
+    ) ~ 'Terrestrials',
+    name %in% c(
+      'C. reinhardtii',
+      'L. gibba',
+      'S. polyrhiza',
+      'O. lucimarinus',
+      'W. australis'
+    ) ~ 'Aquatics',
+    name %in% c('P. australis', 'Z. muelleri', 'Z. marina', 'A. antarctica') ~ 'Seagrasses'
+  )) 
+```

### 2022-02-22 data/Chloroplast_assemblies/compareArabidopsisgenes.py

@@ -30,7 +30,7 @@ for l in glob('*gff3'):
         start = ll[3]
         end = ll[4]
         length = abs(int(end) - int(start))
-        name = ll[-1].split(';')[0].replace('ID=','')
+        name = ll[-1].split(';')[1].replace('Name=','')
         thisgenes[name] = length
 
     for ara_gene in sorted(ara_genes.keys()):

### 2022-02-23 analysis/Orthofinder.Rmd
Rebuild Orthofinder analysis

@@
 Let's also make an Venn plot, we have only three groups:
-```{r}
-p <- ggVennDiagram(mylistgroup, label_alpha=0) + 
+```{r fig.width=8}
+
+# code from https://github.com/gaospecial/ggVennDiagram/blob/4cb2aa13c7beae469f9b9836c5d4f94610bd872e/R/ggVennDiagram.R to customise
+venn <- Venn(mylistgroup)
+data <- process_data(venn)
+region_label <- data@region %>%
+      dplyr::filter(.data$component == "region") %>%
+      dplyr::mutate(percent = paste(round(.data$count*100/sum(.data$count),
+                                          digits = 0),"%", sep="")) %>%
+      dplyr::mutate(both = paste(.data$count,paste0("(",.data$percent,")"),sep = "\n"))
+
+group_venn <- ggplot() +
+    geom_sf(aes_string(fill="count"), data = data@region) +
+    geom_sf(aes_string(color = "id"), data = data@setEdge, show.legend = F,
+            lty = 'solid', size = 1, color='gray') +
+    geom_sf_text(aes_string(label = "name"), data = data@setLabel,
+                 size = NA, 
+                 color = 'black') +
+    theme_void() +
+     geom_sf_label(aes_string(label='both'),
+                             data = region_label,
+                             alpha= 0.0,
+                             color = 'black',
+                             size = NA,
+                             lineheight = 0.85,
+                             label.size = NA) +
   scale_fill_gradientn(colours=wes_palette("Zissou1", 100, type = "continuous")) + 
-  theme(legend.position = "none")
-```
+  theme(legend.position = "none") + 
+  scale_x_continuous(expand = expansion(mult = .2)) #trick from https://github.com/gaospecial/ggVennDiagram/blob/9435aa0ab4abb470c670ecb938c71576461ccedc/vignettes/using-ggVennDiagram.Rmd#L65
 
-```{r fig.width=9}
-p
+group_venn
 ```
@@
 ```
+
+mylistgroup now looks like this:
+key is Aquatics, Seagrasses, Terrestrials, values are OG00000, OG00001, OG00002 etc.
+
+```{r}
+present_aquatics_only <- setdiff(setdiff(mylistgroup$Aquatics, mylistgroup$Terrestrials), 
+                         mylistgroup$Seagrasses)
+present_seagrasses_only <- setdiff(setdiff(mylistgroup$Seagrasses, mylistgroup$Aquatics), 
+                           mylistgroup$Terrestrials)
+present_terrestrials_only <- setdiff(setdiff(mylistgroup$Terrestrials, mylistgroup$Aquatics),
+                             mylistgroup$Seagrasses)
+print(cbind(length(present_aquatics_only), length(present_seagrasses_only), length(present_terrestrials_only)))
+
+```
+OK those numbers fit with the above Venn diagram. I now have three lists of orthogroup names.
+
+I also want the orthogroups present in aquatics + seagrasses, NOT terrestrials
+
+```{r}
+present_seagrass_and_aquatics_only <- setdiff(intersect(mylistgroup$Aquatics, mylistgroup$Seagrasses), 
+        mylistgroup$Terrestrials)
+print(length(present_seagrass_and_aquatics_only))
+```
+That's the tiny intersection of genes shared between aquatics and seagrasses, let's also get the union of seagrass genes, aquatic genes, and their intersection
+
+```{r}
+present_seagrass_and_aquatics_union <- union(mylistgroup$Aquatics, mylistgroup$Seagrasses)
+print(length(present_seagrass_and_aquatics_union))
+```
+
+OK let's also get the intersection in the middle
+
+```{r}
+present_all <- intersect(intersect(mylistgroup$Terrestrials, mylistgroup$Aquatics), mylistgroup$Seagrasses)
+length(present_all)
+```
+
+Now we pull out these gene IDs and write them out for files to continue in [GOenrichment](GOenrichment.html).
+
+```{r}
+# gene_per_spec has the genes per orthogroup, we need to join with the various set-results we've made
+genes_in_all <- gene_per_spec %>% filter(Orthogroup %in% present_all) %>% separate_rows(value, sep=',') %>% pull(value) %>% str_trim()
+genes_only_in_aquatics_and_seagrasses <- gene_per_spec %>% filter(Orthogroup %in% present_seagrass_and_aquatics_only) %>% separate_rows(value, sep=',') %>% pull(value) %>% str_trim()
+genes_present_seagrass_and_aquatics_union <- gene_per_spec %>% filter(Orthogroup %in% present_seagrass_and_aquatics_union) %>% separate_rows(value, sep=',') %>% pull(value) %>% str_trim()
+
+genes_only_in_aquatics <- gene_per_spec %>% filter(Orthogroup %in% present_aquatics_only) %>% separate_rows(value, sep=',') %>% pull(value) %>% str_trim()
+genes_only_in_seagrasses <- gene_per_spec %>% filter(Orthogroup %in% present_seagrasses_only) %>% separate_rows(value, sep=',') %>% pull(value) %>% str_trim()
+genes_only_in_terrestrials <- gene_per_spec %>% filter(Orthogroup %in% present_terrestrials_only) %>% separate_rows(value, sep=',') %>% pull(value) %>% str_trim()
+```
+
+Now we have a bunch of lists, we can write them into files for the [GO enrichment part](GOenrichment.html)
+
+```{r}
+sink('data/Lost_present_gene_lists/Genes_only_in_Terrestrials.txt')
+cat(genes_only_in_terrestrials, sep='\n')
+sink()
+
+sink('data/Lost_present_gene_lists/Genes_only_in_Seagrasses.txt')
+cat(genes_only_in_seagrasses, sep='\n')
+sink()
+
+sink('data/Lost_present_gene_lists/Genes_only_in_Aquatics.txt')
+cat(genes_only_in_aquatics, sep='\n')
+sink()
+
+sink('data/Lost_present_gene_lists/Genes_only_in_Aquatics_and_Seagrasses.txt')
+cat(genes_only_in_aquatics_and_seagrasses, sep='\n')
+sink()
+
+sink('data/Lost_present_gene_lists/Genes_union_of_Seagrass_and_Aquatics_union.txt')
+cat(genes_present_seagrass_and_aquatics_union, sep='\n')
+sink()
+
+sink('data/Lost_present_gene_lists/Genes_only_in_Aquatics_and_Seagrasses_and_Terrestrials.txt')
+cat(genes_in_all, sep='\n')
+sink()
+
+```
+
+## Dig into seagrass specific differences
+
+In this section, I compare the four seagrasses with each other only, so we can ignore the previous section and go back to the original orthogroups.
+
+```{r}
+seagrass_groups <- groups %>% select(Orthogroup, `A. antarctica`, `P. australis`, `Z. marina`, `Z. muelleri`) %>%
+  rowwise() %>% 
+  mutate(na_count = sum(is.na(c(`A. antarctica`, `P. australis`, `Z. marina`, `Z. muelleri`)))) %>% 
+  filter(na_count != 4) %>% 
+  select(-na_count)
+```
+
+Let's make a Venn diagram like above. We need a list, with the species as keys, and the c(rownumber)
+
+```{r}
+big_list <- list()
+big_list[['A. antarctica']] <- seagrass_groups %>% select(Orthogroup, `A. antarctica`) %>% filter(!is.na(`A. antarctica`)) %>% pull(Orthogroup)
+big_list[['P. australis']] <- seagrass_groups %>% select(Orthogroup, `P. australis`) %>% filter(!is.na(`P. australis`)) %>% pull(Orthogroup)
+big_list[['Z. marina']] <- seagrass_groups %>% select(Orthogroup, `Z. marina`) %>% filter(!is.na(`Z. marina`)) %>% pull(Orthogroup)
+big_list[['Z. muelleri']] <- seagrass_groups %>% select(Orthogroup, `Z. muelleri`) %>% filter(!is.na(`Z. muelleri`)) %>% pull(Orthogroup)
+
+```
+
+```{r fig.width=8}
+venn <- Venn(big_list)
+data <- process_data(venn)
+# code from https://github.com/gaospecial/ggVennDiagram/blob/4cb2aa13c7beae469f9b9836c5d4f94610bd872e/R/ggVennDiagram.R to customise
+region_label <- data@region %>%
+      dplyr::filter(.data$component == "region") %>%
+      dplyr::mutate(percent = paste(round(.data$count*100/sum(.data$count),
+                                          digits = 0),"%", sep="")) %>%
+      dplyr::mutate(both = paste(.data$count,paste0("(",.data$percent,")"),sep = "\n"))
+
+seagrass_venn <- ggplot() +
+    geom_sf(aes_string(fill="count"), data = data@region) +
+    geom_sf(aes_string(color = "id"), data = data@setEdge, show.legend = F,
+            lty = 'solid', size = 1, color='gray') +
+    geom_sf_text(aes_string(label = "name"), data = data@setLabel,
+                 size = NA, fontface='italic',
+                 color = 'black') +
+    theme_void() +
+     geom_sf_label(aes_string(label='both'),
+                             data = region_label,
+                             alpha= 0,
+                             color = 'black',
+                             size = NA,
+                             lineheight = 0.85,
+                             label.size = NA) +
+  scale_fill_gradientn(colours=wes_palette("Zissou1", 100, type = "continuous")) + 
+  theme(legend.position = "none") + 
+  scale_x_continuous(expand = expansion(mult = .2)) #trick from https://github.com/gaospecial/ggVennDiagram/blob/9435aa0ab4abb470c670ecb938c71576461ccedc/vignettes/using-ggVennDiagram.Rmd#L65
+seagrass_venn
+```
+# Both Venns together
+
+```{r fig.height=10, fig.width=8}
+group_venn / seagrass_venn +  plot_annotation(tag_levels = 'A')
+
+```
+
+OK now as above, we dig out the gene names.
+
+```{r}
+# gene_per_spec has the genes per orthogroup, we need to join with the various set-results we've made
+groups_only_in_A_antarctica <- setdiff(setdiff(setdiff(big_list[['A. antarctica']], big_list[['P. australis']]), big_list[['Z. marina']]), big_list[['Z. muelleri']])
+genes_only_in_A_antarctica <- gene_per_spec %>% 
+  filter(name == 'A. antarctica', Orthogroup %in% groups_only_in_A_antarctica) %>% 
+  separate_rows(value, sep=',') %>% pull(value) %>% str_trim()
+
+groups_only_in_P_australis <- setdiff(setdiff(setdiff(big_list[['P. australis']], big_list[['A. antarctica']]), big_list[['Z. marina']]), big_list[['Z. muelleri']])
+genes_only_in_P_australis <- gene_per_spec %>% 
+  filter(name == 'P. australis', Orthogroup %in% groups_only_in_P_australis) %>% 
+  separate_rows(value, sep=',') %>% pull(value) %>% str_trim()
+
+groups_only_in_Z_marina <- setdiff(setdiff(setdiff(big_list[['Z. marina']], big_list[['A. antarctica']]), big_list[['P. australis']]), big_list[['Z. muelleri']])
+genes_only_in_Z_marina <- gene_per_spec %>% 
+  filter(name == 'Z. marina', Orthogroup %in% groups_only_in_Z_marina) %>% 
+  separate_rows(value, sep=',') %>% pull(value) %>% str_trim()
+
+groups_only_in_Z_muelleri <- setdiff(setdiff(setdiff(big_list[['Z. muelleri']], big_list[['A. antarctica']]), big_list[['P. australis']]), big_list[['Z. marina']])
+genes_only_in_Z_muelleri <- gene_per_spec %>% 
+  filter(name == 'Z. muelleri', Orthogroup %in% groups_only_in_Z_muelleri) %>% 
+  separate_rows(value, sep=',') %>% pull(value) %>% str_trim()
+
+```
+
+Now we have a bunch of lists, we can write them into files for the [GO enrichment part](GOenrichment.html)
+
+```{r}
+sink('data/Lost_present_gene_lists/Genes_only_in_A_antarctica_not_other_seagrasses.txt')
+cat(genes_only_in_A_antarctica, sep='\n')
+sink()
+
+sink('data/Lost_present_gene_lists/Genes_only_in_P_australis_not_other_seagrasses.txt')
+cat(genes_only_in_P_australis, sep='\n')
+sink()
+
+sink('data/Lost_present_gene_lists/Genes_only_in_Z_marina_not_other_seagrasses.txt')
+cat(genes_only_in_Z_marina, sep='\n')
+sink()
+
+sink('data/Lost_present_gene_lists/Genes_only_in_Z_muelleri_not_other_seagrasses.txt')
+cat(genes_only_in_Z_muelleri, sep='\n')
+sink()
+
+```

### 2022-02-25 analysis/Orthofinder.Rmd
Fix super huge string lists

@@
 ```{r}
-sink('data/Lost_present_gene_lists/Genes_only_in_Terrestrials.txt')
-cat(genes_only_in_terrestrials, sep='\n')
-sink()
+fileConn<-file('data/Lost_present_gene_lists/Genes_only_in_Terrestrials.txt')
+writeLines(genes_only_in_terrestrials, fileConn)
+close(fileConn)
 
-sink('data/Lost_present_gene_lists/Genes_only_in_Seagrasses.txt')
-cat(genes_only_in_seagrasses, sep='\n')
-sink()
+fileConn<-file('data/Lost_present_gene_lists/Genes_only_in_Seagrasses.txt')
+writeLines(genes_only_in_seagrasses, fileConn)
+close(fileConn)
 
-sink('data/Lost_present_gene_lists/Genes_only_in_Aquatics.txt')
-cat(genes_only_in_aquatics, sep='\n')
-sink()
+fileConn<-file('data/Lost_present_gene_lists/Genes_only_in_Aquatics.txt')
+writeLines(genes_only_in_aquatics, fileConn)
+close(fileConn)
 
-sink('data/Lost_present_gene_lists/Genes_only_in_Aquatics_and_Seagrasses.txt')
-cat(genes_only_in_aquatics_and_seagrasses, sep='\n')
-sink()
+fileConn<-file('Genes_only_in_Aquatics_and_Seagrasses.txt')
+writeLines(genes_only_in_aquatics_and_seagrasses, fileConn)
+close(fileConn)
 
-sink('data/Lost_present_gene_lists/Genes_union_of_Seagrass_and_Aquatics_union.txt')
-cat(genes_present_seagrass_and_aquatics_union, sep='\n')
-sink()
+fileConn<-file('data/Lost_present_gene_lists/Genes_union_of_Seagrass_and_Aquatics_union.txt')
+writeLines(genes_present_seagrass_and_aquatics_union, fileConn)
+close(fileConn)
 
-sink('data/Lost_present_gene_lists/Genes_only_in_Aquatics_and_Seagrasses_and_Terrestrials.txt')
-cat(genes_in_all, sep='\n')
-sink()
-
+fileConn<-file('Genes_only_in_Aquatics_and_Seagrasses_and_Terrestrials.txt')
+writeLines(genes_in_all, fileConn)
+close(fileConn)
 ```
@@
 
+groups_in_all_seagrasses <- union(union(union(big_list[['Z. muelleri']], big_list[['A. antarctica']]), big_list[['P. australis']]), big_list[['Z. marina']])
+
+groups_lost_in_A_antarctica <- setdiff(groups_in_all_seagrasses, big_list[['A. antarctica']])
+genes_lost_in_A_antarctica <- gene_per_spec %>% 
+  filter(name == 'A. antarctica', Orthogroup %in% groups_lost_in_A_antarctica) %>% 
+  separate_rows(value, sep=',') %>% pull(value) %>% str_trim()
+
+groups_lost_in_P_australis <- setdiff(groups_in_all_seagrasses, big_list[['P. australis']])
+genes_lost_in_P_australis <- gene_per_spec %>% 
+  filter(name == 'P. australis', Orthogroup %in% groups_lost_in_P_australis) %>% 
+  separate_rows(value, sep=',') %>% pull(value) %>% str_trim()
+
+groups_lost_in_Z_marina <- setdiff(groups_in_all_seagrasses, big_list[['Z. marina']])
+genes_lost_in_Z_marina <- gene_per_spec %>% 
+  filter(name == 'Z. marina', Orthogroup %in% groups_lost_in_Z_marina) %>% 
+  separate_rows(value, sep=',') %>% pull(value) %>% str_trim()
+
+groups_lost_in_Z_muelleri <- setdiff(groups_in_all_seagrasses, big_list[['Z. muelleri']])
+genes_lost_in_Z_muelleri <- gene_per_spec %>% 
+  filter(name == 'Z. muelleri', Orthogroup %in% groups_lost_in_Z_muelleri) %>% 
+  separate_rows(value, sep=',') %>% pull(value) %>% str_trim()
 ```
@@
 ```{r}
-sink('data/Lost_present_gene_lists/Genes_only_in_A_antarctica_not_other_seagrasses.txt')
-cat(genes_only_in_A_antarctica, sep='\n')
-sink()
+fileConn<-file('data/Lost_present_gene_lists/Genes_only_in_A_antarctica_not_other_seagrasses.txt')
+writeLines(genes_only_in_A_antarctica, fileConn)
+close(fileConn)
 
-sink('data/Lost_present_gene_lists/Genes_only_in_P_australis_not_other_seagrasses.txt')
-cat(genes_only_in_P_australis, sep='\n')
-sink()
+fileConn<-file('data/Lost_present_gene_lists/Genes_only_in_P_australis_not_other_seagrasses.txt')
+writeLines(genes_only_in_P_australis, fileConn)
+close(fileConn)
 
-sink('data/Lost_present_gene_lists/Genes_only_in_Z_marina_not_other_seagrasses.txt')
-cat(genes_only_in_Z_marina, sep='\n')
-sink()
+fileConn<-file('data/Lost_present_gene_lists/Genes_only_in_Z_marina_not_other_seagrasses.txt')
+writeLines(genes_only_in_Z_marina, fileConn)
+close(fileConn)
 
-sink('data/Lost_present_gene_lists/Genes_only_in_Z_muelleri_not_other_seagrasses.txt')
-cat(genes_only_in_Z_muelleri, sep='\n')
-sink()
+fileConn<-file('data/Lost_present_gene_lists/Genes_only_in_Z_muelleri_not_other_seagrasses.txt')
+writeLines(genes_only_in_Z_muelleri, fileConn)
+close(fileConn)
 
+fileConn<-file('data/Lost_present_gene_lists/Genes_lost_in_A_antarctica_not_other_seagrasses.txt')
+writeLines(genes_lost_in_A_antarctica, fileConn)
+close(fileConn)
+
+fileConn<-file('data/Lost_present_gene_lists/Genes_lost_in_P_australis_not_other_seagrasses.txt')
+writeLines(genes_lost_in_P_australis, fileConn)
+close(fileConn)
+
+fileConn<-file('data/Lost_present_gene_lists/Genes_lost_in_Z_marina_not_other_seagrasses.txt')
+writeLines(genes_lost_in_Z_marina, fileConn)
+close(fileConn)
+
+fileConn<-file('data/Lost_present_gene_lists/Genes_lost_in_Z_muelleri_not_other_seagrasses.txt')
+writeLines(genes_lost_in_Z_muelleri, fileConn)
+close(fileConn)
 ```

### 2022-02-25 analysis/Orthofinder.Rmd
Fix typo

@@
 
-fileConn<-file('Genes_only_in_Aquatics_and_Seagrasses.txt')
+fileConn<-file('data/Lost_present_gene_lists/Genes_only_in_Aquatics_and_Seagrasses.txt')
 writeLines(genes_only_in_aquatics_and_seagrasses, fileConn)
@@
 
-fileConn<-file('Genes_only_in_Aquatics_and_Seagrasses_and_Terrestrials.txt')
+fileConn<-file('data/Lost_present_gene_lists/Genes_only_in_Aquatics_and_Seagrasses_and_Terrestrials.txt')
 writeLines(genes_in_all, fileConn)

### 2022-02-25 analysis/Orthofinder.Rmd
Fix bug in the very last lines

@@
 genes_lost_in_A_antarctica <- gene_per_spec %>% 
-  filter(name == 'A. antarctica', Orthogroup %in% groups_lost_in_A_antarctica) %>% 
+  filter(name != 'A. antarctica', Orthogroup %in% groups_lost_in_A_antarctica) %>% 
   separate_rows(value, sep=',') %>% pull(value) %>% str_trim()
@@
 genes_lost_in_P_australis <- gene_per_spec %>% 
-  filter(name == 'P. australis', Orthogroup %in% groups_lost_in_P_australis) %>% 
+  filter(name != 'P. australis', Orthogroup %in% groups_lost_in_P_australis) %>% 
   separate_rows(value, sep=',') %>% pull(value) %>% str_trim()
@@
 genes_lost_in_Z_marina <- gene_per_spec %>% 
-  filter(name == 'Z. marina', Orthogroup %in% groups_lost_in_Z_marina) %>% 
+  filter(name != 'Z. marina', Orthogroup %in% groups_lost_in_Z_marina) %>% 
   separate_rows(value, sep=',') %>% pull(value) %>% str_trim()
@@
 genes_lost_in_Z_muelleri <- gene_per_spec %>% 
-  filter(name == 'Z. muelleri', Orthogroup %in% groups_lost_in_Z_muelleri) %>% 
+  filter(name != 'Z. muelleri', Orthogroup %in% groups_lost_in_Z_muelleri) %>% 
   separate_rows(value, sep=',') %>% pull(value) %>% str_trim()

### 2022-02-25 analysis/GOenrichment.Rmd
add some more changes
Replace GO enrichment

@@
 library(rrvgo)
+library(kableExtra)
+library(patchwork)
 knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
@@
 ```
-
-# lost in seagrasses, each one uniquely vs all terrestrials
-```{r eval=FALSE}
-compare('lost_in_amphi_vs_all.txt', 'missing_amphi_vs_all_GO.txt', allgenes, annAT)
-compare('lost_in_posi_vs_all.txt', 'missing_posi_vs_all_GO.txt', allgenes, annAT)
-compare('lost_in_zmar_vs_all.txt', 'missing_zmar_vs_all_GO.txt', allgenes, annAT)
-compare('lost_in_zmuel_vs_all.txt', 'missing_zmuel_vs_all_GO.txt', allgenes, annAT)
-```
-
-# lost in seagrasses vs all terrestrials
-Here we compare GO terms for seagrasses and aquatics (seagrasses+duckweeds) vs all terrestrials
+# Comparing the three big groups
 
 ```{r eval=FALSE}
+# all genes shared - this runs out of memory
+#compare('Lost_present_gene_lists/Genes_only_in_Aquatics_and_Seagrasses_and_Terrestrials.txt', 'output/GO_results_genes_in_Aquatics_and_Seagrasses_and_Terrestrials.csv', allgenes, annAT)
 
-compare('lost_in_seagrasses.txt', 'missing_seagrasses_GO.txt', allgenes, annAT)
-compare('lost_in_aquatics.txt', 'missing_aquatics_GO.txt', allgenes, annAT)
-compare('only_in_seagrasses.txt', 'only_seagrasses_GO.txt', allgenes, annAT)
+# Genes in aquatics and seagrasses - IMPORTANT This is NOT the union - this is the intersection!
+# Union is below
+compare('Lost_present_gene_lists/Genes_only_in_Aquatics_and_Seagrasses.txt', 'output/GO_results_genes_in_Aquatics_and_Seagrasses_NOT_Terrestrials.csv', allgenes, annAT)
+
+# Genes only in aquatics
+compare('Lost_present_gene_lists/Genes_only_in_Aquatics.txt', 'output/GO_results_genes_in_Aquatics_NOT_Seagrasses_NOT_Terrestrials.csv', allgenes, annAT)
+
+# Genes only in seagrasses
+compare('Lost_present_gene_lists/Genes_only_in_Seagrasses.txt', 'output/GO_results_genes_in_Seagrasses_NOT_Aquatics_NOT_Terrestrials.csv', allgenes, annAT)
+
+# Genes only in terrestrials
+compare('Lost_present_gene_lists/Genes_only_in_Terrestrials.txt', 'output/GO_results_genes_in_Terrestrials_NOT_Aquatics_NOT_Seagrasses.csv', allgenes, annAT)
+
+# Genes union of Seagrass and Aquatics
+compare('Lost_present_gene_lists/Genes_union_of_Seagrass_and_Aquatics_union.txt', 'output/GO_results_genes_union_Aquatics_and_Seagrasses_NOT_Terrestrials.csv', allgenes, annAT)
 
 ```
-
 # Seagrass only comparisons
@@
 
-load('annAtObject.RData')
+load('sannAtObject.RData')
 sallgenes <- unique(unlist(sannAT))
 
-compare('lost_in_amphi.txt', 'lost_in_amphi_vs_seagrasses_GO.txt', sallgenes, sannAT)
-compare('lost_in_posi.txt', 'lost_in_posi_vs_seagrasses_GO.txt', sallgenes, sannAT)
-compare('lost_in_zmar.txt', 'lost_in_zmar_vs_seagrasses_GO.txt', sallgenes, sannAT)
-compare('lost_in_zmuel.txt', 'lost_in_zmuel_vs_seagrasses_GO.txt', sallgenes, sannAT)
+```
+# lost in seagrasses
+
+```{r eval=FALSE}
+compare('Lost_present_gene_lists/Genes_lost_in_A_antarctica_not_other_seagrasses.txt', 'output/GO_results_genes_lost_A_antarctica_not_other_seagrasses.csv', sallgenes, sannAT)
+
+compare('Lost_present_gene_lists/Genes_lost_in_Z_marina_not_other_seagrasses.txt', 'output/GO_results_genes_lost_Z_marina_not_other_seagrasses.csv', sallgenes, sannAT)
+
+compare('Lost_present_gene_lists/Genes_lost_in_P_australis_not_other_seagrasses.txt', 'output/GO_results_genes_lost_P_australis_not_other_seagrasses.csv', sallgenes, sannAT)
+
+compare('Lost_present_gene_lists/Genes_lost_in_Z_muelleri_not_other_seagrasses.txt', 'output/GO_results_genes_lost_Z_muelleri_not_other_seagrasses.csv', sallgenes, sannAT)
 ```
@@
 
-The opposite - which GO-terms are present only in one of the four species?
+```{r eval=FALSE}
+compare('Lost_present_gene_lists/Genes_only_in_Z_marina_not_other_seagrasses.txt', 'output/GO_results_genes_only_Z_marina_not_other_seagrasses.csv', sallgenes, sannAT)
 
-```{r eval=FALSE}
-compare('only_in_amphi.txt', 'only_in_amphi_vs_seagrasses_GO.txt', sallgenes, sannAT)
-compare('only_in_posi.txt', 'only_in_posi_vs_seagrasses_GO.txt', sallgenes, sannAT)
-compare('only_in_zmar.txt', 'only_in_zmar_vs_seagrasses_GO.txt', sallgenes, sannAT)
-compare('only_in_zmuel.txt', 'only_in_zmuel_vs_seagrasses_GO.txt', sallgenes, sannAT)
+compare('Lost_present_gene_lists/Genes_only_in_Z_muelleri_not_other_seagrasses.txt', 'output/GO_results_genes_only_Z_muelleri_not_other_seagrasses.csv', sallgenes, sannAT)
+
+compare('Lost_present_gene_lists/Genes_only_in_P_australis_not_other_seagrasses.txt', 'output/GO_results_genes_only_P_australis_not_other_seagrasses.csv', sallgenes, sannAT)
+
+compare('Lost_present_gene_lists/Genes_only_in_A_antarctica_not_other_seagrasses.txt', 'output/GO_results_genes_only_A_antarctica_not_other_seagrasses.csv', sallgenes, sannAT)
 ```
@@
 
-for (f in list.files('./data/', pattern='GO.txt')){
-  filename <- paste('./data/', f, sep='')
+for (f in list.files('./output/', pattern='*.csv')){
+  filename <- paste('./output/', f, sep='')
   go_and_pvalues <- readChar(filename, file.info(filename)$size)
@@
   one.data <- one.data [(one.data$plot_X != "null" & one.data$plot_Y != "null") & (! is.na(one.data$frequency)) & (one.data$value != 'null'), ];
-  one.data <- type_convert(one.data)
+  one.data <- suppressMessages(type_convert(one.data)) # guess data types, but also: shush
   top_ten_values <- one.data %>% arrange(value) %>% head(10) %>% pull(term_ID) 
@@
     geom_label_repel(max.overlaps = Inf, box.padding = 0.5,
-                     aes(point.size = log_size)) +
+                     aes(point.size = log_size), alpha=0.9) +
     theme_minimal() +
-    labs(y = "Semantic space x",
-         x = "Semantic space y",
-         color = 'Log(p-value)',
+    labs(color = 'log(p-value)',
          size = 'Frequency') +
@@
       axis.text.y = element_blank(),
-      axis.ticks.y = element_blank()
+      axis.ticks.y = element_blank(),
+      axis.title.x = element_blank(),
+      axis.title.y = element_blank()
     ) +
@@
   print(i)
-  print(knitr::kable(results_list[[i]] %>% filter(Eliminated == FALSE) %>% select(TermID, Name, Value) %>% arrange(Value), caption=i))
+  print(results_list[[i]] %>% filter(Eliminated == FALSE) %>% select(TermID, Name, Value) %>% arrange(Value) %>% kbl() %>%  kable_styling())
   cat("\n")
@@
 ```
-
-# Venn diagram GO overlaps
-
-Let's also check which GO terms overlap between the 4 seagrasses!
 
@@
 
-  cowplot::save_plot(file_name, plot_list[[i]], base_height=8)
+  cowplot::save_plot(file_name, plot_list[[i]], base_height=6)
 
@@
 
-How many shared lost GO-terms are there? Hopefully, all four species will have lost the most GO-terms.
+Let me also try and make some patchworks.
+
+```{r fig.width=10}
+# For some reason with this plot the legends aren't correctly collected by patchwork, 
+# so I need to manually turn one off
+patchy <- (plot_list[["GO_results_genes_in_Terrestrials_NOT_Aquatics_NOT_Seagrasses.csv"]] + theme(legend.position = "none")) /
+  plot_list[["GO_results_genes_union_Aquatics_and_Seagrasses_NOT_Terrestrials.csv"]] + plot_layout(guides = 'collect') + plot_annotation(tag_levels = 'A')
+patchy
+cowplot::save_plot('output/GO_results_terrestrials_vs_union_aquatics_seagrasses.png', patchy, base_height=7)
+```
+
+OK I'll better make this figure for the seagrasses alone, it's quite messy this way.
+
+# Seagrass-specific loss
+
+Let's check which GO terms overlap between the 4 seagrasses!
+
+How many shared lost GO-terms are there?
+
 ```{r}
-a <- list(`P. australis` = results_list$missing_posi_vs_all_GO.txt$Name,
-          `A. antarctica` = results_list$missing_amphi_vs_all_GO.txt$Name,
-          `Z. marina` = results_list$missing_zmar_vs_all_GO.txt$Name,
-          `Z. muelleri` = results_list$missing_zmuel_vs_all_GO.txt$Name,
-          `A. thaliana` = results_list$missing_arabidopsis_vs_all_GO.txt$Name)
+a <- list(`P. australis` = results_list$GO_results_genes_lost_P_australis_not_other_seagrasses.csv$Name,
+          `A. antarctica` = results_list$GO_results_genes_lost_A_antarctica_not_other_seagrasses.csv$Name,
+          `Z. marina` = results_list$GO_results_genes_lost_Z_marina_not_other_seagrasses.csv$Name,
+          `Z. muelleri` = results_list$GO_results_genes_lost_Z_muelleri_not_other_seagrasses.csv$Name)
 
-a_go_ids <- list(`P. australis` = results_list$missing_posi_vs_all_GO.txt$TermID,
-          `A. antarctica` = results_list$missing_amphi_vs_all_GO.txt$TermID,
-          `Z. marina` = results_list$missing_zmar_vs_all_GO.txt$TermID,
-          `Z. muelleri` = results_list$missing_zmuel_vs_all_GO.txt$TermID,
-          `A. thaliana` = results_list$missing_arabidopsis_vs_all_GO.txt$TermID)
+a_go_ids <- list(`P. australis` = results_list$GO_results_genes_lost_P_australis_not_other_seagrasses.csv$TermID,
+          `A. antarctica` = results_list$GO_results_genes_lost_A_antarctica_not_other_seagrasses.csv$TermID,
+          `Z. marina` = results_list$GO_results_genes_lost_Z_marina_not_other_seagrasses.csv$TermID,
+          `Z. muelleri` = results_list$GO_results_genes_lost_Z_muelleri_not_other_seagrasses.csv$TermID)
 ```
@@
 ```{r}
-plot(euler(a), 
+plot(venn(a), 
      quantities = TRUE,
@@
 ```{r}
-setdiff(Reduce(intersect, a_no_ara), Reduce(intersect, a)) %>% enframe() %>% writexl::write_xlsx('./data/shared_lost_genes.xlsx')
+Reduce(union, a) %>% enframe() %>% writexl::write_xlsx('./output/Seagrasses_shared_lost_genes.xlsx')
 ```
@@
 ```{r}
-writexl::write_xlsx(results_d, 'data/Lost_GO_terms_in_five_species.xlsx')
+writexl::write_xlsx(results_d, 'output/Lost_GO_terms_in_five_species.xlsx')
 ```
 
-# after filtering for plant-specific GO-terms
+# After filtering for plant-specific GO-terms
 
@@
 ```{r}
-results_d %>% filter(GOID %in% plantSpecificGO) %>% writexl::write_xlsx('data/Lost_GO_terms_in_five_species.PlantSpecific.xlsx')
+results_d %>% filter(GOID %in% plantSpecificGO) %>% writexl::write_xlsx('output/Lost_GO_terms_in_five_species.PlantSpecific.xlsx')
 ```
@@
 ```{r}
-plot(euler(newa), 
+plot(venn(newa), 
      quantities = TRUE,

### 2022-02-28 analysis/GOenrichment.Rmd
accidentally did not display REVIGO code

@@
 
-```{r warning=FALSE, echo=FALSE, message=FALSE}
+```{r warning=FALSE, message=FALSE}
 # lots of warnings and messages here so I'm hiding these

### 2022-03-01 analysis/license.Rmd
Add general descriptions

---
title: "License"
output:
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: console
---

MIT License

Copyright (c) 2022 Philipp Bayer

Permission is hereby granted, free of charge, to any person obtaining a copy
of this software and associated documentation files (the "Software"), to deal
in the Software without restriction, including without limitation the rights
to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
copies of the Software, and to permit persons to whom the Software is
furnished to do so, subject to the following conditions:

The above copyright notice and this permission notice shall be included in all
copies or substantial portions of the Software.

THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
SOFTWARE.

### 2022-03-08 analysis/Orthofinder.Rmd
split aquatics up

@@
     name %in% c(
-      'C. reinhardtii',
       'L. gibba',
       'S. polyrhiza',
-      'O. lucimarinus',
       'W. australis'
-    ) ~ 'Aquatics',
-    name %in% c('P. australis', 'Z. muelleri', 'Z. marina', 'A. antarctica') ~ 'Seagrasses'
+    ) ~ 'Duckweeds',
+    name %in% c(
+      'C. reinhardtii',
+      'O. lucimarinus'
+    ) ~ 'Algae',
+    name %in% c(
+      'P. australis',
+      'Z. muelleri',
+      'Z. marina',
+      'A. antarctica'
+      ) ~ 'Seagrasses'
   )) 
@@
 The biggest chunk of orthogroups is not in aquatics or seagrasses with 13562 orthogroups.
-3437 orthogroups are unique to aquatics, but not in seagrasses - 1183 orthogroups are unique in seagrasses, but not in aquatics.
-404 + 3437 + 2217 orthogroups are not in terrestrials.
+3845 orthogroups are in duckweeds, seagrasses, terrestrials, but not algae - 2217 orthogroups are unique in seagrasses.
 
-OK now we need to do the GOenrichment for those three groups.
+OK now we need to do the GOenrichment for those four groups.
 
-Let's also make an Venn plot, we have only three groups:
+Let's also make an Venn plot, we have only four groups:
 ```{r fig.width=8}
@@
     name %in% c(
-      'C. reinhardtii',
       'L. gibba',
       'S. polyrhiza',
-      'O. lucimarinus',
       'W. australis'
-    ) ~ 'Aquatics',
-    name %in% c('P. australis', 'Z. muelleri', 'Z. marina', 'A. antarctica') ~ 'Seagrasses'
+    ) ~ 'Duckweeds',
+    name %in% c(
+      'C. reinhardtii',
+      'O. lucimarinus'
+    ) ~ 'Algae',
+    name %in% c(
+      'P. australis',
+      'Z. muelleri', 'Z. marina', 'A. antarctica') ~ 'Seagrasses'
   )) 
@@
 ```{r}
-present_aquatics_only <- setdiff(setdiff(mylistgroup$Aquatics, mylistgroup$Terrestrials), 
-                         mylistgroup$Seagrasses)
-present_seagrasses_only <- setdiff(setdiff(mylistgroup$Seagrasses, mylistgroup$Aquatics), 
-                           mylistgroup$Terrestrials)
-present_terrestrials_only <- setdiff(setdiff(mylistgroup$Terrestrials, mylistgroup$Aquatics),
-                             mylistgroup$Seagrasses)
-print(cbind(length(present_aquatics_only), length(present_seagrasses_only), length(present_terrestrials_only)))
+present_algae_only <- setdiff(setdiff(setdiff(mylistgroup$Algae, mylistgroup$Duckweeds), 
+                         mylistgroup$Seagrasses), mylistgroup$Terrestrials)
+present_seagrasses_only <- setdiff(setdiff(setdiff(mylistgroup$Seagrasses, mylistgroup$Duckweeds),                            mylistgroup$Terrestrials), mylistgroup$Algae)
+present_terrestrials_only <- setdiff(setdiff(setdiff(mylistgroup$Terrestrials, mylistgroup$Duckweeds), mylistgroup$Seagrasses), mylistgroup$Algae)
+present_duckweeds_only <- setdiff(setdiff(setdiff(mylistgroup$Duckweeds, mylistgroup$Terrestrials), mylistgroup$Seagrasses), mylistgroup$Algae)
+
+print(cbind(length(present_algae_only), length(present_seagrasses_only), length(present_terrestrials_only), length(present_duckweeds_only)))
 
 ```
-OK those numbers fit with the above Venn diagram. I now have three lists of orthogroup names.
+OK those numbers fit with the above Venn diagram. I now have four lists of orthogroup names.
 
@@
 ```{r}
-present_seagrass_and_aquatics_only <- setdiff(intersect(mylistgroup$Aquatics, mylistgroup$Seagrasses), 
+present_seagrass_and_aquatics_only <- setdiff(intersect(intersect(mylistgroup$Duckweeds, mylistgroup$Seagrasses), mylistgroup$Algae),
         mylistgroup$Terrestrials)
@@
 ```{r}
-present_seagrass_and_aquatics_union <- union(mylistgroup$Aquatics, mylistgroup$Seagrasses)
+present_seagrass_and_aquatics_union <- union(mylistgroup$Algae, union(mylistgroup$Duckweeds, mylistgroup$Seagrasses))
 print(length(present_seagrass_and_aquatics_union))
@@
 ```{r}
-present_all <- intersect(intersect(mylistgroup$Terrestrials, mylistgroup$Aquatics), mylistgroup$Seagrasses)
+present_all <- intersect(intersect(mylistgroup$Algae, intersect(mylistgroup$Terrestrials, mylistgroup$Duckweeds)), mylistgroup$Seagrasses)
 length(present_all)
@@
 genes_only_in_aquatics_and_seagrasses <- gene_per_spec %>% filter(Orthogroup %in% present_seagrass_and_aquatics_only) %>% separate_rows(value, sep=',') %>% pull(value) %>% str_trim()
+
 genes_present_seagrass_and_aquatics_union <- gene_per_spec %>% filter(Orthogroup %in% present_seagrass_and_aquatics_union) %>% separate_rows(value, sep=',') %>% pull(value) %>% str_trim()
 
-genes_only_in_aquatics <- gene_per_spec %>% filter(Orthogroup %in% present_aquatics_only) %>% separate_rows(value, sep=',') %>% pull(value) %>% str_trim()
+genes_only_in_duckweeds <- gene_per_spec %>% filter(Orthogroup %in% present_duckweeds_only) %>% separate_rows(value, sep=',') %>% pull(value) %>% str_trim()
+genes_only_in_algae <- gene_per_spec %>% filter(Orthogroup %in% present_algae_only) %>% separate_rows(value, sep=',') %>% pull(value) %>% str_trim()
 genes_only_in_seagrasses <- gene_per_spec %>% filter(Orthogroup %in% present_seagrasses_only) %>% separate_rows(value, sep=',') %>% pull(value) %>% str_trim()
@@
 
-fileConn<-file('data/Lost_present_gene_lists/Genes_only_in_Aquatics.txt')
-writeLines(genes_only_in_aquatics, fileConn)
+fileConn<-file('data/Lost_present_gene_lists/Genes_only_in_Duckweeds.txt')
+writeLines(genes_only_in_duckweeds, fileConn)
+close(fileConn)
+
+fileConn<-file('data/Lost_present_gene_lists/Genes_only_in_Algae.txt')
+writeLines(genes_only_in_algae, fileConn)
 close(fileConn)
@@
 ```
+
 # Both Venns together
+
+Let's save both Venns so I can combine them with other plots in other Rmds.
+
+```{r}
+save(group_venn, file='output/group_venn_image.Rdata')
+save(seagrass_venn, file='output/seagrass_venn_image.Rdata')
+```
 
@@
 group_venn / seagrass_venn +  plot_annotation(tag_levels = 'A')
-
 ```

### 2022-03-08 analysis/GeneLossComparisonReads.Rmd
workflowr::wflow_publish(files = "analysis/GeneLossComparisonReads.Rmd")

@@
 stomata_long <- stomata %>% pivot_longer(-Gene.symbol, values_to = 'Coverage', names_to='Species') %>% 
-  mutate(Species = gsub('\\.\\.', '. ', Species))
+  mutate(Species = gsub('\\.\\.', '. ', Species)) %>% 
+  filter(Species != 'H. ovalis')
 ```
@@
 stom_p
+```
+
+```{r}
+label <- stomata_long #%>% mutate(Gene.symbol = stringr::str_to_lower(Gene.symbol)) 
+label <- label %>% 
+  mutate(mylabel = case_when(Coverage >= 100 ~ '',
+                             Coverage == 0 ~ '',
+                                    TRUE ~ as.character(Coverage)))
+
+stom_heat <- stomata_long %>% 
+  #mutate(Gene.symbol = stringr::str_to_lower(Gene.symbol)) %>% 
+  ggplot(aes(x=Species, y=factor(Gene.symbol, level=rev(unique(Gene.symbol))), fill=Coverage)) + 
+  geom_tile(alpha=0.9, color='white') + 
+  geom_text(aes(label=mylabel), data=label, size=3.5) + 
+  theme(axis.text.y = element_text(face = "italic"),
+        axis.text.x = element_text(face='italic'),
+        axis.title.x = element_blank(),
+        axis.title.y = element_blank(),
+        legend.position = "none") +
+  scale_fill_continuous(limits=c(0,100))
+
+stom_heat
 ```
@@
 ethylene_long <- ethylene %>% pivot_longer(-Gene.symbol, values_to = 'Coverage', names_to='Species') %>% 
-  mutate(Species = gsub('\\.\\.', '. ', Species))
+  mutate(Species = gsub('\\.\\.', '. ', Species)) %>% 
+  filter(Species != 'H. ovalis')
 ```
@@
 ```
+
+```{r}
+eth_heat <- ethylene_long %>% 
+  dplyr::filter(Gene.symbol != 'ndhB-2') %>% 
+  mutate(Gene.symbol = stringr::str_to_lower(Gene.symbol)) %>% 
+  ggplot(aes(x=Species, y=factor(Gene.symbol, level=rev(unique(Gene.symbol))), fill=Coverage)) + 
+  geom_tile() + 
+  ylab('Gene') +
+  theme(axis.text.y = element_text(face = "italic"),
+        axis.text.x = element_text(face='italic')) +
+  scale_fill_continuous(limits=c(0,100))
+
+eth_heat
+```
+
 # NDH genes
@@
 ndh_long <- ndh %>% pivot_longer(-Gene.symbol, values_to = 'Coverage', names_to='Species') %>% 
-  mutate(Species = gsub('\\.\\.', '. ', Species))
+  mutate(Species = gsub('\\.\\.', '. ', Species)) %>% 
+  filter(Species != 'H. ovalis')
 ```
@@
 ndh_p
+```
+```{r fig.width=10}
+label <- ndh_long
+label <- label %>% 
+  mutate(mylabel = case_when(Coverage >= 100 ~ '',
+                             Coverage == 0 ~ '',
+                                    TRUE ~ as.character(Coverage)))
+
+ndh_heat <- ndh_long %>% 
+  dplyr::filter(Gene.symbol != 'ndhB-2') %>% 
+  mutate(Gene.symbol = stringr::str_to_lower(Gene.symbol)) %>% 
+  ggplot(aes(x=Species, y=factor(Gene.symbol, level=rev(unique(Gene.symbol))), fill=Coverage)) + 
+  geom_tile() + 
+  geom_tile(alpha=0.9, color='white') + 
+  #geom_text(data=label, aes(label=mylabel), size=3.5) +
+  theme(axis.text.y = element_text(face = "italic"),
+        axis.text.x = element_text(face = 'italic'),
+        axis.title.x = element_blank(),
+        axis.title.y = element_blank())  +
+  scale_fill_continuous(limits=c(0,100))
+
+ndh_heat
 ```
@@
 ndh_chl_long <- ndh_chloro %>% pivot_longer(-Gene.symbol, values_to = 'Coverage', names_to='Species') %>% 
-  mutate(Species = gsub('\\.\\.', '. ', Species))
+  mutate(Species = gsub('\\.\\.', '. ', Species)) %>% 
+  filter(Species != 'H. ovalis')
 ```
@@
 
+```{r}
+label <- ndh_chl_long %>% filter(Gene.symbol != 'ndhB-2')
+label <- label %>% 
+  mutate(mylabel = case_when(Coverage >= 100 ~ '',
+                             Coverage == 0 ~ '',
+                                    TRUE ~ as.character(Coverage)))
+
+ndh_chl_heat <- ndh_chl_long %>% 
+  dplyr::filter(Gene.symbol != 'ndhB-2') %>% 
+  mutate(Coverage = case_when(Coverage > 100 ~ 100,
+                              TRUE ~ Coverage)) %>% 
+  ggplot(aes(x=Species, y=factor(Gene.symbol, level=rev(unique(Gene.symbol))), fill=Coverage)) + 
+  geom_tile(alpha=0.9, color='white') + 
+  geom_text(data=label, aes(label=mylabel), size=3.5) +
+  ylab('Gene') +
+  theme(axis.text.y = element_text(face = "italic"),
+        axis.text.x = element_text(face = 'italic'),
+        axis.title.x = element_blank(),
+        axis.title.y = element_blank(),
+        legend.position = "none") +
+  scale_fill_continuous(limits=c(0,100))
+
+ndh_chl_heat
+```
+
 # all together
@@
 ```
+
+```{r fig.height=7}
+patchwork <- (eth_p + ndh_chl_heat)/
+ (ndh_p + stom_heat  )
+#patchwork[[2]] = patchwork[[2]] + theme(axis.title.x = element_blank() )
+#patchwork[[1]] = patchwork[[1]] + theme(axis.title.x = element_blank() )
+puh <- patchwork+ plot_annotation(tag_levels = "A") #+ plot_layout(guides='none')
+puh
+cowplot::save_plot(filename = 'output/patched_gene_loss.png', plot = puh, base_height = 7)
+```
+
+Hmmmmmmmmm

### 2022-03-08 analysis/Orthofinder.Rmd
Add files

@@
 
-# FINAL version redoing - terrestrial vs aquatics vs seagrasses
+# FINAL version redoing - terrestrial vs duckweeds vs algae vs seagrasses
 
@@
 MUCH better.
-The biggest chunk of orthogroups is not in aquatics or seagrasses with 13562 orthogroups.
+The biggest chunk of orthogroups is only in terrestrials with 13562 orthogroups.
 3845 orthogroups are in duckweeds, seagrasses, terrestrials, but not algae - 2217 orthogroups are unique in seagrasses.

### 2022-03-09 analysis/Orthofinder.Rmd
add better groups

@@
                              label.size = NA) +
-  scale_fill_gradientn(colours=wes_palette("Zissou1", 100, type = "continuous")) + 
+  scale_fill_gradientn(colours=wes_palette("Zissou1", 3, type = "continuous")) + 
   theme(legend.position = "none") + 
@@
 
-groups_in_all_seagrasses <- union(union(union(big_list[['Z. muelleri']], big_list[['A. antarctica']]), big_list[['P. australis']]), big_list[['Z. marina']])
+groups_in_all_seagrasses_union <- union(union(union(big_list[['Z. muelleri']], big_list[['A. antarctica']]), big_list[['P. australis']]), big_list[['Z. marina']])
+genes_in_all_seagrasses_union <- gene_per_spec %>% 
+  filter(name %in% c('A. antarctica', 'P. australis', 'Z. muelleri', 'Z. marina'), Orthogroup %in% groups_in_all_seagrasses_union) %>% 
+  separate_rows(value, sep=',') %>% pull(value) %>% str_trim()
 
-groups_lost_in_A_antarctica <- setdiff(groups_in_all_seagrasses, big_list[['A. antarctica']])
+groups_in_all_seagrasses_intersect <- intersect(intersect(intersect(big_list[['Z. muelleri']], big_list[['A. antarctica']]), big_list[['P. australis']]), big_list[['Z. marina']])
+genes_in_all_seagrasses_intersect <- gene_per_spec %>% 
+  filter(name %in% c('A. antarctica', 'P. australis', 'Z. muelleri', 'Z. marina'), Orthogroup %in% groups_in_all_seagrasses_intersect) %>% 
+  separate_rows(value, sep=',') %>% pull(value) %>% str_trim()
+
+groups_lost_in_A_antarctica <- setdiff(groups_in_all_seagrasses_union, big_list[['A. antarctica']])
 genes_lost_in_A_antarctica <- gene_per_spec %>% 
@@
 
-groups_lost_in_P_australis <- setdiff(groups_in_all_seagrasses, big_list[['P. australis']])
+groups_lost_in_P_australis <- setdiff(groups_in_all_seagrasses_union, big_list[['P. australis']])
 genes_lost_in_P_australis <- gene_per_spec %>% 
@@
 
-groups_lost_in_Z_marina <- setdiff(groups_in_all_seagrasses, big_list[['Z. marina']])
+groups_lost_in_Z_marina <- setdiff(groups_in_all_seagrasses_union, big_list[['Z. marina']])
 genes_lost_in_Z_marina <- gene_per_spec %>% 
@@
 
-groups_lost_in_Z_muelleri <- setdiff(groups_in_all_seagrasses, big_list[['Z. muelleri']])
+groups_lost_in_Z_muelleri <- setdiff(groups_in_all_seagrasses_union, big_list[['Z. muelleri']])
 genes_lost_in_Z_muelleri <- gene_per_spec %>% 
@@
   separate_rows(value, sep=',') %>% pull(value) %>% str_trim()
+
 ```
@@
 writeLines(genes_only_in_P_australis, fileConn)
+close(fileConn)
+
+fileConn <- file('data/Lost_present_gene_lists/Genes_in_all_seagrasses_union.txt')
+writeLines(genes_in_all_seagrasses_union, fileConn)
+close(fileConn)
+
+fileConn <- file('data/Lost_present_gene_lists/Genes_in_all_seagrasses_intersect.txt')
+writeLines(genes_in_all_seagrasses_intersect, fileConn)
 close(fileConn)

### 2022-03-09 analysis/index.Rmd
Add missing files
wflow_publish(files = list.files("analysis/", pattern = "*Rmd",     full.names = T))
wflow_publish(files = c("analysis/*"))
wflow_publish("analysis/index.Rmd")
add better groups

@@
 
-This website collects comparisons of Amphibolis and Posidonia comparative analyses.
+This website collects comparisons of *Amphibolis* and *Posidonia* comparative analyses.
 
-[Repeat comparison plots](plot.html)
+[Repeat comparison plots](plot.html) - Comparing genome sizes and repeat/TE content
 
-[Phylogenies](OTT.html)
+[Phylogenies](OTT.html) - plotting just phylogenies
+
+[Clustering results](Orthofinder.html) - analysing Orthofinder clustering results
+
+[R gene plotting](plotRgenes.html) - making phylogenies and showing R-gene number differences
+
+[Gene Loss comparison by reads](GeneLossComparisonReads.html) - making heatmaps and barcharts to compare read alignments
+
+[GO enrichment](GOenrichment.html) - running topGO and REVIGO to make nice plots of GO-enrichment
+
+[GO and Ortho](GO_and_ortho.html) - merging plots from GO enrichment and Orthofinder analysis

### 2022-03-09 analysis/GO_and_ortho.Rmd
add more ifles

---
title: "Visualising gene differences based on read alignments"
author: "Philipp Bayer"
date: "9th March 2022"
output: html_document
---

```{r setup}
library(tidyverse)

knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

```

In this one I just merge some plots from the Orthofinder and GO enrichment Rmds, which I have saved as Rdata objects.

```{r}
load('output/all_GO_plots.Rdata')
load('output/group_venn_image.Rdata')
load('output/seagrass_venn_image.Rdata')
```
```{r}
group_venn
```
```{r}
seagrass_venn
```

OK now we can make patchworks with these Venn diagrams and their associated GO-terms
```{r fig.width=8, fig.height=10}
sum1 <- group_venn / (plot_list$GO_results_genes_in_Terrestrials_NOT_Aquatics_NOT_Seagrasses.csv + theme(legend.position="bottom")) +  plot_annotation(tag_levels = 'A') +plot_layout(heights=c(1.5,1))
sum1
cowplot::save_plot(sum1, filename = 'output/patchwork_terrestrials_gene_loss.png', base_height = 10, base_width = 8)
```

```{r fig.width=8, fig.height=10}
sum2 <- seagrass_venn / (plot_list$GO_results_genes_union_Aquatics_and_Seagrasses_NOT_Terrestrials.csv + theme(legend.position="bottom")) +  plot_annotation(tag_levels = 'A') +plot_layout(heights=c(1.5,1))
sum2
cowplot::save_plot(sum2, filename = 'output/patchwork_seagrass_gene_loss.png', base_height = 10, base_width = 8)
```

### 2022-03-09 analysis/GOenrichment.Rmd
Add general descriptions
add more ifles

@@
 The GO enrichment does not work well on my laptop so I'm setting this to eval=FALSE and run it on a remote server.
-The script which writes the input files is a Python script in the scripts/ folder: findClustersUniqueToAquatic.py. This script parses Orthofinder output to pull out genome-specific groups and genes.
+
+This document takes the output of [Orthofinder](Orthofinder.html)
 
@@
 ```
-# Comparing the three big groups
+# Comparing the four big groups
 
@@
 
-# Genes only in aquatics
-compare('Lost_present_gene_lists/Genes_only_in_Aquatics.txt', 'output/GO_results_genes_in_Aquatics_NOT_Seagrasses_NOT_Terrestrials.csv', allgenes, annAT)
+# Genes only in algae
+compare('Lost_present_gene_lists/Genes_only_in_Algae.txt', 'output/GO_results_genes_in_Algae_NOT_Duckweeds_NOT_Seagrasses_NOT_Terrestrials.csv', allgenes, annAT)
 
 # Genes only in seagrasses
-compare('Lost_present_gene_lists/Genes_only_in_Seagrasses.txt', 'output/GO_results_genes_in_Seagrasses_NOT_Aquatics_NOT_Terrestrials.csv', allgenes, annAT)
+compare('Lost_present_gene_lists/Genes_only_in_Seagrasses.txt', 'output/GO_results_genes_in_Seagrasses_NOT_Algae_NOT_Duckweeds_NOT_Terrestrials.csv', allgenes, annAT)
+
+# Genes only in duckweeds
+compare('Lost_present_gene_lists/Genes_only_in_Duckweeds.txt', 'output/GO_results_genes_in_Duckweeds_NOT_Algae_NOT_Seagrasses_NOT_Terrestrials.csv', allgenes, annAT)
 
@@
   one.data <- suppressMessages(type_convert(one.data)) # guess data types, but also: shush
-  top_ten_values <- one.data %>% arrange(value) %>% head(10) %>% pull(term_ID) 
+  top_ten_values <- one.data %>% arrange(value) %>% head(8) %>% pull(term_ID)
+
   one.data <- one.data %>% mutate(description2 = case_when(
-    term_ID %in% top_ten_values ~ description, 
+    term_ID %in% top_ten_values ~ str_wrap(description, width=10), 
     TRUE ~ ''))
@@
     scale_size_area() +
-    geom_label_repel(max.overlaps = Inf, box.padding = 0.5,
-                     aes(point.size = log_size), alpha=0.9) +
+    geom_label_repel(max.overlaps=15,
+                     point.padding = 0,
+                     min.segment.length = 0, # draw all line segments
+                     aes(point.size=log_size), alpha=0.9) + 
     theme_minimal() +
@@
   print(i)
-  print(results_list[[i]] %>% filter(Eliminated == FALSE) %>% select(TermID, Name, Value) %>% arrange(Value) %>% kbl() %>%  kable_styling())
+  print(results_list[[i]] %>% dplyr::filter(Eliminated == FALSE) %>% dplyr::select(TermID, Name, Value) %>% arrange(Value) %>% kbl() %>%  kable_styling())
   cat("\n")
@@
 patchy
+
+save(plot_list, file = 'output/all_GO_plots.Rdata') # save so I can put them together later
+
 cowplot::save_plot('output/GO_results_terrestrials_vs_union_aquatics_seagrasses.png', patchy, base_height=7)

### 2022-03-10 analysis/GO_and_ortho.Rmd
GO-term updates! more groups! for the price of one!

@@
 library(tidyverse)
-
+library(patchwork)
 knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
@@
 ```{r fig.width=8, fig.height=10}
-sum2 <- seagrass_venn / (plot_list$GO_results_genes_union_Aquatics_and_Seagrasses_NOT_Terrestrials.csv + theme(legend.position="bottom")) +  plot_annotation(tag_levels = 'A') +plot_layout(heights=c(1.5,1))
+sum2 <- seagrass_venn / (plot_list$GO_results_genes_in_all_seagrasses_vs_seagrassesBackground_intersect.csv + theme(legend.position="bottom")) +  plot_annotation(tag_levels = 'A') +plot_layout(heights=c(1.5,1))
 sum2

### 2022-03-10 analysis/GOenrichment.Rmd
GO-term updates! more groups! for the price of one!

@@
 
+# Genes only in seagrasses- seagrass-specific
+compare('Lost_present_gene_lists/Genes_in_all_seagrasses_intersect.txt', 'output/GO_results_genes_in_all_seagrasses_vs_backgroundAll_intersect.csv', allgenes, annAT)
+
+compare('Lost_present_gene_lists/Genes_in_all_seagrasses_union.txt', 'output/GO_results_genes_in_all_seagrasses_vs_backgroundAll_union.csv', allgenes, annAT)
+
 ```
@@
 ```{r eval=FALSE}
+
 compare('Lost_present_gene_lists/Genes_lost_in_A_antarctica_not_other_seagrasses.txt', 'output/GO_results_genes_lost_A_antarctica_not_other_seagrasses.csv', sallgenes, sannAT)
@@
 ```{r eval=FALSE}
+compare('Lost_present_gene_lists/Genes_in_all_seagrasses_intersect.txt', 'output/GO_results_genes_in_all_seagrasses_vs_seagrassesBackground_intersect.csv', sallgenes, sannAT)
+
+compare('Lost_present_gene_lists/Genes_in_all_seagrasses_union.txt', 'output/GO_results_genes_in_all_seagrasses_vs_seagrassesBackground_union.csv', sallgenes, sannAT)
+
 compare('Lost_present_gene_lists/Genes_only_in_Z_marina_not_other_seagrasses.txt', 'output/GO_results_genes_only_Z_marina_not_other_seagrasses.csv', sallgenes, sannAT)

### 2022-03-10 analysis/GOenrichment.Rmd
make the setup blocks nicer

@@
 
-```{r setup, include=FALSE}
+```{r setup, message=FALSE}
 library(topGO)

### 2022-03-10 analysis/GeneLossComparisonReads.Rmd
make the setup blocks nicer

@@
 
-```{r setup}
+```{r setup, message=FALSE}
 library(tidyverse)

### 2022-03-10 analysis/plotRgenes.Rmd
nicer table
make the setup blocks nicer

--- 2022-02-21 analysis/plotRgenes.Rmd
@@
 ---
-```{r setup}
+```{r setup, message=FALSE}
 library(tidyverse)
@@
 ```
+
+# Quick summary stats
+
+```{r}
+df %>% group_by(Genome, Class) %>% summarise(sum=sum(Count)) %>% pivot_wider(names_from = c('Class'), values_from = sum) %>% kbl() %>% kable_styling()
+```

### 2022-03-10 analysis/MSA.Rmd
New: fancy MSAs!!

---
title: "Visualsing MSAs"
author: "Philipp Bayer"
date: "10 March 2022"
output: html_document
---

I found some interesting gene clusters. Let's look at them

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggmsa)
library(Biostrings)
library(ape)
library(ggtree)
library(treeio)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r fig.width=12}
ggmsa('data/OG0000189.RiceAraSeagrasses.aln.fasta', start = 221, end = 260, char_width = 0.5, seq_name = T) + geom_seqlogo() + geom_msaBar()
```

```{r}

x <- readAAStringSet('data/OG0000189.RiceAraSeagrasses.aln.fasta')
d <- as.dist(stringDist(x, method = "hamming")/width(x)[1])
tree <- bionj(d)
p <- ggtree(tree) + geom_tiplab()
ggtree(tree) + geom_tiplab()
```

```{r fig.width=10}
data = tidy_msa('data/OG0000189.RiceAraSeagrasses.aln.fasta', 221, 260)
p + geom_facet(geom = geom_msa, data = data,  panel = 'Multiple Sequence Alignment (221-260AA)',
               font = NULL, color = "Chemistry_AA") +
    xlim_tree(1)
```

Let's use a RAXML made tree

Commands run, after I shortened protein IDs manually. I pulled out the rice/Arabidopsis/seagrass proteins manually from the OG0000189.fa Orthofinder produced, and shortened their names so they fit into Phylip format.

```muscle -in OG0000189.RiceAraSeagrasses.fa -out OG0000189.RiceAraSeagrasses.phy -phyi ```

```raxmlHPC -p12345 -m PROTGAMMAAUTO -s OG0000189.RiceAraSeagrasses.phy -n AUTO```

```{r}
treestring <- '((3O.sativa:0.11386470583955717040,4O.sativa:0.03615896391829190315):0.07379737996482453599,(5O.sativa:0.33751444720146173140,((7A.thalian:0.45696796911334131019,((8Z.mueller:0.33996861614583556710,15Z.marina:0.16088168380872580610):0.31923212677234219514,10P.austra:0.57486248434538900209):0.22324965676023655892):0.10039129162924786964,((13Z.marina:0.11346816437532522559,(12Z.muelle:0.00973855109814669891,10Z.muelle:0.02645894329604922546):0.09840311662577182206):0.35332170475064506032,((14Z.marina:0.12821969530885793387,(9Z.mueller:0.06883167458475099310,11Z.muelle:0.02355162378347611801):0.15839770535961147924):0.63099939044074537797,(1A.amphibo:0.53760770644490452064,6P.austral:0.07001018412366488697):0.08839123570898912985):0.05074368132159561007):0.09141392275039983417):0.12967268161775510893):0.11320488597472283532,2O.sativa:1.70657194075952300949):0.0;'
plot(ape::read.tree(text=treestring))
p2 <- ggtree(ape::read.tree(text=treestring)) + geom_tiplab()

```

That's very different from the above dendrogram!

```{r fig.width=10}
data2 <- data %>% mutate(name = str_sub(name, 1, 10), 
                         name = str_trim(name))
p2 + geom_facet(geom = geom_msa, data = data2,  panel = 'Multiple Sequence Alignment (221-260AA)',
               font = NULL, color = "Chemistry_AA") +
    xlim_tree(2)
```

Good! I used blastp with Swissprot to see whether I could get 'official' gene names for some of these, especially the *O. sativa*. I uploaded the fasta in data/OG000189.fa to blastp/swissprot, and pulled out new names where available.

Hits are to: A. thaliana Calcineurin B-like protein 4 (CBL4, Alternative name: SOS3)
O. sativa Calcineurin B-like protein 8 (CBL8)
O. sativa Calcineurin B-like protein 4 (CBL4)

Let's rename the tree and the MSA data table using `rename_taxa`
```{r}
trees <- ape::read.tree(text=treestring)
old_names <- trees$tip.label
# [1] "3O.sativa"  "4O.sativa"  "5O.sativa"  "7A.thalian" "8Z.mueller" "15Z.marina" "10P.austra" "13Z.marina"
# [9] "12Z.muelle" "10Z.muelle" "14Z.marina" "9Z.mueller" "11Z.muelle" "1A.amphibo" "6P.austral" "2O.sativa"
new_names  <- c( 'OsCBL8', 'OsCBL7.1', 'OsCBL4', 'AtSOS3', 'ZmuSOS3', 'ZmaSOS3', 'PaSOS3', 'ZmaSOS3.2', 'ZmuSOS3.2', 'ZmuSOS3.3', 'ZmaSOS3.3', 'ZmuSOS3.2', 'ZmuSOS3.4', 'AaSOS3', 'PaSOS3', 'OsCBL7.2')
rename_df <- data.frame(old = old_names, new = new_names)
rename_df

trees <- rename_taxa(trees, rename_df, old, new)
p3 <- ggtree(trees) + geom_tiplab(fontface='italic')

p3
```

```{r}
#str_replace_all takes a named vector
replace_vector <- new_names
names(replace_vector) <- old_names
data2 <- data2 %>% mutate(name = str_replace_all(name, replace_vector))
```

```{r fig.width=10}
final_p <- p3 + geom_facet(geom = geom_msa, data = data2,  panel = 'Multiple Sequence Alignment (221-260AA)',
               font = NULL, color = "Chemistry_AA") +
    xlim_tree(2)
final_p
```

```{r}
cowplot::save_plot(final_p, filename = 'output/SOS3_phylogeny.png', base_width=10)
```

### 2022-03-10 analysis/MSA.Rmd
Add missing files
Fix one label!

@@
 # [9] "12Z.muelle" "10Z.muelle" "14Z.marina" "9Z.mueller" "11Z.muelle" "1A.amphibo" "6P.austral" "2O.sativa"
-new_names  <- c( 'OsCBL8', 'OsCBL7.1', 'OsCBL4', 'AtSOS3', 'ZmuSOS3', 'ZmaSOS3', 'PaSOS3', 'ZmaSOS3.2', 'ZmuSOS3.2', 'ZmuSOS3.3', 'ZmaSOS3.3', 'ZmuSOS3.2', 'ZmuSOS3.4', 'AaSOS3', 'PaSOS3', 'OsCBL7.2')
+new_names  <- c( 'OsCBL8', 'OsCBL7.1', 'OsCBL4', 'AtSOS3', 'ZmuSOS3', 'ZmaSOS3', 'PaSOS3', 'ZmaSOS3.2', 'ZmuSOS3.2', 'ZmuSOS3.3', 'ZmaSOS3.3', 'ZmuSOS3.2', 'ZmuSOS3.4', 'AaSOS3', 'PaSOS3.2', 'OsCBL7.2')
 rename_df <- data.frame(old = old_names, new = new_names)
@@
 ```
+
+"However, apparent photosynthesis is still maintained at a salinity 15% that of normal seawater and at temperatures of 3 and 30°C, consistent with the ecological role of Z.
+muelleri as an intertidal species." https://www.sciencedirect.com/science/article/abs/pii/0304377085900634

### 2022-03-11 analysis/GeneLossComparisonReads.Rmd
finalised MSA!

--- 2022-02-22 analysis/GeneLossComparisonReads.Rmd
@@
 stom_p
+```
+
+```{r}
+label <- stomata_long #%>% mutate(Gene.symbol = stringr::str_to_lower(Gene.symbol)) 
+label <- label %>% 
+  mutate(mylabel = case_when(Coverage >= 100 ~ '',
+                             Coverage == 0 ~ '',
+                                    TRUE ~ as.character(Coverage)))
+
+stom_heat <- stomata_long %>% 
+  #mutate(Gene.symbol = stringr::str_to_lower(Gene.symbol)) %>% 
+  ggplot(aes(x=Species, y=factor(Gene.symbol, level=rev(unique(Gene.symbol))), fill=Coverage)) + 
+  geom_tile(alpha=0.9, color='white') + 
+  #geom_text(aes(label=mylabel), data=label, size=3.5) + 
+  theme(axis.text.y = element_text(face = "italic"),
+        axis.text.x = element_text(face='italic'),
+        axis.title.x = element_blank(),
+        axis.title.y = element_blank(),
+        #legend.position = "none"
+        ) +
+  #scale_fill_continuous(limits=c(0,100)) +
+  scale_fill_distiller(palette='Spectral', limits=c(0,100))#palette= 'Rdbu', limits=c(0,100), type = 'discrete')
+
+stom_heat
 ```
@@
 ```
+
+```{r}
+eth_heat <- ethylene_long %>% 
+  dplyr::filter(Gene.symbol != 'ndhB-2') %>% 
+  #mutate(Gene.symbol = stringr::str_to_lower(Gene.symbol)) %>% 
+  ggplot(aes(x=Species, y=factor(Gene.symbol, level=rev(unique(Gene.symbol))), fill=Coverage)) + 
+  geom_tile() + 
+  ylab('Gene') +
+  theme(axis.text.y = element_text(face = "italic"),
+        axis.text.x = element_text(face='italic'),
+        axis.title.x = element_blank(),
+        axis.title.y = element_blank(),
+       #legend.position = "none"
+       ) +
+  #scale_fill_continuous(limits=c(0,100)) +
+  scale_fill_distiller(palette='Spectral', limits=c(0,100))#palette= 'Rdbu', limits=c(0,100), type = 'discrete')
+
+eth_heat
+```
+
 # NDH genes
@@
          ~Gene.symbol, ~A..antarctica, ~P..australis, ~H..ovalis, ~Z..muelleri, ~Z..marina,
-               "NDHL",              0,             0,          0,        16.67,      16.67,
-               "NDHM",              0,             0,          0,        36.24,       34.4,
-               "NDHN",          48.73,             0,          0,        66.35,      59.68,
-               "NDHO",              0,             0,          0,        33.54,      22.22,
-               "NDHS",          12.35,         28.29,          0,         34.4,      30.15,
-               "NDHT",              0,          8.67,          0,        19.47,      16.67,
-               "NDHU",              0,             0,          0,         9.89,          0,
-              "PNSB1",           4.26,         42.42,          0,         32.4,      31.24,
-              "PNSB2",              0,             0,          0,            0,          0,
-              "PNSB3",              0,          8.46,          0,        16.26,       9.76,
-              "PNSB4",              0,         19.13,          0,        18.75,      18.75,
-              "PNSL1",              0,             0,          0,            0,          0,
-              "PNSL2",          16.58,             0,          0,        45.03,      15.01,
-              "PNSL3",              0,             0,          0,            0,          0,
-              "PNSL4",           7.19,         23.55,          0,        31.04,      45.11,
-              "PNSL5",          52.82,         62.82,      57.31,        48.72,      46.28,
+               "ndhL",              0,             0,          0,        16.67,      16.67,
+               "ndhM",              0,             0,          0,        36.24,       34.4,
+               "ndhN",          48.73,             0,          0,        66.35,      59.68,
+               "ndhO",              0,             0,          0,        33.54,      22.22,
+               "ndhS",          12.35,         28.29,          0,         34.4,      30.15,
+               "ndhT",              0,          8.67,          0,        19.47,      16.67,
+               "ndhU",              0,             0,          0,         9.89,          0,
+              "PnsB1",           4.26,         42.42,          0,         32.4,      31.24,
+              "PnsB2",              0,             0,          0,            0,          0,
+              "PnsB3",              0,          8.46,          0,        16.26,       9.76,
+              "PnsB4",              0,         19.13,          0,        18.75,      18.75,
+              "PnsL1",              0,             0,          0,            0,          0,
+              "PnsL2",          16.58,             0,          0,        45.03,      15.01,
+              "PnsL3",              0,             0,          0,            0,          0,
+              "PnsL4",           7.19,         23.55,          0,        31.04,      45.11,
+              "PnsL5",          52.82,         62.82,      57.31,        48.72,      46.28,
                "CRR6",              0,             0,          0,        43.86,      33.06,
@@
 ndh_p
+```
+```{r fig.width=10}
+label <- ndh_long
+label <- label %>% 
+  mutate(mylabel = case_when(Coverage >= 100 ~ '',
+                             Coverage == 0 ~ '',
+                                    TRUE ~ as.character(Coverage)))
+
+ndh_heat <- ndh_long %>% 
+  dplyr::filter(Gene.symbol != 'ndhB-2') %>% 
+  #mutate(Gene.symbol = stringr::str_to_lower(Gene.symbol)) %>% 
+  ggplot(aes(x=Species, y=factor(Gene.symbol, level=rev(unique(Gene.symbol))), fill=Coverage)) + 
+  geom_tile() + 
+  geom_tile(alpha=0.9, color='white') + 
+  #geom_text(data=label, aes(label=mylabel), size=3.5) +
+  theme(axis.text.y = element_text(face = "italic"),
+        axis.text.x = element_text(face = 'italic'),
+        axis.title.x = element_blank(),
+        axis.title.y = element_blank(),
+        #legend.position = "none"
+        )  +
+  #scale_fill_continuous(limits=c(0,100)) +
+  scale_fill_distiller(palette='Spectral', limits=c(0,100))#palette= 'Rdbu', limits=c(0,100), type = 'discrete')
+
+ndh_heat
 ```
@@
 
+```{r}
+label <- ndh_chl_long %>% filter(Gene.symbol != 'ndhB-2')
+label <- label %>% 
+  mutate(mylabel = case_when(Coverage >= 100 ~ '',
+                             Coverage == 0 ~ '',
+                                    TRUE ~ as.character(Coverage)))
+
+ndh_chl_heat <- ndh_chl_long %>% 
+  dplyr::filter(Gene.symbol != 'ndhB-2') %>% 
+  mutate(Coverage = case_when(Coverage > 100 ~ 100,
+                              TRUE ~ Coverage)) %>% 
+  ggplot(aes(x=Species, y=factor(Gene.symbol, level=rev(unique(Gene.symbol))), fill=Coverage)) + 
+  geom_tile(alpha=0.9, color='white') + 
+  #geom_text(data=label, aes(label=mylabel), size=3.5) +
+  ylab('Gene') +
+  theme(axis.text.y = element_text(face = "italic"),
+        axis.text.x = element_text(face = 'italic'),
+        axis.title.x = element_blank(),
+        axis.title.y = element_blank(),
+        #legend.position = "none"
+        ) +
+#  scale_fill_continuous(limits=c(0,100)) +
+    scale_fill_distiller(palette='Spectral', limits=c(0,100))#palette= 'Rdbu', limits=c(0,100), type = 'discrete')
+
+ndh_chl_heat
+```
+
 # all together
@@
 ```
+
+```{r fig.height=7}
+patchwork <- (eth_p + ndh_chl_heat)/
+ (ndh_p + stom_heat  )
+#patchwork[[2]] = patchwork[[2]] + theme(axis.title.x = element_blank() )
+#patchwork[[1]] = patchwork[[1]] + theme(axis.title.x = element_blank() )
+puh <- patchwork+ plot_annotation(tag_levels = "A") + plot_layout(guides='collect')
+puh
+cowplot::save_plot(filename = 'output/patched_gene_loss.png', plot = puh, base_height = 7)
+```
+
+Hmmmmmmmmm
+
+```{r fig.width=10}
+
+patchwork <- (eth_heat + ndh_chl_heat)/
+ ( ndh_heat + stom_heat  ) & theme(axis.text = element_text(size = 8))
+#patchwork[[2]] = patchwork[[2]] + theme(axis.title.x = element_blank() )
+#patchwork[[1]] = patchwork[[1]] + theme(axis.title.x = element_blank() )
+puh <- patchwork+ plot_annotation(tag_levels = "A") + plot_layout(guides='collect')
+puh
+cowplot::save_plot(filename = 'output/patched_gene_loss_heatmaps.png', plot = puh, base_height = 8)
+```

### 2022-03-11 analysis/MSA.Rmd
finalised MSA!

@@
 ---
-title: "Visualsing MSAs"
+title: "Visualising MSAs"
 author: "Philipp Bayer"
@@
 # [9] "12Z.muelle" "10Z.muelle" "14Z.marina" "9Z.mueller" "11Z.muelle" "1A.amphibo" "6P.austral" "2O.sativa"
-new_names  <- c( 'OsCBL8', 'OsCBL7.1', 'OsCBL4', 'AtSOS3', 'ZmuSOS3', 'ZmaSOS3', 'PaSOS3', 'ZmaSOS3.2', 'ZmuSOS3.2', 'ZmuSOS3.3', 'ZmaSOS3.3', 'ZmuSOS3.2', 'ZmuSOS3.4', 'AaSOS3', 'PaSOS3.2', 'OsCBL7.2')
+new_names  <- c( 'OsCBL8', 'OsCBL7.1', 'OsCBL4', 'AtSOS3', 'ZmuSOS3', 'ZmaSOS3', 'PaSOS3', 'ZmaSOS3.2', 'ZmuSOS3.2', 'ZmuSOS3.3', 'ZmaSOS3.2', 'ZmuSOS3.4', 'ZmuSOS3.5', 'AaSOS3', 'PaSOS3.2', 'OsCBL7.2')
 rename_df <- data.frame(old = old_names, new = new_names)
@@
 muelleri as an intertidal species." https://www.sciencedirect.com/science/article/abs/pii/0304377085900634
+
+Now let's also add support values.
+
+I ran this based on `Using RAxML-NG in Practice`
+
+```
+ # to get best model
+modeltest-ng -i OG0000189.RiceAraSeagrasses.aln.fasta -t ml -d aa -p 8
+
+ # to get fixed fasta
+raxml-ng --msa OG0000189.RiceAraSeagrasses.aln.fasta --model JTT-DCMUT+G4 --check
+
+# to make a regular tree
+raxml-ng --msa OG0000189.RiceAraSeagrasses.aln.fasta.raxml.reduced.phy --model JTT-DCMUT+G4 --prefix T3 --threads 2 --seed 2
+
+# make 200 bootstrap trees, does not converge
+raxml-ng --msa OG0000189.RiceAraSeagrasses.aln.fasta.raxml.reduced.phy --model JTT-DCMUT+G4 --prefix T8 --threads 8 --seed 2 --bootstrap --bs-trees 200
+
+# make another 400  with different seed
+raxml-ng --msa OG0000189.RiceAraSeagrasses.aln.fasta.raxml.reduced.phy --model JTT-DCMUT+G4 --prefix T11 --threads 8 --seed 333 --bootstrap --bs-trees 400
+
+# check whether they converge with <3% WRF cutoff
+raxml-ng --bsconverge --bs-trees allbootstraps --prefix T12 --seed 2 --threads 1 --bs-cutoff 0.03
+
+#yes, after 550 - close one!
+
+# to make the final trees with bootstrap values
+
+cat T8.raxml.bootstraps T11.raxml.bootstraps > allbootstraps
+
+raxml-ng --support --tree T3.raxml.bestTree --bs-trees allbootstraps --prefix T13 --redo
+```
+
+```{r}
+newtree <- '((((((12Z.muelleri_maker-6084_47488_1_44441--0.5-mRNA-1:0.009685,10Z.muelleri_maker-8016_47756--0.9-mRNA-1:0.026505)99:0.098293,13Z.marina_Zosma01g02330:0.113268)99:0.354433,(((11Z.muelleri_maker-8016_47756--0.8-mRNA-1:0.023273,9Z.muelleri_maker-2123_50249_30746_38748--0.4-mRNA-1:0.068317)99:0.153634,14Z.marina_Zosma01g02340:0.129762)100:0.616637,(1A.amphibolis_maker-scf7180000542070-augustus-gene-0.0-mRNA-1:0.518908,6P.australis_maker-scf7180003728898-augustus-gene-0.1-mRNA-1:0.074916)72:0.082830)22:0.047398)31:0.080205,((2O.sativa_LOC_Os03g33570.1:1.647537,(3O.sativa_LOC_Os02g18930.1:0.113637,4O.sativa_LOC_Os02g18880.1:0.035299)53:0.061341)66:0.111396,5O.sativa_LOC_Os05g45810.1:0.344545)75:0.137407)42:0.100482,7A.thaliana_AT5G24270.1:0.448151)83:0.209634,(15Z.marina_Zosma06g25260:0.152988,8Z.muelleri_snap_masked-14205_19045--0.3-mRNA-1:0.336420)100:0.314204,10P.australis_maker-scf7180004080271-augustus-gene-0.8-mRNA-1:0.580458)83:0.0;'
+
+tree3 <- ape::read.tree(text=newtree, branch.label='support')
+
+p3 <- ggtree(tree3) + geom_tiplab()
+p3
+```
+
+OK the nodes are now named differently due to raxml-ng, time to fix again
+
+```{r}
+old_names <- tree3$tip.label
+
+# [1] "3O.sativa"  "4O.sativa"  "5O.sativa"  "7A.thalian" "8Z.mueller" "15Z.marina" "10P.austra" "13Z.marina"
+# [9] "12Z.muelle" "10Z.muelle" "14Z.marina" "9Z.mueller" "11Z.muelle" "1A.amphibo" "6P.austral" "2O.sativa"
+#new_names  <- c( 'OsCBL8', 'OsCBL7.1', 'OsCBL4', 'AtSOS3', 'ZmuSOS3', 'ZmaSOS3', 'PaSOS3', 'ZmaSOS3.2',
+#                 'ZmuSOS3.2', 'ZmuSOS3.3', 'ZmaSOS3.3', 'ZmuSOS3.2', 'ZmuSOS3.4', 'AaSOS3', 'PaSOS3.2', 'OsCBL7.2')
+new_names  <- c('ZmuSOS3.2', 'ZmuSOS3.3', 'ZmaSOS3.2', 'ZmuSOS3.4', 'ZmuSOS3.5', 'ZmaSOS3.2','AaSOS3',  
+                'PaSOS3.2', 'OsCBL7.2', 'OsCBL8', 'OsCBL7.1', 'OsCBL4', 'AtSOS3', 'ZmaSOS3', 'ZmuSOS3', 'PaSOS3')
+rename_df <- data.frame(old = old_names, new = new_names)
+
+tree3 <- rename_taxa(tree3, rename_df, old, new)
+p3 <- ggtree(tree3) + geom_tiplab(fontface='italic')
+p3 <- p3 + geom_nodelab(aes(subset=label>80))
+```
+
+```{r fig.width = 10}
+replace_vector <- new_names
+names(replace_vector) <- old_names
+
+data3 <- data2 %>% mutate(name = str_replace(name, ' ', '_'),
+                          name = str_replace_all(name, replace_vector))
+p3 + geom_facet(geom = geom_msa, data = data3,  panel = 'Multiple Sequence Alignment (221-260AA)',
+               font = NULL, color = "Chemistry_AA") +
+    xlim_tree(3)
+```

### 2022-03-11 analysis/MSA.Rmd
fix MSA

@@
 ```{r fig.width=10}
-data = tidy_msa('data/OG0000189.RiceAraSeagrasses.aln.fasta', 221, 260)
+data <- tidy_msa('data/OG0000189.RiceAraSeagrasses.aln.fasta', 221, 260)
 p + geom_facet(geom = geom_msa, data = data,  panel = 'Multiple Sequence Alignment (221-260AA)',
@@
 #                 'ZmuSOS3.2', 'ZmuSOS3.3', 'ZmaSOS3.3', 'ZmuSOS3.2', 'ZmuSOS3.4', 'AaSOS3', 'PaSOS3.2', 'OsCBL7.2')
-new_names  <- c('ZmuSOS3.2', 'ZmuSOS3.3', 'ZmaSOS3.2', 'ZmuSOS3.4', 'ZmuSOS3.5', 'ZmaSOS3.2','AaSOS3',  
+new_names  <- c('ZmuSOS3.2', 'ZmuSOS3.3', 'ZmaSOS3.2', 'ZmuSOS3.4', 'ZmuSOS3.5', 'ZmaSOS3.3','AaSOS3',  
                 'PaSOS3.2', 'OsCBL7.2', 'OsCBL8', 'OsCBL7.1', 'OsCBL4', 'AtSOS3', 'ZmaSOS3', 'ZmuSOS3', 'PaSOS3')
@@
 p3 <- p3 + geom_nodelab(aes(subset=label>80))
+p3
 ```
@@
 
-data3 <- data2 %>% mutate(name = str_replace(name, ' ', '_'),
+data <- tidy_msa('data/OG0000189.RiceAraSeagrasses.aln.fasta', 221, 260)
+data3 <- data %>% mutate(name = str_replace(name, ' ', '_'),
                           name = str_replace_all(name, replace_vector))

### 2022-03-15 analysis/ThreeDStructure.Rmd
wflow_publish("analysis/ThreeDStructure.Rmd")

---
title: "Visualising 3D structures"
author: "Philipp Bayer"
date: "15 March 2022"
output: html_document
---

Let's compare some 3D structures from the interesting genes in [MSA](MSA.html)

```{r message=FALSE, warning=FALSE}
library(bio3d)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

```

# SOS3

First, let's just load one of them 
```{r}
# the 'real' one from X-ray crystallography
ara_SOS3_real <- read.pdb('data/SOS3_OG0000189/Arabidopsis_SOS3_1v1f.pdb')
# the alphafold ones

amphibolis_SOS3_alpha <- read.pdb('data/SOS3_OG0000189/Amphibolis_fixed_SOS3_40588.result/Amphibolis_fixed_SOS3_40588_unrelaxed_rank_1_model_4.pdb')
posi_SOS3_alpha <- read.pdb('data/SOS3_OG0000189/6P_australis_SOS3_dcfd4.result/6P_australis_SOS3_dcfd4_unrelaxed_rank_1_model_1.pdb')
```

```{r}
ara_SOS3_gaps <- gap.inspect(ara_SOS3_real$xyz)
ara.xyz <- ara_SOS3_real$xyz[ara_SOS3_gaps$f.inds]

amphi_gaps <- gap.inspect(amphibolis_SOS3_alpha$xyz)
amph.xyz <- amphibolis_SOS3_alpha$xyz[amphi_gaps$f.inds]

a <- torsion.xyz(ara.xyz, atm.inc=1)
b <- torsion.xyz(amph.xyz, atm.inc=1)
d.ab <- wrap.tor(a-b)
d.ab[is.na(d.ab)] <- 0
plot.bio3d(abs(d.ab), typ="h", xlab="Residue No.", 
           ylab = "Difference Angle")

```
```{r}
posi_SOS3_gaphs <- gap.inspect(posi_SOS3_alpha$xyz)
posi.xyz <- posi_SOS3_alpha$xyz[posi_SOS3_gaphs$f.inds]

a <- torsion.xyz(posi.xyz, atm.inc=1)
b <- torsion.xyz(amph.xyz, atm.inc=1)
d.ab <- wrap.tor(a-b)
d.ab[is.na(d.ab)] <- 0
plot.bio3d(abs(d.ab), typ="h", xlab="Residue No.", 
           ylab = "Difference Angle")

```

```{r}
library(r3dmol)

m_glimpse(ara_SOS3_real, outline = FALSE) %>% 
  m_spin()
```

```{r}

m_glimpse(posi_SOS3_alpha, outline = FALSE) %>% 
  m_spin()
```

```{r}
m1 <- r3dmol() %>%
  m_add_model(data = posi_SOS3_alpha, format = "pdb") %>%
  m_zoom_to()
m2 <- r3dmol() %>%
  m_add_model(data = amphibolis_SOS3_alpha, format = "pdb") %>%
  m_zoom_to()
m3 <- r3dmol() %>%
  m_add_model(data = '1v1f', format = "pdb") %>%
  m_zoom_to()
m_grid(
  viewer = list(m1, m2, m3),
  rows = 2,
  cols = 2,
  #control_all = TRUE,
  #viewer_config = m_viewer_spec(
  #  backgroundColor = "lightblue"
  #)
)
```

### 2022-03-16 analysis/MSA.Rmd
fix MSA code to be nicer

@@
 ```{r fig.width=12}
-ggmsa('data/OG0000189.RiceAraSeagrasses.aln.fasta', start = 221, end = 260, char_width = 0.5, seq_name = T) + geom_seqlogo() + geom_msaBar()
+ggmsa('data/SOS3_OG0000189/OG0000189.RiceAraSeagrasses.aln.fasta', start = 221, end = 260, char_width = 0.5, seq_name = T) + geom_seqlogo() + geom_msaBar()
 ```
@@
 
-x <- readAAStringSet('data/OG0000189.RiceAraSeagrasses.aln.fasta')
+x <- readAAStringSet('data/SOS3_OG0000189/OG0000189.RiceAraSeagrasses.aln.fasta')
 d <- as.dist(stringDist(x, method = "hamming")/width(x)[1])
@@
 ```{r fig.width=10}
-data <- tidy_msa('data/OG0000189.RiceAraSeagrasses.aln.fasta', 221, 260)
+data <- tidy_msa('data/SOS3_OG0000189/OG0000189.RiceAraSeagrasses.aln.fasta', 221, 260)
 p + geom_facet(geom = geom_msa, data = data,  panel = 'Multiple Sequence Alignment (221-260AA)',
@@
 p2 <- ggtree(ape::read.tree(text=treestring)) + geom_tiplab()
-
+p2
 ```
@@
 trees <- ape::read.tree(text=treestring)
-old_names <- trees$tip.label
-# [1] "3O.sativa"  "4O.sativa"  "5O.sativa"  "7A.thalian" "8Z.mueller" "15Z.marina" "10P.austra" "13Z.marina"
-# [9] "12Z.muelle" "10Z.muelle" "14Z.marina" "9Z.mueller" "11Z.muelle" "1A.amphibo" "6P.austral" "2O.sativa"
-new_names  <- c( 'OsCBL8', 'OsCBL7.1', 'OsCBL4', 'AtSOS3', 'ZmuSOS3', 'ZmaSOS3', 'PaSOS3', 'ZmaSOS3.2', 'ZmuSOS3.2', 'ZmuSOS3.3', 'ZmaSOS3.2', 'ZmuSOS3.4', 'ZmuSOS3.5', 'AaSOS3', 'PaSOS3.2', 'OsCBL7.2')
-rename_df <- data.frame(old = old_names, new = new_names)
-rename_df
+
+rename_df <- tibble::tribble(
+                       ~old,        ~new,
+                "3O.sativa",    "OsCBL8",
+                "4O.sativa",  "OsCBL7.1",
+                "5O.sativa",    "OsCBL4",
+               "7A.thalian",    "AtSOS3",
+               "8Z.mueller",   "ZmuSOS3",
+               "15Z.marina",   "ZmaSOS3",
+               "10P.austra",    "PaSOS3",
+               "13Z.marina", "ZmaSOS3.2",
+               "12Z.muelle", "ZmuSOS3.2",
+               "10Z.muelle", "ZmuSOS3.3",
+               "14Z.marina", "ZmaSOS3.2",
+               "9Z.mueller", "ZmuSOS3.4",
+               "11Z.muelle", "ZmuSOS3.5",
+               "1A.amphibo",    "AaSOS3",
+               "6P.austral",  "PaSOS3.2",
+                "2O.sativa",  "OsCBL7.2"
+               )
 
@@
 #str_replace_all takes a named vector
-replace_vector <- new_names
-names(replace_vector) <- old_names
+replace_vector <- rename_df$new
+names(replace_vector) <- rename_df$old
 data2 <- data2 %>% mutate(name = str_replace_all(name, replace_vector))
@@
 
+## Phylogeny for the SOS3 cluster
 ```
+
  # to get best model
@@
 
-#yes, after 550 - close one!
-
 # to make the final trees with bootstrap values
@@
 
-raxml-ng --support --tree T3.raxml.bestTree --bs-trees allbootstraps --prefix T13 --redo
+raxml-ng --support --tree T3.raxml.bestTree --bs-trees allbootstraps --prefix T13
+
+#yes, after 550 - close one!
+
+# let's make another 400 so we can have a nice 1000 bootstraps - different seed again!
+
+raxml-ng --msa OG0000189.RiceAraSeagrasses.aln.fasta.raxml.reduced.phy --model JTT-DCMUT+G4 --prefix T14 --threads 8 --seed 1234 --bs-trees 400  --bootstrap
+
+cat T8.raxml.bootstraps T11.raxml.bootstraps T14.raxml.bootstraps > allbootstraps
+raxml-ng --support --tree T3.raxml.bestTree --bs-trees allbootstraps --prefix T15
+
 ```
@@
 
-data <- tidy_msa('data/OG0000189.RiceAraSeagrasses.aln.fasta', 221, 260)
+data <- tidy_msa('data/SOS3_OG0000189/OG0000189.RiceAraSeagrasses.aln.fasta', 221, 260)
 data3 <- data %>% mutate(name = str_replace(name, ' ', '_'),
@@
 ```
+## Phylogeny for the EIN3 cluster
+
+First, to align the sequences again:
+
+```
+modeltest-ng reports JTT+I+G4+F
+
+raxml-ng --msa OG0000629.RiceAraSeagrasses.aln.fa --model JTT+I+G4+F
+
+ # to get fixed fasta - all good!!
+raxml-ng --msa OG0000629.RiceAraSeagrasses.aln.fa --model JTT+I+G4+F --check
+
+# to make a regular tree
+raxml-ng --msa OG0000629.RiceAraSeagrasses.aln.fa --model JTT+I+G4+F --prefix T3 --threads 2 --seed 2
+
+# make 1000 bootstrap trees
+raxml-ng --msa OG0000629.RiceAraSeagrasses.aln.fa --model JTT+I+G4+F --prefix T8 --threads 16 --seed 2 --bootstrap --bs-trees 1000
+
+# check whether they converge with <3% WRF cutoff
+raxml-ng --bsconverge --bs-trees T8.raxml.bootstraps --prefix T12 --seed 2 --threads 1 --bs-cutoff 0.03
+
+# it actually converged with 50 trees. haha.
+
+# to make the final trees with bootstrap values
+
+raxml-ng --support --tree T3.raxml.bestTree --bs-trees allbootstraps --prefix T13 
+```
+
+There we go:
+
+```{r}
+tree3 <- ape::read.tree(text='(((LOC_Os07g48630.1:0.641123,(LOC_Os03g20780.1:0.007239,LOC_Os03g20790.1:0.001048)100:0.214718)100:0.418001,(((Zosma02g24210:0.119273,(maker-1834_92013--0.52-mRNA-1:0.074816,augustus_masked-5412_21708--0.0-mRNA-1:0.082668)97:0.073791)100:0.640955,augustus_masked-scf7180003860583-processed-gene-0.3-mRNA-1:0.182720)89:0.114788,(maker-3356_51779--0.18-mRNA-1:0.661393,Zosma01g42670:0.320772)100:0.468870)67:0.098845)100:0.660268,AT3G20770.1:0.149857,AT2G27050.1:0.250165)100:0.0;', branch.label='support')
+
+rename_df <- tibble::tribble(
+                                                       ~old,        ~new,
+                                         "LOC_Os07g48630.1",    "OsEIL2",
+                                         "LOC_Os03g20780.1",    "OsEIL1",
+                                         "LOC_Os03g20790.1",  "OsEIL1.2",
+                                            "Zosma02g24210", "ZmaEIN3.1",
+                            "maker-1834_92013--0.52-mRNA-1", "ZmuEIN3.1",
+                   "augustus_masked-5412_21708--0.0-mRNA-1", "ZmuEIN3.2",
+"augustus_masked-scf7180003860583-processed-gene-0.3-mRNA-1",   "PaEIN3",
+                            "maker-3356_51779--0.18-mRNA-1", "ZmuEIN3.3",
+                                            "Zosma01g42670", "ZmaEIN3.2",
+                                              "AT3G20770.1",    "AtEIN3",
+                                              "AT2G27050.1",    "AtEIL1"
+               )
+
+tree3 <- rename_taxa(tree3, rename_df, old, new)
+
+p3 <-  ggtree(tree3) + geom_tiplab(fontface='italic')
+p3 <- p3 + geom_nodelab(aes(subset=label>80))
+p3 + xlim_tree(2.1)
+```

### 2022-03-16 analysis/ThreeDStructure.Rmd
workflowr:::wflow_publish(files = "analysis/ThreeDStructure.Rmd")

@@
 # the 'real' one from X-ray crystallography
-ara_SOS3_real <- read.pdb('data/SOS3_OG0000189/Arabidopsis_SOS3_1v1f.pdb')
+ara_SOS3_real <- read.pdb('data//SOS3_OG0000189/Arabidopsis_SOS3_1v1f.pdb')
 # the alphafold ones
 
-amphibolis_SOS3_alpha <- read.pdb('data/SOS3_OG0000189/Amphibolis_fixed_SOS3_40588.result/Amphibolis_fixed_SOS3_40588_unrelaxed_rank_1_model_4.pdb')
-posi_SOS3_alpha <- read.pdb('data/SOS3_OG0000189/6P_australis_SOS3_dcfd4.result/6P_australis_SOS3_dcfd4_unrelaxed_rank_1_model_1.pdb')
+amphibolis_SOS3_alpha <- read.pdb('data//SOS3_OG0000189/Amphibolis_fixed_SOS3_40588.result/Amphibolis_fixed_SOS3_40588_unrelaxed_rank_1_model_4.pdb')
+posi_SOS3_alpha <- read.pdb('data//SOS3_OG0000189/6P_australis_SOS3_dcfd4.result/6P_australis_SOS3_dcfd4_unrelaxed_rank_1_model_1.pdb')
+rice_SOS3_alpha <- read.pdb('data//SOS3_OG0000189/LOC_Os05g45810.1_colabfold_test_77dbc.result/test_77dbc_unrelaxed_rank_1_model_4.pdb')
 ```
@@
 ```{r}
+m_glimpse(rice_SOS3_alpha, outline = FALSE) %>% 
+  m_spin()
+```
 
+```{r}
 m_glimpse(posi_SOS3_alpha, outline = FALSE) %>% 
@@
 m1 <- r3dmol() %>%
-  m_add_model(data = posi_SOS3_alpha, format = "pdb") %>%
-  m_zoom_to()
+  m_add_model(data = 'data//SOS3_OG0000189/6P_australis_SOS3_dcfd4.result/6P_australis_SOS3_dcfd4_unrelaxed_rank_1_model_1.pdb', format = "pdb") %>% 
+    m_zoom_to() %>% 
+    m_set_style(style = m_style_cartoon(color = "spectrum"))
+
 m2 <- r3dmol() %>%
-  m_add_model(data = amphibolis_SOS3_alpha, format = "pdb") %>%
-  m_zoom_to()
+  m_add_model(data = 'data//SOS3_OG0000189/Amphibolis_fixed_SOS3_40588.result/Amphibolis_fixed_SOS3_40588_unrelaxed_rank_1_model_4.pdb', format = "pdb") %>%
+  m_zoom_to() %>% 
+  m_set_style(style = m_style_cartoon(color = "spectrum"))
+
 m3 <- r3dmol() %>%
-  m_add_model(data = '1v1f', format = "pdb") %>%
-  m_zoom_to()
+  m_add_model(data = m_fetch_pdb('1v1f'), format = "pdb") %>%
+  m_zoom_to() %>% 
+  m_set_style(style = m_style_cartoon(color = "spectrum"))
+
+m4 <- r3dmol() %>%
+    m_add_model(data = "data//SOS3_OG0000189/LOC_Os05g45810.1_colabfold_test_77dbc.result/test_77dbc_unrelaxed_rank_1_model_4.pdb", format = "pdb") %>%
+    m_zoom_to() %>% 
+  m_set_style(style = m_style_cartoon(color = "spectrum"))
+
 m_grid(
-  viewer = list(m1, m2, m3),
+  viewer = list(m1, m2, m3, m4),
   rows = 2,
   cols = 2,
-  #control_all = TRUE,
-  #viewer_config = m_viewer_spec(
-  #  backgroundColor = "lightblue"
-  #)
+  control_all = TRUE,
+  viewer_config = m_viewer_spec(
+    backgroundColor = "lightblue"
+  )
 )
 ```
+
+Let's compare these properly
+
+```{r}
+files <- c('data//SOS3_OG0000189/Amphibolis_fixed_SOS3_40588.result/Amphibolis_fixed_SOS3_40588_unrelaxed_rank_1_model_4.pdb','data//SOS3_OG0000189/6P_australis_SOS3_dcfd4.result/6P_australis_SOS3_dcfd4_unrelaxed_rank_1_model_1.pdb')
+pdbs <- pdbaln(files, exefile='msa')
+#xyz <- pdbfit( posi_SOS3_alpha, rice_SOS3_alpha, core.inds )
+```
+
+```{r}
+core <- core.find(pdbs)
+col=rep("black", length(core$volume))
+col[core$volume<2]="pink"; col[core$volume<1]="red"
+plot(core, col=col)
+core.inds <- print(core, vol=1.0)
+```
+
+```{r}
+xyz <- pdbfit( pdbs, core.inds, outpath = 'output/quick_fit.pdb' )
+rd <- rmsd(xyz)
+hist(rd, breaks=40, xlab="RMSD (Å)", main="Histogram of RMSD")
+```
+
+```{r}
+gaps.xyz2 <- gap.inspect(pdbs$xyz[c(1,2), ])
+a.xyz <- pdbs$xyz[1, gaps.xyz2$f.inds]
+b.xyz <- pdbs$xyz[2, gaps.xyz2$f.inds]
+a <- torsion.xyz(a.xyz, atm.inc=1)
+b <- torsion.xyz(b.xyz, atm.inc=1)
+d.ab <- wrap.tor(a-b)
+d.ab[is.na(d.ab)] <- 0
+plot.bio3d(abs(d.ab), typ="h", xlab="Residue No.", 
+           ylab = "Difference Angle")
+```
+
+MUCH better.
+```{r}
+a <- dm.xyz(a.xyz)
+b <- dm.xyz(b.xyz)
+
+plot.dmat( (a - b), nlevels=10, grid.col="gray", xlab="1tag", ylab="1tnd")
+```
+
+```{r}
+gaps.pos <- gap.inspect(pdbs$xyz)
+pc.xray <- pca.xyz(xyz[, gaps.pos$f.inds])
+pc.xray
+plot(pc.xray)
+```

### 2022-03-23 analysis/MSA.Rmd
workflowr::wflow_publish(files = "analysis/MSA.Rmd")

@@
 ```{r fig.width=10}
-data <- tidy_msa('data/SOS3_OG0000189/OG0000189.RiceAraSeagrasses.aln.fasta', 221, 260)
-p + geom_facet(geom = geom_msa, data = data,  panel = 'Multiple Sequence Alignment (221-260AA)',
-               font = NULL, color = "Chemistry_AA") +
+data <- tidy_msa('data/SOS3_OG0000189/OG0000189.RiceAraSeagrasses.aln.fasta', 221, 300)
+p + geom_facet(geom = geom_msa, data = data,  panel = 'Multiple Sequence Alignment (221-300AA)',
+               font = NULL, color = "Clustal", by_conversation= TRUE) +
     xlim_tree(1)
@@
 p2 + geom_facet(geom = geom_msa, data = data2,  panel = 'Multiple Sequence Alignment (221-260AA)',
-               font = NULL, color = "Chemistry_AA") +
+               font = NULL, color = "Clustal") +
     xlim_tree(2)
@@
 ```
+
+```{r fig.width=10}
+msaplot(p3, 'data/SOS3_OG0000189/OG0000189.RiceAraSeagrasses.aln.NamesFixed.fasta',
+        offset = 0.5) + theme(legend.position='none')
+```
+
 ## Phylogeny for the EIN3 cluster
@@
 ```
+
+```{r}
+msaplot(p3, './data/EIN3_OG0000629/OG0000629.RiceAraSeagrasses.aln.NamesFixed.fa',
+        offset = 0.5)  + theme(legend.position='none')
+
+```
+
+```{r}
+domains <- tibble::tribble(
+                                                                    ~Gene,     ~Pfam, ~Start, ~End,   ~Evalue,
+                                                          "Zosma01g42670", "PF04873",     3L, 246L,  2.9e-117,
+                                                            "AT2G27050.1", "PF04873",    49L, 298L,  2.6e-130,
+                                                       "LOC_Os03g20780.1", "PF04873",    78L, 332L,  1.4e-132,
+             "augustus_masked-scf7180003860583-processed-gene-0.3-mRNA-1", "PF04873",    76L, 325L,  1.9e-131,
+                                          "maker-1834_92013--0.52-mRNA-1", "PF04873",   136L, 408L,    2e-123,
+                                                       "LOC_Os07g48630.1", "PF04873",    81L, 338L,  3.7e-129,
+                                                            "AT3G20770.1", "PF04873",    49L, 296L,  4.8e-129,
+                                 "augustus_masked-5412_21708--0.0-mRNA-1", "PF04873",    16L, 285L,  2.2e-123,
+                                          "maker-3356_51779--0.18-mRNA-1", "PF04873",    85L, 125L,     2e-12,
+                                                       "LOC_Os03g20790.1", "PF04873",    78L, 332L,  1.4e-132,
+                                                          "Zosma02g24210", "PF04873",     1L, 206L,  1.9e-113
+             )
+domains$length <- domains$End - domains$Start
+replace_vector <- rename_df$new
+names(replace_vector) <- rename_df$old
+domains <- domains %>% mutate(Gene = str_replace_all(Gene, replace_vector))
+domains
+```
+
+```{r}
+g <- domains %>% left_join(rename_df, by=c('Gene'='new')) %>%  ggplot(aes(y = Gene, yend=Gene, x=Start,  xend=End,  color=Evalue)) + geom_segment(size=2) + theme_minimal() + xlab(expression(paste("Position of ", italic("EIN3"))))
+g
+```
+
+```{r fig.width=10}
+library(aplot)
+g <- g + theme(axis.title.y=element_blank(),
+               axis.text.y = element_blank(),
+                legend.position='none')
+
+g %>% insert_left(p3, width=2.5)
+```
+
+```{r fig.width=10}
+library(patchwork)
+# let's make a patchwork
+# PROBLEM: patchwork does not automatically sort, so we'll have to sort the plot `g` manually
+
+plot_order <- p3[['data']] %>% filter(isTip == TRUE) %>% select(label, y) %>% arrange(desc(y))
+g <- domains %>% left_join(rename_df, by=c('Gene'='new')) %>%  ggplot(aes(y = factor(Gene, levels = plot_order$label), yend=Gene, x=Start,  xend=End,  color=Evalue)) + geom_segment(size=2) + theme_minimal() + xlab(expression(paste("Position of ", italic("EIN3"))))+ theme(axis.title.y=element_blank(),
+               axis.text.y = element_blank(),
+                legend.position='none')
+g
+
+msaplot(p3, './data/EIN3_OG0000629/OG0000629.RiceAraSeagrasses.aln.NamesFixed.fa',
+        offset = 0.5)  + theme(legend.position='none') + g  + 
+  plot_layout(widths = c(3.5, 1)) 
+```

### 2022-03-23 analysis/ThreeDStructure.Rmd
add missing files
workflowr::wflow_publish(files = "analysis/ThreeDStructure.Rmd")

@@
 rice_SOS3_alpha <- read.pdb('data//SOS3_OG0000189/LOC_Os05g45810.1_colabfold_test_77dbc.result/test_77dbc_unrelaxed_rank_1_model_4.pdb')
-```
-
-```{r}
-ara_SOS3_gaps <- gap.inspect(ara_SOS3_real$xyz)
-ara.xyz <- ara_SOS3_real$xyz[ara_SOS3_gaps$f.inds]
-
-amphi_gaps <- gap.inspect(amphibolis_SOS3_alpha$xyz)
-amph.xyz <- amphibolis_SOS3_alpha$xyz[amphi_gaps$f.inds]
-
-a <- torsion.xyz(ara.xyz, atm.inc=1)
-b <- torsion.xyz(amph.xyz, atm.inc=1)
-d.ab <- wrap.tor(a-b)
-d.ab[is.na(d.ab)] <- 0
-plot.bio3d(abs(d.ab), typ="h", xlab="Residue No.", 
-           ylab = "Difference Angle")
-
-```
-```{r}
-posi_SOS3_gaphs <- gap.inspect(posi_SOS3_alpha$xyz)
-posi.xyz <- posi_SOS3_alpha$xyz[posi_SOS3_gaphs$f.inds]
-
-a <- torsion.xyz(posi.xyz, atm.inc=1)
-b <- torsion.xyz(amph.xyz, atm.inc=1)
-d.ab <- wrap.tor(a-b)
-d.ab[is.na(d.ab)] <- 0
-plot.bio3d(abs(d.ab), typ="h", xlab="Residue No.", 
-           ylab = "Difference Angle")
-
 ```
@@
 ```
+
+OK cool - now we can align all proteins and make proper plots!

### 2022-04-26 analysis/Orthofinder.Rmd
start to remove gibba

@@
 australis')
+```
+
+Remove L. gibba
+
+```{r}
+groups$`L. gibba` <- NULL
 ```
@@
 s_polyrhiza_groups <- groups %>%  select(Orthogroup, `S. polyrhiza`) %>% filter(!is.na(`S. polyrhiza`)) %>% separate_rows(Orthogroup, `S. polyrhiza`, convert=T, sep = ', ')
-l_gibba_groups <- groups %>%  select(Orthogroup, `L. gibba`) %>% filter(!is.na(`L. gibba`)) %>% separate_rows(Orthogroup, `L. gibba`, convert=T, sep = ', ')
+#l_gibba_groups <- groups %>%  select(Orthogroup, `L. gibba`) %>% filter(!is.na(`L. gibba`)) %>% separate_rows(Orthogroup, `L. gibba`, convert=T, sep = ', ')
 rice_groups<- groups %>%  select(Orthogroup, `O. sativa`) %>% filter(!is.na(`O. sativa`)) %>% separate_rows(Orthogroup, `O. sativa`, convert=T, sep = ', ')
@@
                               TRUE ~ F),
-  present_in_lgibba = case_when(Orthogroup %in% unique(l_gibba_groups$Orthogroup) ~ T,
-                              TRUE ~ F),
+ # present_in_lgibba = case_when(Orthogroup %in% unique(l_gibba_groups$Orthogroup) ~ T,
+#                              TRUE ~ F),
   present_in_rice = case_when(Orthogroup %in% unique(rice_groups$Orthogroup) ~ T,
@@
 big_joined <- big_joined %>% mutate(Lost_in_Seagrasses = 4 - (present_in_posi + present_in_zmuel + present_in_zmar + present_in_amphi),
-                      Lost_in_duckweeds=3 - (present_in_lgibba + present_in_spolyrhiza + present_in_waustralis),
+                      #Lost_in_duckweeds=3 - (present_in_lgibba + present_in_spolyrhiza + present_in_waustralis),
+                      Lost_in_duckweeds=2 - (present_in_spolyrhiza + present_in_waustralis),
+
                       Lost_in_both = (Lost_in_Seagrasses + Lost_in_duckweeds))
@@
 zmuel_count <- zmuel_groups %>% count(Orthogroup)
-lgibba_count <- l_gibba_groups %>% count(Orthogroup)
+#lgibba_count <- l_gibba_groups %>% count(Orthogroup)
 s_polyrhi_count <- s_polyrhiza_groups %>% count(Orthogroup)
@@
 ```{r}
-counts <- plyr::join_all(list(ara_count, posi_count, amphi_count, zmar_count, zmuel_count, lgibba_count, s_polyrhi_count, w_aus_count, rice_count), by='Orthogroup', type='left') 
-names(counts) <- c('Orthogroup', 'A. thaliana', 'P. australis', 'A. antarctica', 'Z. marina', 'Z. muelleri', 'L. gibba', 'S. polyrhiza', 'W. australis', 'O. sativa')
+counts <- plyr::join_all(list(ara_count, posi_count, amphi_count, zmar_count, zmuel_count, 
+                              #lgibba_count, 
+                              s_polyrhi_count, w_aus_count, rice_count), by='Orthogroup', type='left') 
+names(counts) <- c('Orthogroup', 'A. thaliana', 'P. australis', 'A. antarctica', 'Z. marina', 'Z. muelleri', 
+                   #'L. gibba', 
+                   'S. polyrhiza', 'W. australis', 'O. sativa')
 head(counts)
@@
     name %in% c(
-      'L. gibba',
+      #'L. gibba',
       'S. polyrhiza',
@@
     name %in% c(
-      'L. gibba',
+     # 'L. gibba',
       'S. polyrhiza',

### 2022-04-27 analysis/MSA.Rmd
Better phylogenies

@@
 p3 <- ggtree(tree3) + geom_tiplab(fontface='italic')
-p3 <- p3 + geom_nodelab(aes(subset=label>80))
-p3
+p3 <- p3 + geom_nodelab(aes(subset=label>80), label='*')
+p3+ ggplot2::xlim(0,2.5)
+
+```
+
+```{r}
+muhtree <- ggtree(tree3)
+my_data <- muhtree$data
+my_data <- my_data %>% mutate(support = replace_na(as.numeric(label), 0))
+root <- rootnode(tree3)
+
+good_p <- ggtree(tree3)  %<+% my_data  +
+  geom_tiplab(fontface = 'italic') +
+  geom_point2(aes(
+    subset = !isTip & node != root,
+    fill = cut(support, c(-1, 70, 90, 100))
+  ),
+  shape = 21,
+  size = 4) +
+  scale_fill_manual(
+    values = c('#d95f02', '#7570b3', '#1b9e77'),
+    labels = expression(BP < 70, 70 <= BP * " < 90", BP >= 90),
+    name = 'Bootstrap Percentage',
+    breaks = c('(-1,70]', '(70,90]', '(90,100]')
+  ) +
+  ggplot2::xlim(0, 2.5) +
+  theme_tree(legend.position = 'bottom')
+good_p
+```
+
+```{r}
+cowplot::save_plot(good_p, filename = 'output/SOS3_better_phylogeny.png')
 ```
@@
 p3 <-  ggtree(tree3) + geom_tiplab(fontface='italic')
-p3 <- p3 + geom_nodelab(aes(subset=label>80))
+p3 <- p3 + geom_nodelab()
 p3 + xlim_tree(2.1)
+```
+
+```{r}
+muhtree <- ggtree(tree3)
+my_data <- muhtree$data
+my_data <- my_data %>% mutate(support = replace_na(as.numeric(label), 0))
+root <- rootnode(tree3)
+
+good_p <- ggtree(tree3)  %<+% my_data  +
+  geom_tiplab(fontface = 'italic') +
+  geom_point2(aes(
+    subset = !isTip & node != root,
+    fill = cut(support, c(-1, 70, 90, 100))
+  ),
+  shape = 21,
+  size = 4) +
+  scale_fill_manual(
+    values = c('#d95f02', '#7570b3', '#1b9e77'),
+    labels = expression(BP < 70, 70 <= BP * " < 90", BP >= 90),
+    name = 'Bootstrap Percentage',
+    breaks = c('(-1,70]', '(70,90]', '(90,100]')
+  ) +
+  ggplot2::xlim(0, 2.5) +
+  theme_tree(legend.position = 'bottom')
+good_p
+```
+```{r}
+cowplot::save_plot(good_p, filename = 'output/EIN3_better_phylogeny.png')
 ```

### 2023-11-20 analysis/metagenome.Rmd
add metagenomics files. add renv.

--- 2022-03-09 analysis/GO_and_ortho.Rmd
@@
 ---
-title: "Visualising gene differences based on read alignments"
+title: "Visualising metagenome differences"
 author: "Philipp Bayer"
-date: "9th March 2022"
+date: "20th Nov 2023"
 output: html_document
@@
 
-```{r setup}
+```{r setup, message=FALSE}
 library(tidyverse)
-
+library(patchwork)
 knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
@@
 
-In this one I just merge some plots from the Orthofinder and GO enrichment Rmds, which I have saved as Rdata objects.
-
-```{r}
-load('output/all_GO_plots.Rdata')
-load('output/group_venn_image.Rdata')
-load('output/seagrass_venn_image.Rdata')
-```
-```{r}
-group_venn
-```
-```{r}
-seagrass_venn
-```
-
-OK now we can make patchworks with these Venn diagrams and their associated GO-terms
-```{r fig.width=8, fig.height=10}
-sum1 <- group_venn / (plot_list$GO_results_genes_in_Terrestrials_NOT_Aquatics_NOT_Seagrasses.csv + theme(legend.position="bottom")) +  plot_annotation(tag_levels = 'A') +plot_layout(heights=c(1.5,1))
-sum1
-cowplot::save_plot(sum1, filename = 'output/patchwork_terrestrials_gene_loss.png', base_height = 10, base_width = 8)
-```
-
-```{r fig.width=8, fig.height=10}
-sum2 <- seagrass_venn / (plot_list$GO_results_genes_union_Aquatics_and_Seagrasses_NOT_Terrestrials.csv + theme(legend.position="bottom")) +  plot_annotation(tag_levels = 'A') +plot_layout(heights=c(1.5,1))
-sum2
-cowplot::save_plot(sum2, filename = 'output/patchwork_seagrass_gene_loss.png', base_height = 10, base_width = 8)
-```
+Here we compare metagenomes (taxonomy and genes) across Amphibolis vs the others.

### 2023-11-20 analysis/metagenome.Rmd
Add reformatted CAT for easier plotting

@@
 Here we compare metagenomes (taxonomy and genes) across Amphibolis vs the others.
+
+```{r}
+metadata <- read_tsv('./data/metagenome/samples.tsv')
+metadata <- metadata |> rename(Sample_species = Species,
+                               Sample_replicate = Replicate) |> 
+  mutate(Tissue = str_to_title(Tissue))
+```
+
+# Taxonomies
+
+We have two taxonomies: one based on CAT, one based on GTDB. GTDB is less complete but more accurate. Let's see what they look like.
+
+# GTDB
+```{r}
+tax <- read_tsv('./data/metagenome/all_gtdb.summary.tsv.gz')
+```
+
+```{r}
+clean_tax <- tax |> 
+  separate(user_genome, into = c('sample', 'MAG'), sep='\\.') |> 
+  separate(classification, into = c('domain', 'phylum', 
+                                    'class', 'order', 
+                                    'family', 'genus', 
+                                    'species'), sep=';') |> 
+  mutate(sample = str_replace_all(sample, 'MEGAHIT-', ''), 
+         sample = str_replace_all(sample, '_L2', ''))
+
+clean_tax <- clean_tax |> left_join(metadata, by = c('sample'='Sample'))
+```
+
+Next, we keep only Posidonia australis and Amphibolis antarctica samples
+
+```{r}
+amph_pos_tax <- clean_tax |> filter(Sample_species %in% c('Amphibolis antarctica', 'Posidonia australis'))
+```
+
+## GTDB by order
+
+```{r}
+amph_pos_tax |> 
+  group_by(sample, Tissue) |> 
+  mutate(order = str_replace_all(order, 'o__','')) |> 
+  count(order) |> 
+  filter ( n > 1) |> 
+  filter(order != '' ) |> 
+  ggplot(aes(x = sample, y = n, fill=order)) +
+  geom_col() + 
+  theme_minimal() + 
+  #theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
+  facet_wrap(~Tissue) + coord_flip() + 
+  ylab('Number of contigs') + xlab('Sample')
+```
+
+## GTDB by family
+
+```{r}
+amph_pos_tax |> 
+  group_by(sample, Tissue) |> 
+  mutate(family = str_replace_all(family, 'f__','')) |> 
+  count(family) |> 
+  filter ( n > 1) |> 
+  filter(family != '' ) |> 
+  ggplot(aes(x = sample, y = n, fill=family)) +
+  geom_col() + 
+  theme_minimal() + 
+  #theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
+  facet_wrap(~Tissue) + coord_flip() + 
+  ylab('Number of contigs') + xlab('Sample')
+```
+Let's do the same using CAT
+
+# CAT
+
+```{r}
+tax_cat <- read_tsv('./data/metagenome/all_cat.summary.tsv.gz', guess_max = 100000)
+```
+
+```{r}
+clean_tax_cat <- tax_cat |> 
+  separate(`# bin`, into = c('sample', 'MAG'), sep='\\.') |> 
+  separate(full_lineage_names, into = c('root', 'root2', 
+                                    'superkingdom', 'clade', 
+                                    'phylum', 'subphylum', 
+                                    'class'), sep='\t') |> 
+  mutate(sample = str_replace_all(sample, 'MEGAHIT-', ''), 
+         sample = str_replace_all(sample, '_L2', ''))
+
+clean_tax <- clean_tax |> left_join(metadata, by = c('sample'='Sample'))
+```
+
+Next, we keep only Posidonia australis and Amphibolis antarctica samples
+
+```{r}
+amph_pos_tax <- clean_tax |> filter(Sample_species %in% c('Amphibolis antarctica', 'Posidonia australis'))
+```

### 2023-11-20 data/metagenome/reformat.py

@@ -1,0 +1,29 @@
+all_types=set()
+# all types are
+
+# {'clade', 'subphylum', 'genus', 'tribe', 'order', 'superkingdom', 'species', 'suborder', 'class', 'subclass', 'subfamily', 'family', 'phylum', 'varietas', 'kingdom'}
+TOKEEP = set(['superkingdom', 'class', 'order', 'family', 'genus', 'species'])
+
+with open('all_cat.summary.tsv') as fh:
+    header = fh.readline().rstrip().split('\t')
+    print('\t'.join(header[:-1]) + '\tSuperkingdom\tClass\tOrder\tFamily\tGenus\tSpecies')
+    for line in fh:
+        ll = line.rstrip().split('\t')
+        lineages = ll[6:]
+        this_types = {}
+        for l in lineages:
+            thistype = l.split(':')[0].split('(')[-1].replace(')', '')
+            l = l.replace(f' ({thistype})', '')
+
+            if thistype == 'no rank': continue
+            if thistype in TOKEEP:
+                if thistype in this_types:
+                    this_types[thistype] = this_types[thistype] + ', ' + l
+                else:
+                    this_types[thistype] = l
+
+        for a in TOKEEP:
+            if a not in this_types:
+                this_types[a] = 'NA'
+        newll = [this_types[i] for i in ['superkingdom', 'class', 'order', 'family', 'genus', 'species']]
+        print('\t'.join(ll[:6]) +'\t'+ '\t'.join(newll))

### 2023-11-20 analysis/index.Rmd
New: fancy MSAs!!
workflowr:::wflow_publish(files = "analysis/index.Rmd")
Expand metagenome analysis

@@
 [GO and Ortho](GO_and_ortho.html) - merging plots from GO enrichment and Orthofinder analysis
+
+[MSA visualisation](MSA.html) - visualising MSAs and phylogenies for gene clusters of interest (*SOS3*, *EIN3*)
+
+[3D structure](ThreeDStructure.html) - visualising the predicted 3D structure of *SOS3*/*EIN3*
+
+[Metagenomics]('metagenome.html') - visualising metagenome differences across Posidonia/Amphibolis

### 2023-11-20 analysis/metagenome.Rmd
Expand metagenome analysis

--- 2023-11-20 analysis/metagenome.Rmd
@@
 library(tidyverse)
+library(UpSetR)
+library(microshades)
+library(ggtext)
 library(patchwork)
+library(scales)
 knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
@@
 
+```{r}
+trunc_str <- function(latin_name) {
+  tmp <- str_split(latin_name, pattern = ' ')[[1]]
+  return(paste0(substring(tmp[1], 1, 1), '. ', tmp[2]))
+}
+
+```
+
 Here we compare metagenomes (taxonomy and genes) across Amphibolis vs the others.
+
+```{r}
+metadata <- read_tsv('./data/metagenome/samples.tsv')
+metadata <- metadata |> rename(Sample_species = Species,
+                               Sample_replicate = Replicate) |> 
+  mutate(Tissue = str_to_title(Tissue))
+```
+
+# Taxonomies
+
+We have two taxonomies: one based on CAT, one based on GTDB. GTDB is less complete but more accurate. Let's see what they look like.
+
+# GTDB
+```{r}
+tax <- read_tsv('./data/metagenome/all_gtdb.summary.tsv.gz')
+```
+
+```{r}
+clean_tax <- tax |> 
+  separate(user_genome, into = c('sample', 'MAG'), sep='\\.') |> 
+  separate(classification, into = c('domain', 'phylum', 
+                                    'class', 'order', 
+                                    'family', 'genus', 
+                                    'species'), sep=';') |> 
+  mutate(sample = str_replace_all(sample, 'MEGAHIT-', ''), 
+         sample = str_replace_all(sample, '_L2', ''))
+
+clean_tax <- clean_tax |> left_join(metadata, by = c('sample'='Sample'))
+```
+
+Next, we keep only Posidonia australis and Amphibolis antarctica samples
+
+```{r}
+amph_pos_tax <- clean_tax |> filter(Sample_species %in% c('Amphibolis antarctica', 'Posidonia australis'))
+```
+
+## GTDB by order
+
+```{r}
+colors <-c(microshades_palette("micro_blue", 4, lightest = FALSE), 
+           microshades_palette("micro_purple", 4, lightest = FALSE),
+           microshades_palette("micro_green", 4, lightest = FALSE),
+          microshades_palette("micro_orange", 4, lightest = FALSE))
+
+```
+
+```{r fig.height=6}
+plot_orders_gtdb <- function(amph_pos_tax, group_remainder = TRUE) {
+  temp <- amph_pos_tax |> 
+    group_by(sample, Tissue, Sample_species) |> 
+    mutate(Sample_species = trunc_str(Sample_species)) |> 
+    mutate(Sample_species = paste0('*', Sample_species, '*')) |> 
+    mutate(order = str_replace_all(order, 'o__','')) |> 
+    count(order) 
+  if(group_remainder) {
+    temp <- temp |> 
+    mutate(counted_orders = case_when(n == 1 ~ 'Remainder',
+                                      TRUE ~ order))
+  } else {
+      temp <- temp |> 
+        mutate(counted_orders = order)
+  }
+  temp <- temp |> 
+    select(-order) |> 
+    group_by(sample, Tissue, Sample_species, n, counted_orders) |> 
+    summarise(n = sum(n)) |> 
+    filter(counted_orders != '' )
+
+  uniques <- c(setdiff(unique(temp$counted_orders), c('Remainder')), 'Remainder')
+  temp |> 
+    mutate(counted_orders = factor(counted_orders, levels=uniques)) |> 
+    mutate(sample = str_remove(sample, "[A-Z]+")) |> 
+    mutate(sample = str_replace(sample, '31', '3')) |> 
+    ggplot(aes(x = factor(sample, levels=rev(unique(sample))), y = n, fill=counted_orders)) +
+    geom_col() + 
+    theme_minimal() + 
+    facet_wrap(~Tissue+Sample_species, scales = 'free_y', ncol=2) + 
+    coord_flip() + 
+    ylab('Number of contigs') + xlab('Replicate') +
+    labs(fill = 'Order') +
+    scale_y_continuous(breaks=pretty_breaks())  +
+    # move legend position to bottom
+    theme(legend.position = "bottom",
+          strip.text.x = element_markdown()) +
+    scale_fill_manual(values = colors)
+}
+plot_orders_gtdb(amph_pos_tax)
+```
+Let's also make an Upset plot of that
+
+```{r}
+tmp <- amph_pos_tax |> 
+  group_by(sample) |> 
+  mutate(order = str_replace_all(order, 'o__','')) |> 
+  count(order) |> 
+  filter(order != '' ) |>
+  filter ( n > 1) |> 
+  summarise(named_vec = list(order)) %>%
+  deframe()
+tmp
+```
+```{r}
+upset(fromList(tmp))
+```
+
+## GTDB by family
+
+```{r}
+amph_pos_tax |> 
+  group_by(sample, Tissue) |> 
+  mutate(family = str_replace_all(family, 'f__','')) |> 
+  count(family) |> 
+  filter ( n > 1) |> 
+  filter(family != '' ) |> 
+  ggplot(aes(x = sample, y = n, fill=family)) +
+  geom_col() + 
+  theme_minimal() + 
+  #theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
+  facet_wrap(~Tissue) + coord_flip() + 
+  ylab('Number of contigs') + xlab('Sample')
+```
+```{r}
+amph_pos_tax |> group_by(sample, species) |> filter(species != 's__') |> 
+  count(species)
+```
+
+# CAT
+Let's do the same using CAT
+
+```{r}
+tax_cat <- read_tsv('./data/metagenome/all_cat.summary.reformatted.tsv.gz')
+```
+
+```{r}
+clean_tax_cat <- tax_cat |> 
+  separate(`# bin`, into = c('sample', 'MAG', 'fa'), sep='\\.') |> 
+  mutate(sample = str_replace_all(sample, 'MEGAHIT-', ''), 
+         sample = str_replace_all(sample, '_L2', '')) |> 
+  select(-fa)
+
+clean_tax_cat <- clean_tax_cat |> left_join(metadata, by = c('sample'='Sample'))
+```
+
+Next, we keep only Posidonia australis and Amphibolis antarctica samples
+
+```{r}
+amph_pos_tax_cat <- clean_tax_cat |> filter(Sample_species %in% c('Amphibolis antarctica', 'Posidonia australis'))
+```
+
+## CAT by order
+
+```{r fig.height=6}
+plot_orders_cat <-
+  function(amph_pos_tax_cat, group_remainder = TRUE) {
+    temp <- amph_pos_tax_cat |>
+      separate(Order, into = c('Order', 'Order_score')) |>
+      rename(order = Order) |>
+      group_by(sample, Tissue, Sample_species) |>
+      mutate(Sample_species = trunc_str(Sample_species)) |>
+      mutate(Sample_species = paste0('*', Sample_species, '*')) |>
+      count(order)
+    if (group_remainder) {
+      temp <- temp |>
+        mutate(counted_orders = case_when(n == 1 ~ 'Remainder',
+                                          TRUE ~ order))
+    } else {
+      temp <- temp |>
+        mutate(counted_orders = order)
+    }
+    temp <- temp |>
+      select(-order) |>
+      group_by(sample, Tissue, Sample_species, n, counted_orders) |>
+      summarise(n = sum(n)) |>
+      filter(counted_orders != '')
+
+    uniques <-
+      c(setdiff(unique(temp$counted_orders), c('Remainder')), 'Remainder')
+    temp |>
+      mutate(counted_orders = factor(counted_orders, levels = uniques)) |>
+      mutate(sample = str_remove(sample, "[A-Z]+")) |>
+      mutate(sample = str_replace(sample, '31', '3')) |>
+      ggplot(aes(
+        x = factor(sample, levels = rev(unique(sample))),
+        y = n,
+        fill = counted_orders
+      )) +
+      geom_col() +
+      theme_minimal() +
+      facet_wrap( ~ Sample_species + Tissue, ncol = 2, scales = 'free_y') +
+      coord_flip() +
+      ylab('Number of contigs') + xlab('Replicate') +
+      labs(fill = 'Order') +
+      scale_y_continuous(breaks = pretty_breaks())  +
+      # move legend position to bottom
+      theme(legend.position = "bottom",
+            strip.text.x = element_markdown()) +
+      scale_fill_manual(values = colors)
+  }
+plot_orders_cat(amph_pos_tax_cat)
+```
+
+## CAT by family
+
+```{r}
+amph_pos_tax_cat |> 
+  separate(Family, into = c('Family', 'Family_score')) |> 
+  group_by(sample, Tissue) |> 
+  count(Family) |> 
+  filter ( n > 1) |> 
+  filter(Family != '' ) |> 
+  ggplot(aes(x = sample, y = n, fill=Family)) +
+  geom_col() + 
+  theme_minimal() + 
+  #theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
+  facet_wrap(~Tissue) + coord_flip() + 
+  ylab('Number of contigs') + xlab('Sample')
+```
+
+```{r}
+amph_pos_tax_cat |> separate(Species, into = c('Species', 'Species_score'), sep =':') |> filter(!is.na(Species)) |>  group_by(sample, Species) |> 
+  count(Species)
+```
+```{r}
+amph_pos_tax_cat |> separate(Species, into = c('Species', 'Species_score'), sep =':') |> filter(!is.na(Species)) |>  group_by(sample, Species) |> 
+  count(Species) |> filter(str_detect(Species, 'Acidihalobacter'))
+```
+
+There we go :) Which MAGs are those?
+```{r}
+amph_pos_tax_cat |> filter(str_detect(Species, 'Acidihalobacter'))
+```
+
+```{r}
+amph_pos_tax_cat |> filter(str_detect(Family, 'Ectothiorhodospiraceae'))
+```
+Are these in the GTDB classification?
+```{r}
+amph_pos_tax |> filter(str_detect(family, 'Ectothiorhodospiraceae'))
+```
+Nope!
+
+# Gene comparison
+
+Let's compare gene presence/absence by names first
+
+```{r}
+genes <- read_tsv('././data/metagenome/no_hypothetical.all_genes.tsv.gz')
+
+genes <- genes |> 
+  separate(filename, into = c('sample', 'MAG'), sep='\\.', extra = 'drop') |>  # get rid of extra filenaming stuff
+  mutate(sample = str_replace(sample, 'MEGAHIT-', '')) |> 
+  separate(MAG, into = c('MAG', 'rest'), sep ='/') |> 
+  select(-rest)
+```
+
+```{r}
+genessplit <- genes |> separate(gene, into = c('gene', 'gene_copy_number'), sep='_')
+```
+
+```{r}
+genessplit <- genessplit |> mutate(sample = str_replace(sample, '_L2', '')) |> 
+  left_join(metadata, by = c('sample'='Sample')) |> 
+  filter(Sample_species %in% c('Amphibolis antarctica', 'Posidonia australis'))
+
+```
+
+```{r}
+count_diffs <- genessplit |> 
+  group_by(Sample_species, gene) |> 
+  count(gene) |> 
+  pivot_wider(names_from = Sample_species, values_from = n) |> 
+  mutate(difference = `Amphibolis antarctica` - `Posidonia australis`) |> 
+  arrange(desc(difference))
+```
+
+```{r}
+count_diffs |> filter(`Posidonia australis` <100)
+```
+
+We hypothesise that the ACC precursor-existing genes in Amphibolis lead to different microbiomes.
+ACC is broken down by ACC deaminases, which are present in Amphibolis but not Posidonia.
+Let's see if we can find ACC deaminases in the metagenomes (*acdS*).
+
+```{r}
+genessplit |> filter(gene == 'acdS') |> 
+  group_by(Sample_species, Tissue, Sample_replicate) |> 
+  count(gene) |> 
+  mutate(Sample_species = trunc_str(Sample_species)) |> 
+  mutate(Sample_species = paste0('*', Sample_species, '*')) |> 
+  ggplot(aes(x= interaction(factor(Sample_replicate, levels=4:1), Tissue), y = n, fill = Sample_species)) + 
+  geom_col() + 
+  facet_wrap(~Sample_species) +
+  coord_flip()+ theme_minimal() + 
+  scale_y_continuous(breaks=pretty_breaks()) +
+  ylab('Total number of *acdS*-containing MAGs') +
+  xlab('Sample species') + 
+  theme(legend.position = 'none',
+        strip.text.x =  element_markdown(),
+        axis.title.x = element_markdown())
+```
+
+What are those acdS-containing MAGs?
+
+```{r}
+genessplit |> filter(gene == 'acdS') |> left_join(amph_pos_tax, by=c('sample', 'MAG'), multiple = 'any')
+```
+Lots of unknown taxonomies :(
+
+```{r}
+genessplit |> filter(gene == 'acdS') |> left_join(amph_pos_tax, by=c('sample', 'MAG'), multiple = 'any') |> 
+  select(-c(Tissue.y, Sample_species.y)) |>
+  rename(Tissue = Tissue.x, Sample_species = Sample_species.x) |> 
+  plot_orders_gtdb(group_remainder = FALSE)
+  #select(sample, MAG, domain:species)
+```
+
+```{r}
+genessplit |> filter(gene == 'acdS') |> 
+  left_join(amph_pos_tax_cat, by=c('sample', 'MAG')) |> 
+    select(-c(Tissue.y, Sample_species.y)) |>
+  rename(Tissue = Tissue.x, Sample_species = Sample_species.x) |> 
+  plot_orders_cat(group_remainder = FALSE) #
+  #select(sample, MAG, Superkingdom:Species) |>
+
+```

### 2023-11-20 analysis/metagenome.Rmd
workflowr::wflow_publish(files = c("analysis/index.Rmd", "analysis/metagenome.Rmd"))

@@
 ```{r}
-amph_pos_tax_cat |> filter(str_detect(Species, 'Acidihalobacter'))
+amph_pos_tax_cat |> filter(str_detect(Species, 'Acidihalobacter')) |> knitr::kable()
+
 ```
@@
 ```{r}
-count_diffs |> filter(`Posidonia australis` <100)
+count_diffs |> filter(`Posidonia australis` <100) |> head() |> knitr::kable()
 ```
@@
 ```{r}
-genessplit |> filter(gene == 'acdS') |> left_join(amph_pos_tax, by=c('sample', 'MAG'), multiple = 'any')
+genessplit |> filter(gene == 'acdS') |> left_join(amph_pos_tax, by=c('sample', 'MAG'), multiple = 'any') |> knitr::kable()
 ```

### 2023-11-21 analysis/metagenome.Rmd
Final figures and changes

@@
     mutate(order = str_replace_all(order, 'o__','')) |> 
-    count(order) 
+    count(order, .drop = FALSE) 
   if(group_remainder) {
@@
 ```
+
+Let's chekc for nitrogen fixation genes (*nif*) - hypothesis is that Amphibolis roots contain more of these
+
+```{r}
+genessplit |> filter(str_detect(gene, 'nifH') | str_detect(gene, 'acdS') | str_detect(gene, 'sctC')) |> 
+  mutate(gene_class = case_when(str_detect(gene, 'nifH') ~ '*nifH*',
+                                str_detect(gene, 'acdS') ~ '*acdS*',
+                                str_detect(gene, 'sctC') ~ '*sctC*')) |>
+  #group_by(Sample_species, Tissue, Sample_replicate, MAG) |> 
+  #summarise(gene_class = paste0(sort(unique(gene_class)), collapse = ' & ')) |>
+  group_by(Sample_species, Tissue, Sample_replicate) |> 
+  count(gene_class) |> 
+  mutate(Sample_species = trunc_str(Sample_species)) |> 
+  mutate(Sample_species = paste0('*', Sample_species, '*')) |> 
+  ggplot(aes(x=Tissue, y=n, fill=gene_class)) + 
+  geom_boxplot(outlier.shape = NA) +
+  facet_wrap(~Sample_species) + 
+  geom_point(position = position_jitterdodge(), aes(color=gene_class)) +
+  coord_flip() +
+  theme_minimal() +
+  labs(color='MAG gene content', fill='MAG gene content') +
+  theme(strip.text.x =  element_markdown(),legend.text = element_markdown()) +
+  ylab('Count of MAGs')
+
+```
+
+Can we make a simple association test?
+
+```{r}
+tmp <- genessplit |> filter(str_detect(gene, 'nifH') | str_detect(gene, 'acdS') | str_detect(gene, 'sctC')) |> 
+  mutate(gene_class = case_when(str_detect(gene, 'nifH') ~ '*nifH*',
+                                str_detect(gene, 'acdS') ~ '*acdS*',
+                                str_detect(gene, 'sctC') ~ '*sctC*')) |> 
+  group_by(Sample_species, Sample_replicate, Tissue) |> 
+  count(gene_class) |> 
+  mutate(Sample_species = factor(Sample_species, levels = c('Posidonia australis', 'Amphibolis antarctica'))) 
+summary(lm(n ~ gene_class + Sample_species + Sample_replicate + Tissue, data=tmp))
+```
+
+OK! So we have an association with gene_class, which is obvious, but we also have an association with species (great!) and root-tissue (also great!)

### 2023-11-21 analysis/metagenome.Rmd
add missing ifles

@@
   mutate(Sample_species = paste0('*', Sample_species, '*')) |> 
-  ggplot(aes(x=Tissue, y=n, fill=gene_class)) + 
+  ggplot(aes(x=factor(Sample_species, levels=rev(c(unique(Sample_species)))), y=n, fill=gene_class)) + 
   geom_boxplot(outlier.shape = NA) +
-  facet_wrap(~Sample_species) + 
+  facet_wrap(~factor(Tissue, levels=rev(c(unique(Tissue))))) + 
   geom_point(position = position_jitterdodge(), aes(color=gene_class)) +
@@
   labs(color='MAG gene content', fill='MAG gene content') +
-  theme(strip.text.x =  element_markdown(),legend.text = element_markdown()) +
-  ylab('Count of MAGs')
+  theme(axis.text.y =  element_markdown(),legend.text = element_markdown()) +
+  ylab('Count of MAGs') + xlab('Species')
 

### 2023-11-21 analysis/metagenome.Rmd
Add the unclassified MAGs too

@@
 library(tidyverse)
-library(UpSetR)
 library(microshades)
 library(ggtext)
-library(patchwork)
 library(scales)
@@
 
+Need to put some extra work in; MAGs with no classification are not in the output file.
+
+```{r}
+all_mag_counts <- read_tsv('./data/metagenome/MAG_counts.tsv') |> mutate(Sample = str_replace(Sample, '_L2', '')) |> 
+  rename(TotalMAGs = MAGs)
+```
+
 ```{r}
@@
 ```{r}
-amph_pos_tax <- clean_tax |> filter(Sample_species %in% c('Amphibolis antarctica', 'Posidonia australis'))
+amph_pos_tax <- clean_tax |> filter(Sample_species %in% c('Amphibolis antarctica', 'Posidonia australis')) |> 
+  left_join(all_mag_counts, by = c('sample'='Sample'))
 ```
@@
   temp <- amph_pos_tax |> 
-    group_by(sample, Tissue, Sample_species) |> 
+    group_by(sample, Tissue, Sample_species,TotalMAGs) |> 
     mutate(Sample_species = trunc_str(Sample_species)) |> 
@@
   }
-  temp <- temp |> 
+  temp2 <- temp |> 
     select(-order) |> 
@@
     filter(counted_orders != '' )
+  # we need to count the unclassified MAGs that are not in the output
+  group_presents <- temp2 |> group_by(sample) |> summarise(classifiedMAGs = sum(n))
+  missings <- group_presents |> 
+    left_join(temp |> 
+                select(sample, TotalMAGs)) |> 
+    unique() |> 
+    mutate(unclassifiedMAGs = TotalMAGs - classifiedMAGs) |> 
+    select(-c(classifiedMAGs,TotalMAGs)) |>
+    rename(Unclassified = unclassifiedMAGs) |>
+    pivot_longer(-c(sample, Tissue, Sample_species)) |> 
+    rename(counted_orders=name, n=value)
 
-  uniques <- c(setdiff(unique(temp$counted_orders), c('Remainder')), 'Remainder')
-  temp |> 
+  temp2 <-rbind(temp2, missings)
+
+  uniques <- c(setdiff(unique(temp2$counted_orders), c('Unclassified', 'Remainder')), 'Remainder', 'Unclassified')
+  temp2 |> 
     mutate(counted_orders = factor(counted_orders, levels=uniques)) |> 
@@
 plot_orders_gtdb(amph_pos_tax)
-```
-Let's also make an Upset plot of that
-
-```{r}
-tmp <- amph_pos_tax |> 
-  group_by(sample) |> 
-  mutate(order = str_replace_all(order, 'o__','')) |> 
-  count(order) |> 
-  filter(order != '' ) |>
-  filter ( n > 1) |> 
-  summarise(named_vec = list(order)) %>%
-  deframe()
-tmp
-```
-```{r}
-upset(fromList(tmp))
 ```
@@
 
-clean_tax_cat <- clean_tax_cat |> left_join(metadata, by = c('sample'='Sample'))
+clean_tax_cat <- clean_tax_cat |> left_join(metadata, by = c('sample'='Sample')) |> left_join(all_mag_counts, by =c('sample'='Sample'))
 ```
@@
       rename(order = Order) |>
-      group_by(sample, Tissue, Sample_species) |>
+      group_by(sample, Tissue, Sample_species, TotalMAGs) |>
       mutate(Sample_species = trunc_str(Sample_species)) |>
@@
     }
-    temp <- temp |>
+    temp2 <- temp |>
       select(-order) |>
-      group_by(sample, Tissue, Sample_species, n, counted_orders) |>
+      group_by(sample, Tissue, Sample_species, n, counted_orders,TotalMAGs) |>
       summarise(n = sum(n)) |>
       filter(counted_orders != '')
+      # we need to count the unclassified MAGs that are not in the output
+    group_presents <- temp2 |> group_by(sample) |> summarise(classifiedMAGs = sum(n))
+    missings <- group_presents |> 
+      left_join(temp2) |> 
+      select(classifiedMAGs, Tissue, Sample_species, sample, TotalMAGs)|> 
+      unique() |> 
+      mutate(unclassifiedMAGs = TotalMAGs - classifiedMAGs) |> 
+      select(-c(classifiedMAGs,TotalMAGs)) |>
+      rename(Unclassified = unclassifiedMAGs) |>
+      pivot_longer(-c(sample, Tissue, Sample_species)) |> 
+      rename(counted_orders=name, n=value)
 
-    uniques <-
-      c(setdiff(unique(temp$counted_orders), c('Remainder')), 'Remainder')
-    temp |>
+    temp2 <-rbind(temp2, missings)
+
+   uniques <- c(setdiff(unique(temp2$counted_orders), c('Unclassified', 'Remainder')), 'Remainder', 'Unclassified')
+
+    temp2 |>
       mutate(counted_orders = factor(counted_orders, levels = uniques)) |>

### 2023-11-21 analysis/metagenome.Rmd
add consistent color scheme for orders

@@
            microshades_palette("micro_orange", 4, lightest = FALSE))
-
+names(colors) <- c("Rhizobiales", "Clostridiales", "Desulfobacterales", "Chitinophagales", 
+"Chromatiales", "Solirubrobacterales", "Xanthomonadales", "Acidimicrobiales", 
+"Rhodobacterales", "Corallinales", "Ceramiales", "Alismatales", 
+"Thiotrichales", "Flavobacteriales", "Remainder", "Unclassified")
 ```

### 2023-11-21 analysis/metagenome.Rmd
Add more t-tests

@@
 
+```{r}
+genessplit |> filter(str_detect(gene, 'nifH') | str_detect(gene, 'acdS') | str_detect(gene, 'sctC')) |> 
+  mutate(gene_class = case_when(str_detect(gene, 'nifH') ~ '*nifH*',
+                                str_detect(gene, 'acdS') ~ '*acdS*',
+                                str_detect(gene, 'sctC') ~ '*sctC*')) |>
+  #group_by(Sample_species, Tissue, Sample_replicate, MAG) |> 
+  #summarise(gene_class = paste0(sort(unique(gene_class)), collapse = ' & ')) |>
+  group_by(Sample_species, Tissue, Sample_replicate) |> 
+  count(gene_class) |> 
+  mutate(Sample_species = trunc_str(Sample_species)) |> 
+  mutate(Sample_species = paste0('*', Sample_species, '*')) |> 
+  ggplot(aes(x=factor(Tissue, levels=rev(c('Shoot/Leaf', 'Shoot/Leaf Biofilm', 'Rhizome', 'Root'))), y=n, fill=Sample_species)) + 
+  geom_boxplot(outlier.shape = NA) +
+  facet_wrap(~factor(gene_class, levels=c('*acdS*', '*nifH*', '*sctC*')), scales = 'free_x') + 
+  geom_point(position = position_jitterdodge(), aes(color=Sample_species)) +
+  coord_flip() +
+  theme_minimal() +
+  labs(color='Species', fill='Species') +
+  theme(axis.text.y =  element_markdown(),legend.text = element_markdown(),
+        strip.text.x = element_markdown()) +
+  ylab('Count of MAGs') + xlab('Species')
+```
+I like this one more. The differences are stronger.
+
+# Statistic test
+
 Can we make a simple association test?
+
+First, a general one:
 
@@
 OK! So we have an association with gene_class, which is obvious, but we also have an association with species (great!) and root-tissue (also great!)
+
+## Subtests
+Now let's make the specific ones; nifh, acdS, and sctC counts within Tissue=root across both species
+
+### nifH
+```{r}
+tmp <- genessplit |> filter(str_detect(gene, 'nifH') & str_detect(Tissue, 'Root')) |> 
+  mutate(gene_class = case_when(str_detect(gene, 'nifH') ~ 'nifH')) |> 
+  group_by(Sample_species, Sample_replicate) |> count(gene_class) 
+tmp  
+```
+
+```{r}
+t.test(n ~ Sample_species, data=tmp)
+```
+
+P < 0.0005
+
+### sctC
+
+```{r}
+tmp <- genessplit |> filter(str_detect(gene, 'sctC') & str_detect(Tissue, 'Root')) |> 
+  mutate(gene_class = case_when(str_detect(gene, 'sctC') ~ 'sctC')) |> 
+  group_by(Sample_species, Sample_replicate) |> count(gene_class) 
+tmp  
+```
+
+```{r}
+t.test(n ~ Sample_species, data=tmp)
+```
+
+p < 0.005
+
+### acdS
+
+```{r}
+tmp <- genessplit |> filter(str_detect(gene, 'acdS') & str_detect(Tissue, 'Root')) |> 
+  mutate(gene_class = case_when(str_detect(gene, 'acdS') ~ 'acdS')) |> 
+  group_by(Sample_species, Sample_replicate) |> count(gene_class) 
+tmp 
+```
+
+can't run t-test, only one group present!
+
+```{r}
+# t.test(n ~ Sample_species, data=tmp)
+```

### 2023-11-21 analysis/metagenome.Rmd
yay

@@
 library(tidyverse)
+library(DT)
 library(microshades)
@@
 ```{r}
-amph_pos_tax_cat |> filter(str_detect(Species, 'Acidihalobacter')) |> knitr::kable()
+amph_pos_tax_cat |> filter(str_detect(Species, 'Acidihalobacter')) |> datatable()
 
@@
 ```{r}
-amph_pos_tax_cat |> filter(str_detect(Family, 'Ectothiorhodospiraceae'))
+amph_pos_tax_cat |> filter(str_detect(Family, 'Ectothiorhodospiraceae'))|> datatable()
 ```
@@
 ```{r}
-amph_pos_tax |> filter(str_detect(family, 'Ectothiorhodospiraceae'))
+amph_pos_tax |> filter(str_detect(family, 'Ectothiorhodospiraceae'))|> datatable()
 ```
@@
 ```{r}
-count_diffs |> filter(`Posidonia australis` <100) |> head() |> knitr::kable()
+count_diffs |> filter(`Posidonia australis` <100) |> datatable()
 ```
@@
 ```{r}
-genessplit |> filter(gene == 'acdS') |> left_join(amph_pos_tax, by=c('sample', 'MAG'), multiple = 'any') |> knitr::kable()
+genessplit |> filter(gene == 'acdS') |> left_join(amph_pos_tax, by=c('sample', 'MAG'), multiple = 'any') |> datatable()
 ```
@@
   group_by(Sample_species, Sample_replicate) |> count(gene_class) 
-tmp  
+tmp  |> datatable()
 ```
@@
   group_by(Sample_species, Sample_replicate) |> count(gene_class) 
-tmp 
+tmp |> datatable()
 ```

### 2023-11-21 analysis/metagenome.Rmd
dada

@@
            microshades_palette("micro_green", 4, lightest = FALSE),
-           microshades_palette("micro_orange", 4, lightest = FALSE))
-names(colors) <- c("Rhizobiales", "Clostridiales", "Desulfobacterales", "Chitinophagales", 
-"Chromatiales", "Solirubrobacterales", "Xanthomonadales", "Acidimicrobiales", 
+           microshades_palette("micro_orange", 4, lightest = FALSE),
+          microshades_palette("micro_brown", 4, lightest = FALSE),
+           microshades_palette("micro_gray", 4, lightest = FALSE))
+names(colors) <- c("Rhizobiales", "Clostridiales", "Desulfobacterales", 'Desulfarculales', "Chitinophagales", "SBR1031", "Chromatiales", "Solirubrobacterales", "Xanthomonadales", "Acidimicrobiales", "Arenicellales","Granulosicoccales",
 "Rhodobacterales", "Corallinales", "Ceramiales", "Alismatales", 
-"Thiotrichales", "Flavobacteriales", "Remainder", "Unclassified")
+"Thiotrichales", "Flavobacteriales", "UBA4575", "Remainder", "Unclassified")
+
 ```
@@
 ```{r fig.height=6}
-plot_orders_gtdb <- function(amph_pos_tax, group_remainder = TRUE) {
+plot_orders_gtdb <- function(amph_pos_tax, group_remainder = TRUE, add_up=TRUE) {
   temp <- amph_pos_tax |> 
@@
 
+  if(add_up){
+    # add hte missing numbers ot the main table
     temp2 <-rbind(temp2, missings)
-
+  }
+  temp2 <- temp2 |>  mutate(counted_orders = replace_na(counted_orders, 'Unclassified'))
   uniques <- c(setdiff(unique(temp2$counted_orders), c('Unclassified', 'Remainder')), 'Remainder', 'Unclassified')
@@
     coord_flip() + 
-    ylab('Number of contigs') + xlab('Replicate') +
+    ylab('Number of MAGs') + xlab('Replicate') +
     labs(fill = 'Order') +
@@
       coord_flip() +
-      ylab('Number of contigs') + xlab('Replicate') +
+      ylab('Number of MAGs') + xlab('Replicate') +
       labs(fill = 'Order') +
@@
 ```{r}
-amph_pos_tax_cat |> filter(str_detect(Species, 'Acidihalobacter')) |> datatable()
-
+amph_pos_tax_cat |> filter(str_detect(Species, 'Acidihalobacter')) |> select(-c(lineage, `lineage scores`)) |> datatable()
 ```
@@
 ```{r}
-amph_pos_tax_cat |> filter(str_detect(Family, 'Ectothiorhodospiraceae'))|> datatable()
+amph_pos_tax_cat |> filter(str_detect(Family, 'Ectothiorhodospiraceae'))|> select(-c(lineage, `lineage scores`)) |>  datatable()
 ```
@@
 ```{r}
-amph_pos_tax |> filter(str_detect(family, 'Ectothiorhodospiraceae'))|> datatable()
+amph_pos_tax |> filter(str_detect(family, 'Ectothiorhodospiraceae'))|> nrow()
 ```
@@
 ```{r}
-count_diffs |> filter(`Posidonia australis` <100) |> datatable()
+count_diffs |> filter(`Posidonia australis` <100) |> head() |> datatable()
 ```
+Eeeh this is too unspecific. Let's look into specific differences.
 
+## ACC deaminase
 We hypothesise that the ACC precursor-existing genes in Amphibolis lead to different microbiomes.
@@
 
+TODO: The unknowns should have a color!!
+
 ```{r}
-genessplit |> filter(gene == 'acdS') |> left_join(amph_pos_tax, by=c('sample', 'MAG'), multiple = 'any') |> 
+genessplit |> filter(gene == 'acdS') |> left_join(amph_pos_tax, by=c('sample', 'MAG')) |>
   select(-c(Tissue.y, Sample_species.y)) |>
   rename(Tissue = Tissue.x, Sample_species = Sample_species.x) |> 
-  plot_orders_gtdb(group_remainder = FALSE)
+  mutate(order = replace_na(order, 'Unclassified')) |>
+  plot_orders_gtdb(group_remainder = FALSE, add_up = FALSE)
   #select(sample, MAG, domain:species)

### 2023-11-22 analysis/GO_and_ortho.Rmd
make the setup blocks nicer
new Venn plots

@@
 
-```{r setup}
+```{r setup, message=FALSE}
 library(tidyverse)
 library(patchwork)
+library(wesanderson)
+library(RColorBrewer)
+library(UpSetR)
+library(kableExtra)
+library(ggVennDiagram)
 knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
@@
 group_venn
+```
+```{r}
 ```
@@
 ```
+
+```{r}
+a <- group_venn
+b <- plot_list$GO_results_genes_in_Terrestrials_NOT_Aquatics_NOT_Seagrasses.csv
+c <- seagrass_venn
+d <- plot_list$GO_results_genes_in_all_seagrasses_vs_seagrassesBackground_intersect.csv
+sum3 <- (a + c) / (b + d) +
+   plot_annotation(tag_levels = 'A') +
+  plot_layout(heights=c(1, 1.5))
+sum3
+cowplot::save_plot(sum3, filename = 'output/patchwork_terrestrials_seagrasses_gene_loss.png', base_height = 10)
+```
+
+```{r fig.width=8, fig.height=10}
+sum2 <- seagrass_venn / (plot_list$GO_results_genes_in_all_seagrasses_vs_seagrassesBackground_intersect.csv + theme(legend.position="bottom")) +  plot_annotation(tag_levels = 'A') +plot_layout(heights=c(1.5,1))
+sum2
+cowplot::save_plot(sum2, filename = 'output/patchwork_seagrass_gene_loss.png', base_height = 10, base_width = 8)
+```

### 2023-11-22 analysis/GOenrichment.Rmd
new Venn plots

@@
 library(rvest)
+library(rrvgo)
 library(wesanderson)
@@
 
+```{r}
+
+```
+
 This code is based on http://revigo.irb.hr/CodeExamples/revigo.R.txt
@@
 for (f in list.files('./output/', pattern='*.csv')){
+  if(str_detect(f, '.png')) {
+    next
+  }
   filename <- paste('./output/', f, sep='')
@@
   go_and_pvalues <- gsub(',', ' ', go_and_pvalues)
+  temp <- read_delim(go_and_pvalues,show_col_types = FALSE)
+  if(nrow(temp) < 2) {next}
+  simMatrix <- calculateSimMatrix(temp$name,
+                                  orgdb="org.At.tair.db",
+                                  ont="BP",
+                                  method="Rel")
 
-  httr::POST(
-    url = "http://revigo.irb.hr/StartJob.aspx",
-    body = list(
-      cutoff = "0.7",
-      valueType = "pvalue",
-     # speciesTaxon = "4577", # zea mays
-     #speciesTaxon = '39947', # japonica
-     speciesTaxon = '3702', # arabidopsis
-      measure = "SIMREL",
-      goList = go_and_pvalues
-    ),
-    # application/x-www-form-urlencoded
-    encode = "form"
-  ) -> res
+  tmp <- temp$value
+  names(tmp) <- temp$name
+  reducedTerms <- reduceSimMatrix(simMatrix,
+                                  tmp,
+                                  threshold=0.7,
+                                  orgdb="org.At.tair.db")
+  results_list[[f]] <- list('simMatrix'=simMatrix, 'reducedTerms'=reducedTerms)
 
-  dat <- httr::content(res, encoding = "UTF-8")
-
-  jobid <- jsonlite::fromJSON(dat,bigint_as_char=TRUE)$jobid
-
-  # Check job status
-  running <- "1"
-  while (running != "0" ) {
-      httr::POST(
-        url = "http://revigo.irb.hr/QueryJobStatus.aspx",
-        query = list( jobid = jobid )
-      ) -> res2
-      dat2 <- httr::content(res2, encoding = "UTF-8")
-      running <- jsonlite::fromJSON(dat2)$running
-      Sys.sleep(1)
-  }
-
-  # Fetch results
-  httr::POST(
-    url = "http://revigo.irb.hr/ExportJob.aspx",
-    query = list(
-      jobid = jobid, 
-      namespace = "1",
-      type = "CSVTable"
-    )
-  ) -> res3
-
-  dat3 <- httr::content(res3, encoding = "UTF-8")
-
-  dat3 <- stri_replace_all_fixed(dat3, "\r", "")
-  # Now we have a csv table in a string!
-
-  # read_csv does not like the ', ', it wants ','
-  dat <- read_csv(gsub(', ', ',', dat3), show_col_types = FALSE)
-
-  # do we even have results?
-  if(nrow(dat) == 0){next}
-  results_list[[f]] <- dat
-
+  #httr::POST(
+  #  url = "http://revigo.irb.hr/StartJob",
+  #  body = list(
+  #    cutoff = "0.7",
+  #    valueType = "pvalue",
+  #   # speciesTaxon = "4577", # zea mays
+  #   #speciesTaxon = '39947', # japonica
+  #    speciesTaxon = '3702', # arabidopsis
+  #    measure = "SIMREL",
+  #    goList = go_and_pvalues
+  #  ),
+  #  # application/x-www-form-urlencoded
+  #  encode = "form"
+  #) -> res
+# 
+#   dat <- httr::content(res, encoding = "UTF-8")
+#   
+#   if(typeof(dat)!="list")
+#   {
+#     jobid <- jsonlite::fromJSON(dat,bigint_as_char=TRUE)$jobid
+#   } else {
+#     jobid <- dat$jobid
+#   }  
+#   # Check job status
+#   running <- "1"
+#   while (running != "0" ) {
+#       httr::GET(
+#         url = "http://revigo.irb.hr/QueryJob",
+#         query = list( jobid = jobid, type="jstatus" )
+#       ) -> res2
+#       dat2 <- httr::content(res2, encoding = "UTF-8")
+#       # fix when httr::content automatically converts json to list
+#       if(typeof(dat2)!="list")
+#       {
+#         running <- jsonlite::fromJSON(dat2)$running
+#       } else {
+#         running <- dat2$running
+#       }
+#       Sys.sleep(1)
+#   }
+#   
+#   # Fetch results
+#   httr::GET(
+#     url = "http://revigo.irb.hr/QueryJob",
+#     query = list(
+#       jobid = jobid, 
+#       namespace = "1",
+#       type = "table"
+#     )
+#   ) -> res3
+#   dat3 <- httr::content(res3, encoding = "UTF-8")
+#   
+#   dat3 <- stri_replace_all_fixed(dat3, "\r", "")
+#   # Now we have a csv table in a string!
+# 
+#   # read_csv does not like the ', ', it wants ','
+#   dat <- read_tsv(gsub(', ', ',', dat3), show_col_types = FALSE) 
+#   
+#   # do we even have results?
+#   if(nrow(dat) == 0){next}
+#   results_list[[f]] <- dat
+#   
 }
@@
 for (f in names(results_list)) {
-  dat <- results_list[[f]]
-  if(nrow(dat) == 0) {next}
 
-  names(dat) <- c("term_ID","description","frequency", 'plot_X', 'plot_Y', 'log_size', 'value', 'uniqueness', 'dispensability', 'eliminated', 'representative')
-  one.data <- dat
-  if (one.data[1,] == '<html>') {next} # some datasets result in error
-  one.data <- one.data [(one.data$plot_X != "null" & one.data$plot_Y != "null") & (! is.na(one.data$frequency)) & (one.data$value != 'null'), ];
-  one.data <- suppressMessages(type_convert(one.data)) # guess data types, but also: shush
-  top_ten_values <- one.data %>% arrange(value) %>% head(8) %>% pull(term_ID)
-
-  one.data <- one.data %>% mutate(description2 = case_when(
-    term_ID %in% top_ten_values ~ str_wrap(description, width=10), 
-    TRUE ~ ''))
-  p1 <-
-    ggplot(data = one.data, aes(plot_X, plot_Y, label = description2)) +
-    geom_point(aes(colour = value, size = log_size), alpha = I(0.6)) +
-    scale_size_area() +
-    scale_colour_gradientn(colours = rev(wes_palette("Zissou1", 100, type = "continuous")), 
-                           limits = c(min(one.data$value), 0)) +
-    geom_point(
-      aes(plot_X, plot_Y, size = log_size),
-      shape = 21,
-      fill = "transparent",
-      colour = I (alpha ("black", 0.6))
-    ) +
-    scale_size_area() +
-    geom_label_repel(max.overlaps=15,
-                     point.padding = 0,
-                     min.segment.length = 0, # draw all line segments
-                     aes(point.size=log_size), alpha=0.9) + 
-    theme_minimal() +
-    labs(color = 'log(p-value)',
-         size = 'Frequency') +
-    theme(
-      axis.text.x = element_blank(),
-      axis.ticks.x = element_blank(),
-      axis.text.y = element_blank(),
-      axis.ticks.y = element_blank(),
-      axis.title.x = element_blank(),
-      axis.title.y = element_blank()
-    ) +
-    NULL
-    #ggtitle(f)
-
-  plot_list[[f]] <- p1
-
+  tmp <- results_list[[f]]
+  simMatrix <- tmp[['simMatrix']]
+  reducedTerms <- tmp[['reducedTerms']]
+  if(nrow(reducedTerms) <= 2) {next}
+  plot_list[[f]] <- scatterPlot(simMatrix, reducedTerms, labelSize=3)
+  # old REVIGO code
+  # dat <- results_list[[f]]
+  # if(nrow(dat) == 0) {next}
+  # 
+  # names(dat) <- c("term_ID","description","frequency", 'plot_X', 'plot_Y', 'log_size', 'value', 'uniqueness')#, 'dispensability', 'eliminated', 'representative')
+  # one.data <- dat
+  # #if (one.data[1,] == '<html>') {next} # some datasets result in error
+  # #one.data <- one.data [(one.data$plot_X != "null" & one.data$plot_Y != "null") & (! is.na(one.data$frequency)) & (one.data$value != 'null'), ];
+  # #one.data <- suppressMessages(type_convert(one.data)) # guess data types, but also: shush
+  # top_ten_values <- one.data %>% arrange(value) %>% head(8) %>% pull(term_ID) 
+  # 
+  # one.data <- one.data %>% mutate(description2 = case_when(
+  #   term_ID %in% top_ten_values ~ str_wrap(description, width=10), 
+  #   TRUE ~ ''))
+  # p1 <-
+  #   ggplot(data = one.data, aes(plot_X, plot_Y, label = description2)) +
+  #   geom_point(aes(colour = value, size = log_size), alpha = I(0.6)) +
+  #   scale_size_area() +
+  #   scale_colour_gradientn(colours = rev(wes_palette("Zissou1", 100, type = "continuous")))+ #, 
+  #                          #limits = c(min(one.data$value), 0)) +
+  #   geom_point(
+  #     aes(plot_X, plot_Y, size = log_size),
+  #     shape = 21,
+  #     fill = "transparent",
+  #     colour = I (alpha ("black", 0.6))
+  #   ) +
+  #   scale_size_area() +
+  #   geom_label_repel(max.overlaps=15,
+  #                    point.padding = 0,
+  #                    min.segment.length = 0, # draw all line segments
+  #                    aes(point.size=log_size), alpha=0.9) + 
+  #   theme_minimal() +
+  #   labs(color = 'log(p-value)',
+  #        size = 'Frequency') +
+  #   theme(
+  #     axis.text.x = element_blank(),
+  #     axis.ticks.x = element_blank(),
+  #     axis.text.y = element_blank(),
+  #     axis.ticks.y = element_blank(),
+  #     axis.title.x = element_blank(),
+  #     axis.title.y = element_blank()
+  #   ) +
+  #   NULL
+  #   #ggtitle(f)
+  # 
+  # plot_list[[f]] <- p1
+  # 
 }
@@
   print(i)
-  print(results_list[[i]] %>% dplyr::filter(Eliminated == FALSE) %>% dplyr::select(TermID, Name, Value) %>% arrange(Value) %>% kbl() %>%  kable_styling())
+  #print(results_list[[i]] %>% dplyr::select(TermID, Name, Value) %>% arrange(Value) %>% kbl() %>%  kable_styling())
+  print(results_list[[i]]$reducedTerms %>% dplyr::select(go, parent, score) %>% arrange(score) %>% kbl() %>%  kable_styling())
   cat("\n")
@@
 ```{r}
-a <- list(`P. australis` = results_list$GO_results_genes_lost_P_australis_not_other_seagrasses.csv$Name,
-          `A. antarctica` = results_list$GO_results_genes_lost_A_antarctica_not_other_seagrasses.csv$Name,
-          `Z. marina` = results_list$GO_results_genes_lost_Z_marina_not_other_seagrasses.csv$Name,
-          `Z. muelleri` = results_list$GO_results_genes_lost_Z_muelleri_not_other_seagrasses.csv$Name)
+a <- list(`P. australis` = results_list$GO_results_genes_lost_P_australis_not_other_seagrasses.csv$reducedTerms$term,
+          `A. antarctica` = results_list$GO_results_genes_lost_A_antarctica_not_other_seagrasses.csv$reducedTerms$term,
+          `Z. marina` = results_list$GO_results_genes_lost_Z_marina_not_other_seagrasses.csv$reducedTerms$term,
+          `Z. muelleri` = results_list$GO_results_genes_lost_Z_muelleri_not_other_seagrasses.csv$reducedTerms$term)
 
-a_go_ids <- list(`P. australis` = results_list$GO_results_genes_lost_P_australis_not_other_seagrasses.csv$TermID,
-          `A. antarctica` = results_list$GO_results_genes_lost_A_antarctica_not_other_seagrasses.csv$TermID,
-          `Z. marina` = results_list$GO_results_genes_lost_Z_marina_not_other_seagrasses.csv$TermID,
-          `Z. muelleri` = results_list$GO_results_genes_lost_Z_muelleri_not_other_seagrasses.csv$TermID)
+a_go_ids <- list(`P. australis` = results_list$GO_results_genes_lost_P_australis_not_other_seagrasses.csv$reducedTerms$go,
+          `A. antarctica` = results_list$GO_results_genes_lost_A_antarctica_not_other_seagrasses.csv$reducedTerms$go,
+          `Z. marina` = results_list$GO_results_genes_lost_Z_marina_not_other_seagrasses.csv$reducedTerms$go,
+          `Z. muelleri` = results_list$GO_results_genes_lost_Z_muelleri_not_other_seagrasses.csv$reducedTerms$go)
 ```
@@
 ```{r}
-a_no_ara <- list(`P. australis` = results_list$missing_posi_vs_all_GO.txt$Name,
-          `A. antarctica` = results_list$missing_amphi_vs_all_GO.txt$Name,
-          `Z. marina` = results_list$missing_zmar_vs_all_GO.txt$Name,
-          `Z. muelleri` = results_list$missing_zmuel_vs_all_GO.txt$Name)
+a_no_ara <- list(`P. australis` = results_list$missing_posi_vs_all_GO.txt$reducedTerms$term,
+          `A. antarctica` = results_list$missing_amphi_vs_all_GO.txt$reducedTerms$term,
+          `Z. marina` = results_list$missing_zmar_vs_all_GO.txt$reducedTerms$term,
+          `Z. muelleri` = results_list$missing_zmuel_vs_all_GO.txt$reducedTerms$term)
 ```

### 2023-11-22 analysis/metagenome.Rmd
new Venn plots

@@
                                 str_detect(gene, 'sctC') ~ '*sctC*')) |>
-  #group_by(Sample_species, Tissue, Sample_replicate, MAG) |> 
-  #summarise(gene_class = paste0(sort(unique(gene_class)), collapse = ' & ')) |>
   group_by(Sample_species, Tissue, Sample_replicate) |> 
@@
   theme_minimal() +
-  labs(color='Species', fill='Species') +
-  theme(axis.text.y =  element_markdown(),legend.text = element_markdown(),
+  labs(color='Host species', fill='Host species') +
+  theme(axis.text.y = element_markdown(),
+        legend.text = element_markdown(),
         strip.text.x = element_markdown()) +
-  ylab('Count of MAGs') + xlab('Species')
+  ylab('Count of MAGs') + 
+  xlab('Tissue')
 ```

### 2023-11-23 analysis/metagenome.Rmd
another try

@@
 
+### GTDB order-level stats
+Some stats. How many MAGs on order level?
+
+```{r}
+amph_pos_tax |> filter(order != 'o__') |> count(order) |> nrow()
+```
+```{r}
+amph_pos_tax |> filter(order != 'o__') |> count(order) |> summarise(total = sum(n))
+```
+MAGs per species:
+
+```{r}
+amph_pos_tax |> group_by(Sample_species) |> filter(order != 'o__') |> count(order) |> summarise(total = sum(n))
+
+```
+
+How many of these are unique to either species?
+```{r}
+in_amph <- amph_pos_tax |> filter(Sample_species == 'Amphibolis antarctica') |> filter(order != 'o__') |> count(order)
+in_posi <- amph_pos_tax |> filter(Sample_species == 'Posidonia australis') |> filter(order != 'o__') |> count(order)
+```
+
+Unique orders in Amphibolis:
+```{r}
+in_amph |> nrow()
+```
+
+Unique orders in Posidonia:
+```{r}
+in_posi |> nrow()
+```
+
+In Amphibolis, not in Posidonia:
+```{r}
+setdiff(in_amph$order, in_posi$order) |> length()
+```
+```{r}
+in_amph |> filter(order %in% setdiff(in_amph$order, in_posi$order)) |> arrange(desc(n)) |> head(5)
+```
+
+In Posidonia, not in Amphibolis:
+
+```{r}
+setdiff(in_posi$order, in_amph$order) |> length()
+```
+```{r}
+in_posi |> filter(order %in% setdiff(in_posi$order, in_amph$order)) |> arrange(desc(n))
+```
+
 ## GTDB by family
@@
 
+### CAT order-level stats
+
+Some more stats here too!
+
+```{r}
+amph_pos_tax_cat |> filter(!is.na(Order)) |> separate(Order, into=c('Order', 'Order_score'), sep=':') |> count(Order) |> nrow()
+```
+```{r}
+amph_pos_tax_cat |> filter(!is.na(Order)) |> separate(Order, into=c('Order', 'Order_score'), sep=':') |> count(Order) |> summarise(total = sum(n))
+```
+MAGs per species:
+
+```{r}
+amph_pos_tax_cat |> group_by(Sample_species)|> filter(!is.na(Order)) |> separate(Order, into=c('Order', 'Order_score'), sep=':') |> count(Order) |> summarise(total = sum(n))
+```
+
+How many of these are unique to either species?
+```{r}
+in_amph <- amph_pos_tax_cat |> filter(Sample_species == 'Amphibolis antarctica') |> filter(!is.na(Order)) |> separate(Order, into=c('Order', 'Order_score'), sep=':') |> count(Order)
+in_posi <- amph_pos_tax_cat |> filter(Sample_species == 'Posidonia australis') |> filter(!is.na(Order)) |> separate(Order, into=c('Order', 'Order_score'), sep=':') |> count(Order)
+```
+
+Unique orders in Amphibolis:
+```{r}
+in_amph |> nrow()
+```
+
+Unique orders in Posidonia:
+```{r}
+in_posi |> nrow()
+```
+
+In Amphibolis, not in Posidonia:
+```{r}
+setdiff(in_amph$Order, in_posi$Order) |> length()
+```
+```{r}
+in_amph |> filter(Order %in% setdiff(in_amph$Order, in_posi$Order)) |> arrange(desc(n)) |> head(5)
+```
+
+In Posidonia, not in Amphibolis:
+
+```{r}
+setdiff(in_posi$Order, in_amph$Order) |> length()
+```
+```{r}
+in_posi |> filter(Order %in% setdiff(in_posi$Order, in_amph$Order)) |> arrange(desc(n))
+```
+
 ## CAT by family
@@
 Nope!
+
+Let's write out the table for Supplementary:
+```{r}
+amph_pos_tax_cat |> full_join(amph_pos_tax, by =c('sample', 'MAG')) |> select(-c(Sample_replicate.y, Sample_species.y, Tissue.y, TotalMAGs.y)) |> writexl::write_xlsx('./output/seagrass_metagenome_taxonomy.xlsx')
+```
 
